### 2019-12-05 docs/isr-article/article.tex
Alterations to ISR-Article

%% fsip2pnupsg.tex
%% V0.1
%% 2019/11/27
%% by Francisco Barros

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}

\usepackage{cite}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{enumerate}
\usepackage{optidef}
\usepackage{subfigure}
\usepackage{algorithmic}
\usepackage{algorithm}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}[theorem]{Remark}

\IEEEoverridecommandlockouts
% This command is only needed if you want to use the \thanks command
\overrideIEEEmargins
% paper

\title {File Survivability in P2P Networks using Probabilistic Swarm Guidance}
\author {Francisco Barros, Daniel Silvestre and Carlos Silvestre
\thanks {F. Barros is with the Instituto Superior Técnico, Universidade de  Lisboa, Lisbon, Portugal, {\tt\small fbarros@isr.ist.utl.pt}}
\thanks {D. Silvestre is with the Department of Electrical and Computer Engineering of the Faculty of Science and Technology of the University of Macau, Macau, China, and with the Institute for Systems and Robotics, Instituto Superior T\'ecnico, Universidade de Lisboa, Lisbon, Portugal., {\tt\small dsilvestre@isr.ist.utl.pt}}
\thanks {C. Silvestre is with the Department of Electrical and Computer Engineering of the Faculty of Science and Technology of the University of Macau, Macau, China, on leave from Instituto Superior T\'ecnico, Universidade de Lisboa, Lisboa, Portugal, {\tt\small csilvestre@umac.mo}}
\thanks {This work was partially supported by the project MYRG2016-00097-FST of the University of Macau; by the Portuguese Fundação para a Ciência e a Tecnologia (FCT) through Institute for Systems and Robotics (ISR), under Laboratory for Robotics and Engineering Systems (LARSyS) project UID/EEA/50009/2019.}}

\begin{document}
\renewcommand{\baselinestretch}{1}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

\begin{abstract}
	\textbf{Qual o problema que vamos considerar}

	\textbf{Principais desafios que o problema coloca}

	\textbf{Abordagem utilizada}

	\textbf{Resultados teóricos e em simulação}
\end{abstract}

\begin{keywords}
	Agents-based Systems; Cooperative Control; Distributed control.
\end{keywords}

\section{Introduction}
Peer-to-Peer (P2P) networking is a distributed systems architecture that can be broadly defined as a set of equally privileged peers that contribute with a portion of their resources in order to achieve common goals. Since the beginning of the century, these networks have gained popularity \cite{ssaroiu:msp2pfss} among file sharing and streaming applications, due to their adaptability, scalability, self-organized behavior, lack of centralization, among others. P2P networks can also be leveraged by communities for availability and storage purposes as peers can replicate each others' files decreasing the odds of file loss due to the failure of a small number of machines. On the other hand, the open and volatile nature of such networks makes security, efficiency and performance guarantee harder to accomplish.\newline

\textbf{Qual a abordagem frequente e o porquê de estarmos a introduzir esta técnica:} todo\newline

\textbf{literature survey de redes P2P principais:} procura por: survaviability, digital preservation, cloud computing\newline

\textbf{The main contributions of this paper can be summarized as follows:}
\begin{itemize}
	\item colocar contribuicao
	\item colocar contribuicao
	\item colocar contribuicao
\end{itemize}

The remainder of this paper is organized as follows. In Section \ref{sec:problemstatement}, it is presented background material and defined the problem being addressed. The rendezvous algorithm for the multi-agent system with limited communication capabilities and uncertain agent position is given in Section \ref{sec:proposedsolution}. Section \ref{sec:convergence} presents results regarding the convergence and guarantees of the method while simulations are provided in Section \ref{sec:simulations}. Conclusions and directions of future work are offered in Section \ref{sec:conclusions}.

\section{Problem statement} \label{sec:problemstatement}
\textbf{definição objectiva do problema com as variáveis, limitações, etc}

\textbf{aqui podes recorrer a figuras para ilustrar o problema}

\section{Proposed Solution} \label{sec:proposedsolution}
\textbf{definir a solução, algortimo, etc. também podes usar figuras}

\textbf{podes usar pseudo code como o exemplo}

\algsetup{indent=2em}

\begin{algorithm}
	\caption{Rendezvous algorithm for a non-connected network topology.}\label{alg:rendezvous}
	\begin{algorithmic}[1]
		\REQUIRE Size of strip $h$, mission plane limits and movement tolerance $\epsilon$.
		\ENSURE Rendezvous of nodes to clusters.

		\medskip
		\STATE /* \textit{Initialize phase variable} */
		\STATE $\textrm{phase} = 0$

		\FOR{{\bf each} $k>0$}
			\FOR{{\bf each} $i$}
				\STATE /* \textit{Tower finds neighbor set $\mathcal{N}_{i}(k)$} */
%				\STATE Compute $\mathcal{N}_{i}(k)$ using \eqref{eq:neighbors}
				\STATE /* \textit{Send measurements to node $i$} */
				\STATE Send $\{X_{j}(k): j \in \mathcal{N}_{i}(k)\}$ to node $i$
				\STATE /* \textit{Node computes direction of movement} */
%				\STATE Compute $v_{i}(k)$ using \eqref{eq:directionVector}
				\IF{$\textrm{phase} == 1$}
					\STATE /* \textit{Reverse $v_{i}(k)$ in Convergence Phase} */
					\STATE $v_{i}(k) = -v_{i}(k)$
				\ENDIF
				\STATE /* \textit{Node finds step size} */
%				\STATE Find $a_{i}(k)$ using \eqref{eq:stepsize}
				\STATE /* \textit{Move action} */
%				\STATE Update position using \eqref{eq:update}
			\ENDFOR
			\STATE /* \textit{Tower checks movements} */
			\IF{$\textrm{totalMovement} \leq \epsilon$}
				\STATE $\textrm{phase} = 1$
			\ENDIF
		\ENDFOR
	\end{algorithmic}
\end{algorithm}

\section{Convergence Analysis} \label{sec:convergence}
\textbf{esta para já não é para mexer}

\section{Simulation Results} \label{sec:simulations}
\textbf{descrever o setup das simulações}

\textbf{descrever o que se pode observar dos gráficos apresentados e que é relevante}

\textbf{devemos mostrar casos que consideremos utéis. Por exemplo v>= 0 e v>0 têm de aparecer, com o sem matriz de markov óptima, etc}

\textbf{codigo exemplo para adicionar figuras}
\begin{figure}
	\centering
%	\includegraphics[scale=0.65, trim=100 265 115 255, clip=true]{images/simulation200.pdf}
	\caption{Statistics of the final number of clusters in 200 simulations using the number of nodes ranging from 3 to 40.}
	\label{fig:sim200}
\end{figure}

\section{Conclusions and Future Work} \label{sec:conclusions}
\textbf{descrever sumario do que se falou no paper}

\textbf{principais resultados e conclusões das simulações}

\textbf{direcções de trabalho futuro.}

\bibliographystyle{unsrt}
\bibliography{p2p}

\end{document}

### 2019-12-07 docs/thesis-project/article.tex
papers

%% fsip2pnupsg.tex
%% V0.1
%% 2019/11/27
%% by Francisco Barros

%%%%%%%%%%%%%%%%%
% BEGIN IMPORTS %
%%%%%%%%%%%%%%%%%
\documentclass[runningheads]{llncs}

\usepackage{graphicx}
\usepackage{cite}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{enumerate}
\usepackage{optidef}
\usepackage{subfigure}
\usepackage{algorithmic}
\usepackage{algorithm}
%%%%%%%%%%%%%%%
% END IMPORTS %
%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%
% BEGIN DOCUMENT %
%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%
% BEGIN HEADER %
%%%%%%%%%%%%%%%%
\title {File Survivability in P2P Networks using Probabilistic Swarm Guidance
\thanks{This work was partially supported by the project MYRG2016-00097-FST of the University of Macau; by the Portuguese Fundação para a Ciência e a Tecnologia (FCT) through Institute for Systems and Robotics (ISR), under Laboratory for Robotics and Engineering Systems (LARSyS) project UID/EEA/50009/2019.}
}
\titlerunning{Hive}
\author{Francisco Barros, Daniel Silvestre \and Carlos Silvestre}
\authorrunning{F. Barros et al.}
\institute{Instituo Superior Técnico - Taguspark\newline Av. Prof. Doutor Cavaco Silva, 2744-016 Porto Salvo, Portugal
\email{fbarros@isr.ist.utl.pt, dsilvestre@isr.ist.utl.pt, csilvestre@umac.mo}\newline
\url{www.tecnico.ulisboa.pt}
}
\maketitle
%%%%%%%%%%%%%%
% END HEADER %
%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%
% BEGIN ABSTRACT %
%%%%%%%%%%%%%%%%%%
\begin{abstract}
The abstract should briefly summarize the contents of the paper in 150--250 words.
\begin{keywords}Agents-based Systems; Cooperative Control; Peer-to-Peer; Cloud Storage; Distributed control; File Availability; Swarm Guidance;\end{keywords}
\end{abstract}
%%%%%%%%%%%%%%%%
% END ABSTRACT %
%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%
% BEGIN INDEX %
%%%%%%%%%%%%%%%
\section{Index}\label{sec:index}
The remainder of this paper is organized as follows. In Section \ref{sec:intro}...
%%%%%%%%%%%%%
% END INDEX %
%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
% BEGIN INTRODUCTION %
%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}
With the growth of the internet and the emergence of distributed systems, two large-scale computing paradigms have gained popularity throughout the last two decades due to their promise of virtually, almost unlimited, scalability. On one hand,  Peer-to-Peer (P2P) networking can be broadly defined as a set of equally privileged peers that contribute with a portion of their resources in order to achieve common goals. These networks are very popular \cite{ssaroiu:msp2pfss} among file sharing and streaming applications, due to their self-organized behavior, lack of centralization and low-cost. P2P networks can be leveraged by communities for availability purposes as peers can replicate each others' files decreasing the odds of file loss due to the failure of a small number of machines, i.e. P2P based-systems can be used for file storage or file backups. On the other hand Cloud-based systems offer unmatched availability and reliability , which are currently very trendy with services such as Amazon S3, Microsoft Azure and Google-Cloud providing all types of IT
\subsection{Goals}\label{subsec:intro}
\textbf{What is usually done and how are we better:} todo\newline
\textbf{The main contributions of this paper can be summarized as follows:}
\begin{itemize}
	\item colocar contribuicoes
\end{itemize}
%%%%%%%%%%%%%%%%%%%%
% END INTRODUCTION %
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
% BEGIN RELATED WORK %
%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}\label{sec:relatedwork}
%%%%%%%%%%%%%%%%%%%%
% END RELATED WORK %
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN SOLUTION PROPOSAL %
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Solution Proposal}\label{sec:proposal}
%%%%%%%%%%%%%%%%%%%%%%%%%
% END SOLUTION PROPOSAL %
%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN TESTING METHODOLOGY %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Testing Methodology}\label{sec:methodology}
%%%%%%%%%%%%%%%%%%%%%%%%%%%
% END TESTING METHODOLOGY %
%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN WORK SCHEDULE %
%%%%%%%%%%%%%%%%%%%%%%%
\section{Work Schedule}\label{sec:workschedule}
%%%%%%%%%%%%%%%%%%%%%
% END WORK SCHEDULE %
%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%
% BEGIN CONCLUSIONS %
%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}\label{sec:conclusion}
%%%%%%%%%%%%%%%%%%%
% END CONCLUSIONS %
%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%
% BEGIN REFERENCES %
%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{splncs04}
\bibliography{bibliography}
%%%%%%%%%%%%%%%%%%
% END REFERENCES %
%%%%%%%%%%%%%%%%%%

\end{document}
%%%%%%%%%%%%%%%%
% END DOCUMENT %
%%%%%%%%%%%%%%%%

### 2019-12-07 docs/thesis-project/article.tex
da

@@
 \documentclass[runningheads]{llncs}
-
+\usepackage{indentfirst}
 \usepackage{graphicx}
@@
 \usepackage{algorithm}
+
+\newcommand{\SubItem}[1]{{\setlength\itemindent{15pt} \item[-] #1}}
 %%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%
-\title {File Survivability in P2P Networks using Probabilistic Swarm Guidance
-\thanks{This work was partially supported by the project MYRG2016-00097-FST of the University of Macau; by the Portuguese Fundação para a Ciência e a Tecnologia (FCT) through Institute for Systems and Robotics (ISR), under Laboratory for Robotics and Engineering Systems (LARSyS) project UID/EEA/50009/2019.}
+\title {Hives: File Survivability in P2P Networks using Probabilistic Swarm Guidance
+\thanks{This work has the support of the project MYRG2016-00097-FST of the University of Macau; by the Portuguese Fundação para a Ciência e a Tecnologia (FCT) through Institute for Systems and Robotics (ISR), under Laboratory for Robotics and Engineering Systems (LARSyS) project UID/EEA/50009/2019.}
 }
-\titlerunning{Hive}
+\titlerunning{Hives}
 \author{Francisco Barros, Daniel Silvestre \and Carlos Silvestre}
 \authorrunning{F. Barros et al.}
-\institute{Instituo Superior Técnico - Taguspark\newline Av. Prof. Doutor Cavaco Silva, 2744-016 Porto Salvo, Portugal
+\institute{Instituto Superior Técnico - Taguspark\newline Av. Prof. Doutor Cavaco Silva, 2744-016 Porto Salvo, Portugal
 \email{fbarros@isr.ist.utl.pt, dsilvestre@isr.ist.utl.pt, csilvestre@umac.mo}\newline
@@
 %%%%%%%%%%%%%%%
-\section{Index}\label{sec:index}
-The remainder of this paper is organized as follows. In Section \ref{sec:intro}...
+\section{Index}\label{sec:index}\newpage
 %%%%%%%%%%%%%
@@
 \section{Introduction}\label{sec:intro}
-With the growth of the internet and the emergence of distributed systems, two large-scale computing paradigms have gained popularity throughout the last two decades due to their promise of virtually, almost unlimited, scalability.
-On one hand,  Peer-to-Peer (P2P) networking can be broadly defined as a set of equally privileged peers that contribute with a portion of their resources in order to achieve common goals.
-These networks are very popular \cite{ssaroiu:msp2pfss} among file sharing and streaming applications, due to their self-organized behavior, lack of centralization and low-cost.
-P2P networks can be leveraged by communities for availability purposes as peers can replicate each others' files decreasing the odds of file loss due to the failure of a small number of machines, i.e.
-P2P based-systems can be used for file storage or file backups.
-On the other hand Cloud-based systems offer unmatched availability and reliability , which are currently very trendy with services such as Amazon S3, Microsoft Azure and Google-Cloud providing all types of IT
-\subsection{Goals}\label{subsec:intro}
-\textbf{What is usually done and how are we better:} todo\newline
-\textbf{The main contributions of this paper can be summarized as follows:}
+
+\subsection{Problem Description}
+With the growth of the internet and the emergence of distributed systems, two large-scale computing paradigms have gained popularity due to their promise of virtually unlimited scalability.
+On the one hand, we have Peer-to-Peer (P2P) networking, which can be defined as a group of equally privileged peers, who contribute with a portion of their resources, to achieve common goals.
+These networks are popular among file sharing and streaming applications \cite{ssaroiu:msp2pfss}, due to their self-organized behavior, lack of centralization and, low-cost.
+On the other hand, Cloud platforms offer unmatched, on-demand, self-service, availability, and reliability, at a higher cost, and are currently trendy, with companies such as Amazon, Google, and Microsoft offering various ITaaS products to individuals and organizations alike.
+Cloud-based systems are centralized architectures, in which a large number of computers are clustered and managed by master entities, which possibly, become bottlenecks.
+Both paradigms can be used to deploy distributed file-storage or file-backup applications; despite the fact that P2P implementations are cheaper for both companies and their clients, these have a hard-time achieving 99.9\% availability seen Cloud-based implementations, furthermore, depending on the P2P application there may also exist the risk of permanent file loss, making them somewhat unappealing for clients who want to store or backup their files remotely.
+
+\subsection{Our approach}
+The focus of this thesis is to create a structured P2P overlay to be used on a distributed backup system, where clients upload their files to remote peers, who are selling their storage space, without persistence guarantees.
+Probabilistic Swarm Guidance (PSG) and Markov Chain theories \cite{markovchain_approach_to_pga} will be used, along with block-based file replication using erasure code algorithms and a reputation system, to hopefully create a network that is able to keep a file available at any given time, regardless of presence of attackers, network partitions and amount of churn, i.e., no matter how many peers leave/join an overlay.
+
+\subsection
+{Frequent approach}There has been much research on the topic of P2P overlays.
+There are two significant categories of overlays, Structured and Unstructured.
+In the former, the overlay has rigid rules regarding peer placement; they favor look-up speed but have higher maintenance overhead under churn, i.e., the process of peers leaving or joining the overlay.
+Some popular P2P Structured overlays include Chord \cite{chord}, Tapestry \cite{tapestry}, Kademlia \cite{kademlia}, and P-Grid \cite{pgrid}.
+In the latter, peer placement is random, which provides the overlay network with a higher degree of resilience and robustness in the advent of failures.
+Unstructured overlays disseminate information in a gossip-like fashion; thus, look-up operations are slower, resource management is less efficient, but overlay maintenance is often straightforward.
+Popular  Unstructured P2P overlays include FreeNet \cite{freenet}, Gnutella \cite{gnutella} and BitTorrent \cite{bittorrent}.
+Research on multi-layer overlays has not yielded satisfactory results, and implementation is difficult \cite{sotart}, but research on bio-inspired overlays did have promising results, some of these include Self-Chord \cite{selfchord}.
+Finally, there have also been some attempts of transparently integrating Cloud Systems into P2P networks, e.g., CLOUDCAST \cite{cloudcast}, and vice-versa, e.g., Spotify \cite{spotify} and Wuala \cite{wuala}.
+From all the mentioned research, perhaps the one that borrows most similarity with our proposal is CLOUDCAST; however, neither it nor any of the remaining literature, try to create a structured overlay composed of completely autonomous peers, who work together to keep the file alive without enforced persistence based on PSG.
+
+\subsection{Proposal Goals}\label{subsec:intro}
 \begin{itemize}
-	\item colocar contribuicoes
+    \item Implement a Tracker server with light-weight responsibilities:
+        \SubItem{Act as peer discovery service for clients;}
+        \SubItem{Act as an indexer for clusters of peers (Hives);}
+        \SubItem{Track peer reputations;}
+    \item Implement an erasure-code algorithm to ensure file persists in the Hive;
+    \item Implement a trust system where:
+        \SubItem Each peer gives feedback to the Tracker about his neighbors;
+        \SubItem Tracker synthesizes the feedback and presents it to the client;
+    \item Calculate the ideal steady-state distribution of file blocks, for a Hive, when:
+        \SubItem{No trust history on a new peer is available;}
+        \SubItem{Peer availability, reliability and other trust factors are known;}
+    \item Hives converge to ideal block distribution in bounded-times using PSG;
+    \item Efficiently update the PSG's Markov Chain as it changes over time;
+    \item Balance the in-degree and out-degree of each peer such that:
+        \SubItem{Message exchange overhead is minimal;}
+        \SubItem{Any changes to the probability of file loss are negligible;}
+        \SubItem{Hive availability, reliability, and confidentiality are not compromised;}
+        \SubItem{Hive is resilient to massive node failures and resistant to churn;}
+    \item Extra miles may include:
+        \SubItem{Peers can reconstruct Hives without Tracker intervention;}
+        \SubItem{Peers trust their neighbors independently of Tracker;}
+        \SubItem{Hives can leverage dedicated Cloud-Storage if they become unreliable;}
 \end{itemize}

### 2019-12-08 docs/thesis-project/article.tex
Updates

@@
 %%%%%%%%%%%%%%%%
-\title {Hives: File Survivability in P2P Networks using Probabilistic Swarm Guidance
+\title {Hives: File Survivability in P2P using Probabilistic Swarm Guidance
 \thanks{This work has the support of the project MYRG2016-00097-FST of the University of Macau; by the Portuguese Fundação para a Ciência e a Tecnologia (FCT) through Institute for Systems and Robotics (ISR), under Laboratory for Robotics and Engineering Systems (LARSyS) project UID/EEA/50009/2019.}
@@
 
-%%%%%%%%%%%%%%%
-% BEGIN INDEX %
-%%%%%%%%%%%%%%%
-\section{Index}\label{sec:index}\newpage
-%%%%%%%%%%%%%
-% END INDEX %
-%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% BEGIN TABLE OF CONTENTS %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\setcounter{tocdepth}{3}
+\tableofcontents
+%%%%%%%%%%%%%%%%%%%%%%%%%
+% END TABLE OF CONTENTS %
+%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}\label{sec:intro}
+\newpage\section{Introduction}\label{sec:intro}
 
@@
         \SubItem{Any changes to the probability of file loss are negligible;}
-        \SubItem{Hive availability, reliability, and confidentiality are not compromised;}
+        \SubItem{Hive availability, reliability, and, confidentiality are not compromised;}
         \SubItem{Hive is resilient to massive node failures and resistant to churn;}
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}\label{sec:relatedwork}
+\newpage\section{Related Work}\label{sec:relatedwork}
+This section as the following structure: In the first subsection, we will study two distributed file systems used in cloud architectures, in the following two subsections we will study some of the previously mentioned peer-to-peer, unstructured and structured, overlays.
+We will follow with considerations to have when choosing erasure code algorithms for file-block replication, two popular works on trust in peer-to-peer networks, and finally, we will mention one cloud-assisted P2P protocol that we consider relevant for being similar to the use-case that led us to study this thesis topic.
+\subsection{Distributed File Systems}
+\textbf{Google File System} (GFS) is designed to provide efficient and reliable access to data using clusters of commodity machines located in geographically spread data-centers.
+In particular, GFS aims to provide high data throughput, low latency, and reliability in the face of individual server failures.
+Because GFS runs over commodity hardware, the proposed model considers machine and network faults as being the norm rather than the exception \cite{gfs}.
+This assumption is typical in modern distributed systems of any kind.
+GFS is the basis for an open-source project, which has seen continuous improvements over the years, called Hadoop Distributed File System (HDFS) \cite{hadoop}; we will not discuss it here because it holds many similarities to GFS with key differences being the approach to security management.
+One advantage of GFS over other architectures is that it leverages a distributed file system that tailors to the designers observation that, once written, files on google provided services, are only read, often sequentially, with random writes being practically non-existent, shifting optimization focus from latency to throughput, from random writes to appends when it comes to atomicity and no caching of data blocks \cite{gfs}.
+In the GFS architecture cluster consists of a single master and multiple chunk-servers.
+The Master maintains a mapping from files to chunks, the current location of chunks, and is responsible for failure detection over the chunk-servers as well as chunk-migration between them.
+Clients communicate, for reads and writes, with the Master for metadata operations and directly with the chunk-servers for data-bearing operations.
+Neither the client nor the chunk-servers cache file data because most of google's client applications stream through huge files.
+A read operation involves obtaining chunk handles and locations from the Master and then retrieving the chunks from one of the many available chunk-severs.
+For writes, the client splits the file into 64MB chunks and asks the Master to assign them an identifier, and then the client uploads his chunks to the chunk-servers indicated by the Master.
+
+GFS has proven its qualities in supporting large-scale data processing.
+However, small files have shown \cite{gfs} to lead to overloads on chunk-servers holding them because chunks have a fixed large value size of 64MB.
+The reasoning for GFS to use such substantial value is to reduce network overload, reduce the number of accesses to the Master, and reduce memory used at the Master due to mappings.
+Concerning our proposal, Hives, since convergence to a desired steady-state using Markov Chains theory requires a large number of chunks to exist in a P2P Hive, we might have to do file-level replication rather than the desired block-level when client files are too small.\newline
+
+\textbf{Ceph} is an open-source, storage system that, like GFS, runs over commodity hardware.
+Ceph aims to deliver reliable, autonomous, distributed object, block, and file storage that is self-healing, self-managing, and has no single point of failure.
+The base architectural component of Ceph is the RADOS object store, and all content a client reads and writes to the system are represented as objects, irrespective of the objects' original data type.
+RADOS is for all purposes a Ceph cluster, and its storage space originates from the disks of multiple machines.
+Ceph architecture is slightly more complicated than that of GFS, so we will summarize the components that are of interest to us.
+Inside a cluster, we find three types of machines, storage devices, abstracted with a layer called Object Software Daemon (OSD), Monitors, and Metadata Servers.
+A small count of Monitors maintain cluster membership and state using PAXOS to make consensual decisions.
+Metadata Servers only required sharing the distributed file system, are used for POSIX semantics, e.g., directory hierarchy and file metadata such as owner and last accessed timestamps.
+Monitors can detect if metadata servers are down and launch new ones as required.
+Finally, OSDs store clients' data, and there are anywhere between dozens to thousands of these in a Ceph cluster.
+
+The innovation Ceph brings is that the OSDs within a cluster are organized in a P2P fashion and are capable of autonomously carrying out replication and recovery tasks.
+When one OSD suspects a peer as died or joined the cluster, it consults the Monitors to obtain an updated cluster map and possibly offloads some of his data to some other living OSD.
+The way an OSD knows based on the cluster map if he needs to send data to another OSD is the same process a client goes through to discover to which OSD he should write or read his data.
+In all cases, a client or OSD process runs the Ceph.CRUSH(placement groups, cluster state, rule set) to obtain the Target OSDs of the named object.
+The last two parameters are obtained directly from the monitors.
+The first parameter is the result of the operation: hash(object name) mod (size of object pool).
+CRUSH \cite{crush} is a light-weight pseudo-random placement algorithm, that provides deterministic object quantization without look-ups to central directories.
+Because it runs independently on clients and peers, it avoids the need to have the monitor constantly updating cluster to object mappings.
+
+Ceph is proof that cloud systems can leverage P2P behavior in order to dynamically adjust replication in a decentralized fashion while meeting high dependability standards.
+\cite{ceph, ceph_benchmarks}
+
+\subsection{Unstructured Overlays}
+\textbf{Kademlia} Kademlia is the first P2P system to combine provable consistency, performance, latency-minimizing routing, and a symmetric topology. \cite{kademlia}
+\subsection{Structured Overlays}
+\subsection{File Block Replication}
+\subsection{Trust in Peer-to-Peer Networks}
+\subsection{Cloud-Assisted P2P Protocol - CLOUDCAST}
 %%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Solution Proposal}\label{sec:proposal}
+\newpage\section{Solution Proposal}\label{sec:proposal}
 %%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% BEGIN TESTING METHODOLOGY %
+% BEGIN EVALUATION METHODOLOGY %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Testing Methodology}\label{sec:methodology}
+\section{Evaluation Methodology}\label{sec:methodology}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-% END TESTING METHODOLOGY %
+% END EVALUATION METHODOLOGY %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-12-10 docs/thesis-project/article.tex
There goes updates

@@
 \usepackage{cite}
-\usepackage{amsmath}
 \usepackage{mathtools}
@@
 \usepackage{algorithm}
+\usepackage{todonotes}
 
@@
 There are two significant categories of overlays, Structured and Unstructured.
-In the former, the overlay has rigid rules regarding peer placement; they favor look-up speed but have higher maintenance overhead under churn, i.e., the process of peers leaving or joining the overlay.
+In the former, the overlay has rigid rules regarding peer placement; they favor lookup speed but have higher maintenance overhead under churn, i.e., the process of peers leaving or joining the overlay.
 Some popular P2P Structured overlays include Chord \cite{chord}, Tapestry \cite{tapestry}, Kademlia \cite{kademlia}, and P-Grid \cite{pgrid}.
 In the latter, peer placement is random, which provides the overlay network with a higher degree of resilience and robustness in the advent of failures.
-Unstructured overlays disseminate information in a gossip-like fashion; thus, look-up operations are slower, resource management is less efficient, but overlay maintenance is often straightforward.
+Unstructured overlays disseminate information in a gossip-like fashion; thus, lookup operations are slower, resource management is less efficient, but overlay maintenance is often straightforward.
 Popular  Unstructured P2P overlays include FreeNet \cite{freenet}, Gnutella \cite{gnutella} and BitTorrent \cite{bittorrent}.
@@
 \newpage\section{Related Work}\label{sec:relatedwork}
-This section as the following structure: In the first subsection, we will study two distributed file systems used in cloud architectures, in the following two subsections we will study some of the previously mentioned peer-to-peer, unstructured and structured, overlays.
-We will follow with considerations to have when choosing erasure code algorithms for file-block replication, two popular works on trust in peer-to-peer networks, and finally, we will mention one cloud-assisted P2P protocol that we consider relevant for being similar to the use-case that led us to study this thesis topic.
+This section as the following structure: In the first subsection, we will study two distributed file systems used in cloud architectures, in the following two subsections we look at four P2P overlays.
+We will then mention one cloud-assisted P2P protocol that we consider relevant for being designed to the same use-case that led us to make this thesis proposal.
+Then we will take a look at two well-known research regarding trust in P2P networks, and finally, we will study erasure code algorithms for file-block replication.
 \subsection{Distributed File Systems}
@@
 Inside a cluster, we find three types of machines, storage devices, abstracted with a layer called Object Software Daemon (OSD), Monitors, and Metadata Servers.
-A small count of Monitors maintain cluster membership and state using PAXOS to make consensual decisions.
+A small count of Monitors maintain cluster membership and state using PAXOS \cite{paxos} to make consensual decisions.
 Metadata Servers only required sharing the distributed file system, are used for POSIX semantics, e.g., directory hierarchy and file metadata such as owner and last accessed timestamps.
@@
 The first parameter is the result of the operation: hash(object name) mod (size of object pool).
-CRUSH \cite{crush} is a light-weight pseudo-random placement algorithm, that provides deterministic object quantization without look-ups to central directories.
+CRUSH \cite{crush} is a light-weight pseudo-random placement algorithm, that provides deterministic object quantization without lookups to central directories.
 Because it runs independently on clients and peers, it avoids the need to have the monitor constantly updating cluster to object mappings.
@@
 
+\subsection{Structured Overlays}
+\textbf{Kademlia} is structured overlay, yet it is fully decentralized, designed for key-value storage systems, which, according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology \cite{kademlia}.
+Kademlia peer-identifiers and data-keys are built as in Chord\cite{chord}, using consistent hashing.
+Inserts and lookups on the overlay are unidirectional, according to the distance between two identifiers, calculated using XOR.
+The unidirectionality allows caching to be done along lookup paths, alleviating hot spots.
+Each peer in the overlay maintains a binary tree, routing table, whose leaves are lists of peers, called k-buckets, at distances $2^{x}$ to $2^{x+1}$, $x \in\hspace{1mm}\mathclose[0,160\mathclose]$ from itself, sorted according to a least recently seen policy.
+As a peer receives messages if the sender is a new acquaintance, that sender is placed on the proper k-bucket if there is enough space in it; otherwise, it will replace the least seen peer only if that peer fails to respond to a ping.
+By preferring old acquaintances, Kademlia seeks to enhance the stability of the overlay based on the observation that the longer a peer remains connected to the network, the higher the probability that it will remain connected at least another hour\cite{ssaroiu:msp2pfss}.
+This policy also prevents some denial of service attacks, where an enemy floods the network with new peers.
+The XOR metric gives Kademlia an advantage over other DHT-based implementations by making routing tables less rigid and by allowing peers in the network to learn useful routing information from discovery and lookup messages.
+Because of the existence of k-buckets, routing can also be done around failed nodes, providing some resistance to churn and partitioning.
+
+The standard Kademlia implementation, however, does not implement an effective mechanism to keep dated entries out of a peer's k-buckets, meaning that the longer a peer stays offline, the worse is k-buckets are.
+Dated entries removed when the peer contacts a peer and obtains no response, meaning that insert and lookup operations may take longer than they could.
+The take-away from Kademlia protocol is that both the liveness and the freshness of an overlay can be improved by passively or actively making peers monitor each other.\newline
+
+\textbf{Self-Chord} is a bio-inspired, structured overlay, specifically designed to be used in cloud computing or grid computing architectures.
+Self-chord foundations are that of the standard Chord, both in terms of identifiers and the way peers organize themselves on the overlay, i.e., in ring-like shapes.
+However, it has a few benefits, which are the result of having mobile agents, called ants, independently roaming the overlay network, but collaboratively re-arranging resource identifiers so that similar data items are clustered around neighboring peers with relatively proportional workloads, meaning that data item placement is not based solely on the numerical-space shared between them and possible placement peers \cite{sotart, selfchord}.
+Hence advantages of Self-Chord over the standard implementation include lack of need to assign data to well-specified peers since keys are independently defined, leading to better behavior under high churn conditions.
+Less overlay maintenance overhead since when a peer joins or leaves the network, ants will autonomously and eventually, distribute data items appropriately; Lastly, since load-balancing is continuously being carried out by ants, the scalability and robustness of the system increases since individual peer failures are less likely to lead to massive file loss.
+According to the authors, the behavior of the ants ensures that even under highly dynamic, unfavorable conditions, rearrangement, and discovery of data items takes only logarithmic time.
+
+Bio-inspired systems like the one above, diminish or eliminate inherent disadvantages typically associated with structured or unstructured overlays, such as the ones mentioned in the introduction of this thesis proposal.
+In these, agents with simple behavior often accomplish complex behavior that would otherwise require time or space consuming algorithms, sometimes with high code complexity.
+In our proposal, we, too, seek to create a system that, just like Self-Chord, has self-healing and self-organizing behaviors.
+
 \subsection{Unstructured Overlays}
-\textbf{Kademlia} Kademlia is the first P2P system to combine provable consistency, performance, latency-minimizing routing, and a symmetric topology. \cite{kademlia}
-\subsection{Structured Overlays}
+\textbf{Freenet} laalalalaa\newline
+
+\textbf{Gia} lalala\newline
+\subsection{Cloud-Assisted P2P Protocol - CLOUDCAST}
 \subsection{File Block Replication}
 \subsection{Trust in Peer-to-Peer Networks}
-\subsection{Cloud-Assisted P2P Protocol - CLOUDCAST}
 %%%%%%%%%%%%%%%%%%%%

### 2019-12-10 docs/thesis-project/article.tex
Updates

@@
 Unstructured overlays disseminate information in a gossip-like fashion; thus, lookup operations are slower, resource management is less efficient, but overlay maintenance is often straightforward.
-Popular  Unstructured P2P overlays include FreeNet \cite{freenet}, Gnutella \cite{gnutella} and BitTorrent \cite{bittorrent}.
+Popular  Unstructured P2P overlays include FreeNet \cite{freenet}, Gnutella \cite{gnutella-rfc} and BitTorrent \cite{bittorrent}.
 Research on multi-layer overlays has not yielded satisfactory results, and implementation is difficult \cite{sotart}, but research on bio-inspired overlays did have promising results, some of these include Self-Chord \cite{selfchord}.
@@
 \subsection{Unstructured Overlays}
-\textbf{Freenet} laalalalaa\newline
+\textbf{Gia} \cite{gia} is a decentralized overlay protocol whose purpose is to overcome the drawbacks of Gnutella\cite{gnutella-rfc, gnutella-cs}.
+Gnutella is the first-ever decentralized, unstructured P2P protocol.
+Gnutella's main issues are lack of scalability, as resource discovery bases itself on simple message flooding mechanisms and how easily some peers become overloaded when faced with high rates of aggregate queries, leading to system degradation.
+Gia proposes the use of super-nodes, which are now also part of the Gnutella protocol.
+The super-nodes receive and route queries to peers holding data, but these super-nodes and the construction of the topology around them is dynamic and adaptive.
+Gia recognizes the heterogeneity and resource constraints of the peers and also uses random walks to alter the topology adaptively as well as lookup operations.
+The results show that Gia is three to five orders of magnitude better than that of Gnutella.
+Random-walks by themselves, while minimizing the network overhead, are less likely to find appropriate responses to the performed lookup, unless they are biased to high-degree peers, but this can make peer overloading more frequent.
+Thus, to tackle both problems, Gia implements a topology adaptation that puts most peers within reach of high capacity peers, while simultaneously ensuring that they can handle the likely incoming requests.
+Lastly, all peers keep pointers to data items kept by their immediate one-hop neighbors.
 
-\textbf{Gia} lalala\newline
+The algorithm Gia uses to ensure high-capacity-peers are also high-degree ones depends on a satisfaction function that returns a value in $\mathclose[0,1\mathclose]$.
+As long as the returned value is not one, the peer takes the initiative of trying to connect itself to another peer at random, preferring those whose capacity is better than his.
+During the handshake, any of the intervenients may abort, based on their capacity and the degrees of their neighbors.
+During this process, the target of the handshake request accepts the requester whenever he does not know enough peers or the requester as more capacity than at least one of his neighbors.
+To avoid disconnecting poorly connected peers, the target of the handshake request will replace the highest capacity peer whose capacity is smaller than that of the requester.
+In order to avoid hot-spots or overloading of peers, a peer is only allowed to send requests to a neighbor if that neighbor has explicitly given him one Token; a token represents one request the neighbor is willing to accept; Each peer allocates tokens at the rate at which they can answer requests.
+
+Gia successfully promotes optimal performance and longevity on the systems overlay topology through continuous optimization.
+Unlike many other unstructured overlays, Gia does not use message flooding to disseminate information, minimizing overhead.
+There is, however, one fundamental problem with Gia, related to the promotion of fairness; according to their capacity, peers, receive tokens from their neighbors, hence decreasing the usability of the system by low capacity peers.
+Either way, free riding is not a concept that directly applies to our proposal.\newline
+
+\textbf{BlatAnt} \cite{blatant} is a bio-inspired, unstructured overlay, designed for grid computing architectures where resource discovery needs to be efficient; to this end, BlatAnt focuses solely on bounded-diameter optimization in order to minimize network overhead and lookup latency, by using swarm intelligence to support flood-like discovery protocols.
+Like in Self-Chord\cite{selfchord}, ant agents are used to optimize and maintain a self-structured P2P overlay, conversely, in BlatAnt the ants do not change the location of data items from peer to peer, but instead build and optimize the overlay such that the network uses a minimal number of connections between peers and such that the diameter of the overlay is bounded.
+The idea is that by doing so, the average number of hops, i.e., average path length, required to obtain a data item from a lookup will be smaller, thus achieving the sought goals.
+BlatAnt also uses local index caching to improve efficiency further; The authors argue that local caches,  as an alternative to semantic clustering such as the ones used in Antares\cite{antares} and Self-Chord, are better for self-organized overlays where a stable topology can not be guaranteed.
+The overlay management can be viewed in three major components.
+
+The first component are the peers themselves.
+Each peer keeps two structures, a   fixed-size $\alpha$-table, i.e., a partial view of the overlay retaining neighborhood information used to evaluate the redundant connections or the need for new ones, and, a neighborhood-set \textit{N}, containing peer identifiers.
+Peers contribute to the optimization by rearranging local connections according to the connection rule, which reduces the diameter of the overlay, and disconnection rule, which discards redundant connections.
+
+The second component are the ant agents, which have multiple species; \textit{discovery ants} are randomly spawned by peers and live for a limited time.
+They wander across the overlay, collecting information about its topology and update the $\alpha$-tables of peers they meet along the way.
+\textit{Construction ants} act as bootstrappers to the system.
+When a new peer \textit{i} wants to connect to a peer \textit{j}, he sends one of these ants, if \textit{j} cannot establish the connection because he would violate a degree constraint, he forwards the ant to his lowest-degree neighbor.
+When a \textit{j'} accepts the ant, he sends it back to \textit{i}, and the procedure complete, i.e., they are now on each others' neighborhood-sets.
+\textit{Optimization ants} establish connections according to the connection rule.
+In this case, a peer \textit{i} wanting to connect to peer \textit{j} sends him this ant, and \textit{j} only accepts or refuses the ant.
+If he accepts, he sends the ant back to \textit{i}.
+\textit{Unlink ants} remove existing connections between peers because of the disconnection rule applies or because a peer wants to leave the overlay.
+When this ant arrives at its destination, it removes all sender's information from the $\alpha$-table and the neighborhood-set.
+\textit{Update Neighbors Ants} are spawned by a peer whenever he establishes or tears-down a connection with some other.
+These ants visit all of that peer's neighbors and update his information on their respective $\alpha$-table.
+Finally, \textit{Ping Ants} are periodically exchanged between peers to keep their connections alive in low traffic situations.
+
+The last component are pheromone trails, which evaporate overtime.
+A trail is a value assigned to a connection between peers, whenever an ant agent, or a query, walks over that connection, they increase the pheromone concentration increases on both ends.
+Discovery ants will tend to follow paths with less pheromone concentration, thus incentivizing full network coverage during the exploration.
+When a peer detects that for a given neighbor, its pheromone concentration has wholly evaporated, that neighbor is assumed to have left the network without warning.
+Eventually, each peer who neighbored the disconnected peer will independently initiate a recovery protocol by sending construction ants to the neighbors of the disconnected peer, reorganizing the overlay, and avoiding network partitioning.
+Thus pheromone trails provide seamless error resolution.
+
+BlatAnt proved to accomplish its goals, quickly converging to stable overlays with bounded-diameter, even under very dynamic conditions \cite{sotart}, resulting in faster lookups due to the reduction of flood messages but also due to caching.
+Note that BlatAnt can leverage random walks such as the ones used in Gia for even less network overhead.
+BlatAnt also introduces optimizations over caching procedures; we do not review them here, because they are based on complex profile similarities and because caching is out of the scope of our proposal.
+
+\subsection{Membership Management Framework - CYCLON}
 \subsection{Cloud-Assisted P2P Protocol - CLOUDCAST}
@@
 \subsection{Trust in Peer-to-Peer Networks}
+\subsection{Probabilistic Swarm Guidance}
 %%%%%%%%%%%%%%%%%%%%

### 2019-12-11 docs/thesis-project/article.tex
updates

@@
 \begin{abstract}
-The abstract should briefly summarize the contents of the paper in 150--250 words.
+Peer-to-Peer (P2P) systems have emerged as a potential choice to build large scale distributed storage systems, at a fraction of the cost of the alternative cloud approaches, which have a better quality of service guarantees.
+We survey general-purpose P2P protocols, distributed file systems, cloud-assisted P2P protocols, and erasure-code algorithms.
+We then draft a solution for the problem of creating a bio-inspired P2P overlay to be used on a cloud-assisted P2P storage and backup system using probabilistic swarm guidance.
+We anticipate that our solution will have high network overhead, in order to provide self-healing properties to the files stored in the system.
+
 \begin{keywords}Agents-based Systems; Cooperative Control; Peer-to-Peer; Cloud Storage; Distributed control; File Availability; Swarm Guidance;\end{keywords}
@@
         \SubItem{Any changes to the probability of file loss are negligible;}
-        \SubItem{Hive availability, reliability, and, confidentiality are not compromised;}
+        \SubItem{Hive availability and reliability are not compromised;}
         \SubItem{Hive is resilient to massive node failures and resistant to churn;}
@@
 Note that BlatAnt can leverage random walks such as the ones used in Gia for even less network overhead.
-BlatAnt also introduces optimizations over caching procedures; we do not review them here, because they are based on complex profile similarities and because caching is out of the scope of our proposal.
+BlatAnt also introduces optimizations over caching procedures; we do not review them here, because they are based on complex profile similarities and because caching is out of the scope of our proposal.\newline
 
-\subsection{Membership Management Framework - CYCLON}
+\textbf{CYCLON} \cite{cyclon} is a protocol that emerges as an inexpensive solution to the maintenance of gossip-based unstructured overlays.
+CYCLON is useable even in highly dynamic environments, providing peer degree symmetry, low diameter, low clustering, resilience to churn and massive node failures, in a fully decentralized manner.
+One fundamental issue of unstructured networks is to avoid partitioning.
+One naive solution would be for each peer to keep a view of all the peers in the network and create connections has needed, but that solution is not scalable.
+
+The protocol is simple.
+It consists of having each peer keeping an ever-changing, partial view, with fixed-size \textit{c}, of the network in a structure called \textit{neighbors}.
+Periodically and independently, a contacts a random neighbor to exchange acquaintances.
+This process is called \textit{shuffling}.
+During a shuffle, a peer \textit{i}, first increments the age of all his known neighbors by one.
+Then, selects the oldest of all his neighbors, \textit{j}, and $l-1$ other random neighbors, forming \textit{s}.
+Then \textit{i} replaces \textit{j}'s network address by his own and sets his age to zero, in \textit{s}.
+After that, \textit{i} sends \textit{s} to \textit{j} and waits for \textit{j} to reply with \textit{s'}.
+Unlike \textit{s}, \textit{s'} is composed of a purely random subset of \textit{j}'s neighbors, without any age modifications.
+Upon reception of subsets \textit{s} and \textit{s'}, \textit{j} and \textit{i}, respectively, update their \textit{neighbors} by first using any empty slots in their partial views and then by replacing entries which they sent to the other, i.e., if \textit{i} has no more slots, he replaces, in its \textit{neighbors} structure, the entries that he sent in \textit{s} to \textit{j}, with the ones he received in \textit{s'} from \textit{j}.
+Note that, after \textit{i} initiates a shuffle with \textit{j}, \textit{i} becomes \textit{j}'s neighbor, but \textit{j} is no longer a neighbor to \textit{i}, i.e., the neighboring relation between \textit{i} and \textit{j} is reversed.
+Unlike many protocols, CYCLON links between peers are not bidirectional.
+
+The age is essential in the shuffling algorithm, for two reasons.
+First, it limits the lifetime of each peer in \textit{neighbors} structures, which globally bounds the number of existing pointers to them, it also limits the time each peers' addresses are passed around until they are selected as shuffling targets, resulting on a more up-to-date overlay as well as uniform distribution of each peers' addresses over the network.
+
+When a peer leaves the overlay for any reason, the remaining peers may have to remove him from their \textit{neighbors} structure.
+Timely removal is fundamental for the robustness of the overlay.
+CYCLON uses passive detection.
+Whenever a peer \textit{i} contacts \textit{j} for shuffling and obtains no response it assumes \textit{j} is disconnected and removes it from his \textit{neighbors}.
+Since the time for \textit{i} to contact \textit{j} is bounded by the age property in \textit{neighbors}, detection is accelerated.
+
+CYCLON simulations show that the average path length and clustering coefficient, for various configurations, converged to values similar to those found in random graphs, i.e., converged to small-diameter topologies with low clustering.
+Gossip-based protocols often have high clustering, which is undesirable, since that may indicate an increased amount of redundant messages passing through the network and a higher chance of network partitioning.
+Another exciting result of CYCLON is that the average path length increased logarithmically, and clustering decreased exponentially, as the number of peers in the network increased, hence demonstrating scalability and increased robustness.
+
+The authors also propose a join method that does not disrupt the randomness of the obtained overlays, while simultaneously making new peers indistinguishable from old ones.
+Whenever \textit{i} wishes to join the overlay, he contacts one peer \textit{j} already in the overlay who initiates \textit{c} random-walks with time-to-live (TTL) equal to the average path length of the overlay.
+The peer \textit{k} where the random-walk ends, replaces one of his \textit{neighbors} entries with \textit{i} network address, setting \textit{i} age set to zero and sends the replaced entry to \textit{i}.
+Note that even if some of the random walks do not complete due to byzantine failures, \textit{i} will remain connected to the overlay and will eventually find new acquaintances.
+
+We conclude by reinforcing the idea that CYCLON is a scalable and robust solution for peer-sampling, providing good self-healing and self-organization capabilities without complex algorithms or apparent downsides.
+In a way, it relates to BlatAnt since the proposal passively bounds the diameter of the network.
+
 \subsection{Cloud-Assisted P2P Protocol - CLOUDCAST}

### 2019-12-12 docs/thesis-project/article.tex
Updates

@@
 Finally, there have also been some attempts of transparently integrating Cloud Systems into P2P networks, e.g., CLOUDCAST \cite{cloudcast}, and vice-versa, e.g., Spotify \cite{spotify} and Wuala \cite{wuala}.
-From all the mentioned research, perhaps the one that borrows most similarity with our proposal is CLOUDCAST; however, neither it nor any of the remaining literature, try to create a structured overlay composed of completely autonomous peers, who work together to keep the file alive without enforced persistence based on PSG.
+From all the mentioned research, perhaps the one that borrows most similarity with our proposal is CLOUDCAST; however, neither it nor any of the remaining literature we present here, try to create a structured overlay composed of completely autonomous peers, who work together to keep the file at all points in time using PSG.
 
 \subsection{Proposal Goals}\label{subsec:intro}
+This work seeks to address the problem of file survivability in P2P networks; our primary goal is to study, design, and evaluate a P2P overlay that uses probabilistic swarm guidance as a way to maintain a file available to service clients at all times.
+Steps will include:
+
 \begin{itemize}
@@
         \SubItem{Track peer reputations;}
-    \item Implement an erasure-code algorithm to ensure file persists in the Hive;
-    \item Implement a trust system where:
-        \SubItem Each peer gives feedback to the Tracker about his neighbors;
-        \SubItem Tracker synthesizes the feedback and presents it to the client;
+    \item Implement an erasure-code algorithm used by Hives to replicate clients' files;
     \item Calculate the ideal steady-state distribution of file blocks, for a Hive, when:
-        \SubItem{No trust history on a new peer is available;}
+        \SubItem{No history information on a peer is available;}
         \SubItem{Peer availability, reliability and other trust factors are known;}
     \item Hives converge to ideal block distribution in bounded-times using PSG;
-    \item Efficiently update the PSG's Markov Chain as it changes over time;
-    \item Balance the in-degree and out-degree of each peer such that:
+    \item Efficiently update routing as Hives' topologies change over time;
+    \item Balance the degree of each peer such that:
         \SubItem{Message exchange overhead is minimal;}
         \SubItem{Any changes to the probability of file loss are negligible;}
-        \SubItem{Hive availability and reliability are not compromised;}
         \SubItem{Hive is resilient to massive node failures and resistant to churn;}
+    \item Evaluating the system using PeerSim \cite{peersim} or equivalent platform;
     \item Extra miles may include:
+        \SubItem{Implement a trust system managed by the Tracker}
         \SubItem{Peers can reconstruct Hives without Tracker intervention;}
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\newpage\section{Related Work}\label{sec:relatedwork}
-This section as the following structure: In the first subsection, we will study two distributed file systems used in cloud architectures, in the following two subsections we look at four P2P overlays.
-We will then mention one cloud-assisted P2P protocol that we consider relevant for being designed to the same use-case that led us to make this thesis proposal.
-Then we will take a look at two well-known research regarding trust in P2P networks, and finally, we will study erasure code algorithms for file-block replication.
-\subsection{Distributed File Systems}
-\textbf{Google File System} (GFS) is designed to provide efficient and reliable access to data using clusters of commodity machines located in geographically spread data-centers.
+\newpage\section{Distributed File Systems}\label{sec:relatedwork}
+\subsection{Google File System}
+Google File System (GFS) \cite{gfs} is designed to provide efficient and reliable access to data using clusters of commodity machines located in geographically spread data-centers.
 In particular, GFS aims to provide high data throughput, low latency, and reliability in the face of individual server failures.
@@
 
-\textbf{Ceph} is an open-source, storage system that, like GFS, runs over commodity hardware.
+\newpage\subsection{Ceph File System}
+Ceph \cite{ceph} is an open-source, storage system that, like GFS, runs over commodity hardware.
 Ceph aims to deliver reliable, autonomous, distributed object, block, and file storage that is self-healing, self-managing, and has no single point of failure.
@@
 
+\newpage\section{P2P Overlays}
 \subsection{Structured Overlays}
-\textbf{Kademlia} is structured overlay, yet it is fully decentralized, designed for key-value storage systems, which, according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology \cite{kademlia}.
+\subsubsection{Kademlia} \cite{kademlia} is structured overlay, yet it is fully decentralized, designed for key-value storage systems, which, according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology \cite{kademlia}.
 Kademlia peer-identifiers and data-keys are built as in Chord\cite{chord}, using consistent hashing.
@@
 
-\textbf{Self-Chord} is a bio-inspired, structured overlay, specifically designed to be used in cloud computing or grid computing architectures.
+\subsubsection{Self-Chord} \cite{selfchord} is a bio-inspired, structured overlay, specifically designed to be used in cloud computing or grid computing architectures.
 Self-chord foundations are that of the standard Chord, both in terms of identifiers and the way peers organize themselves on the overlay, i.e., in ring-like shapes.
@@
 \subsection{Unstructured Overlays}
-\textbf{Gia} \cite{gia} is a decentralized overlay protocol whose purpose is to overcome the drawbacks of Gnutella\cite{gnutella-rfc, gnutella-cs}.
+\subsubsection{Gia} \cite{gia} is a decentralized overlay protocol whose purpose is to overcome the drawbacks of Gnutella\cite{gnutella-rfc, gnutella-cs}.
 Gnutella is the first-ever decentralized, unstructured P2P protocol.
@@
 
-\textbf{BlatAnt} \cite{blatant} is a bio-inspired, unstructured overlay, designed for grid computing architectures where resource discovery needs to be efficient; to this end, BlatAnt focuses solely on bounded-diameter optimization in order to minimize network overhead and lookup latency, by using swarm intelligence to support flood-like discovery protocols.
+\subsubsection{BlatAnt} \cite{blatant} is a bio-inspired, unstructured overlay, designed for grid computing architectures where resource discovery needs to be efficient; to this end, BlatAnt focuses solely on bounded-diameter optimization in order to minimize network overhead and lookup latency, by using swarm intelligence to support flood-like discovery protocols.
 Like in Self-Chord\cite{selfchord}, ant agents are used to optimize and maintain a self-structured P2P overlay, conversely, in BlatAnt the ants do not change the location of data items from peer to peer, but instead build and optimize the overlay such that the network uses a minimal number of connections between peers and such that the diameter of the overlay is bounded.
@@
 
-\textbf{CYCLON} \cite{cyclon} is a protocol that emerges as an inexpensive solution to the maintenance of gossip-based unstructured overlays.
+\newpage\section{Membership Management Protocols}
+\subsection{CYCLON} CYCLON\cite{cyclon} is a protocol that emerges as an inexpensive solution to the maintenance of gossip-based unstructured overlays.
 CYCLON is useable even in highly dynamic environments, providing peer degree symmetry, low diameter, low clustering, resilience to churn and massive node failures, in a fully decentralized manner.
@@
 
-We conclude by reinforcing the idea that CYCLON is a scalable and robust solution for peer-sampling, providing good self-healing and self-organization capabilities without complex algorithms or apparent downsides.
+We conclude by reinforcing that CYCLON is a scalable and robust solution for peer-sampling, providing good self-healing and self-organization capabilities without complex algorithms or apparent downsides other than the one where too few peers in the system may cause it to be under performant.
 In a way, it relates to BlatAnt since the proposal passively bounds the diameter of the network.
 
-\subsection{Cloud-Assisted P2P Protocol - CLOUDCAST}
-\subsection{File Block Replication}
-\subsection{Trust in Peer-to-Peer Networks}
-\subsection{Probabilistic Swarm Guidance}
+\subsection{Newscast}
+Newscast \cite{newscast} is an alternative peer-sampling protocol
+
+\section{Cloud-Assisted P2P Networks}
+
+Many academic and commercial projects have been developed over the years where cloud, grid, and other centralized architectures are partly assisted by P2P networks.
+In these systems, peers are usually machines belonging to the clients of the service, and they help to decrease the load on the system, hence reducing its cost.
+This approach is especially prevalent in streaming and online gaming services, for example, Blizzard Activision uses their player base to help distribute installers and patches in a P2P fashion, and Digital Extremes uses, in their free-to-play game Warframe, P2P networks in particular, but not all, group missions where individuals are in instances of the game inaccessible by players outside the group.
+However, only recently, was the alternative explored, where P2P networks serve the vast majority of the service and are assisted by stable cloud services only to help guarantee some quality of service property of the system under critical situations.
+We present two of them.
+
+\subsection{CLOUDCAST}
+CLOUDCAST \cite{cloudcast} is a P2P-assisted cloud architecture that differs from other proposals because it does not offload the provided service to peers only when service availability is not affected.
+Furthermore, the architecture does not augment P2P systems with elastic computing nodes that perform particular tasks beyond the reach of the P2P network, such has bulk-synchronous content distribution.
+CLOUDCAST functionality depends solely on a passive storage service that transparently integrates into a gossip-based P2P system.
+The reasoning behind not using elastic computing instances is to eliminate fixed renting costs.
+Like that, the focus is on minimizing the costs associated with storage and bandwidth.
+The architecture uses two gossip protocols, one for membership management and another for information dissemination.
+Key results of the proposal include: accesses to the cloud service are bounded regardless of the number of clients in the system; the cloud service only distributes content when few peers are available; the overlay does not suffer from partitioning regardless of the number of peers in the system; the two illustrated use cases have an absurdly low yearly cost when compared to pure cloud approaches, and finally the authors recognize that the vast part of their bill is tied to membership management and not to actual storage costs, thus using strategies that adapt to the size of the network could yield even better savings.
+
+The passive cloud storage service is a key-value store that can be accessed via GET and PUT operations, but it cannot under any circumstance initiate communications with the system peers.
+Two problems are identified, \textit{membership management} and \textit{information diffusion}.
+To solve the first problem, CLOUDCAST uses the CYCLON protocol to do peer-sampling.
+From other alternatives, CYCLON was chosen not for its particular inexpensiveness, but because the protocol maintains a random overlay topology that works even during high churn periods and even in the presence of catastrophic failures where up to 80\% of the peers fail.
+CYCLON properties also guarantee that references to the cloud server are not too many keeping usage cost under control.
+Finally, the cloud server can easily integrate in the protocol, in particular, CYCLON does not require the cloud server to initiate any communications, i.e., its active roles can be carried out by peers interacting with it; Finally, the cloud server can easily integrate into the protocol, in particular, CYCLON does not require the cloud server to initiate any communications, i.e., its active roles can be carried out by peers interacting with it; In this regard, the dummy process carried by the peers occasionally refreshes pointers to the cloud to avoid losing all references to it in the system.
+To deal with the case where all references cloud are lost due to the dynamic characteristics of the environment, e.g.
+churn and lost messages, all peers keep a logical timestamp of the last round they heard about the cloud server updating it whenever they successfully contact it or because one of their a \textit{neighbor} heard about it more recently then them, if after \textit{t} rounds peer has not updated this timestamp he will create a new cloud reference with probability $p\textsubscript{recovery}$.
+
+The second problem is solved using two different implementations of gossip algorithms explained in depth in \cite{epidemic_algorithms}.
+They are a \textit{rumor-mongering, coin and blind} approach which favors fast news spreading and consists in having a peer \textit{i} periodically forwarding updates he learns about to some other peer \textit{j}, then deciding with probability $p\textsubscript{rumor}$ if he should, permanently, stop gossiping the update; And a \textit{anti-entropy, push-pull} approach which guarantees that news are eventually known by all participants of the system and consists in having each peer \textit{i} periodically contacting a random neighbor \textit{j} to exchange update they know about.
+All updates to the data are recorded in the cloud server through a PUT operation before becoming hot-topics, thus triggering rumor-mongering.
+Peers never contact the cloud server for rumor-mongering purposes.
+The anti-entropy approach can be piggybacked to the CYCLON peer-sampling step, thus decreasing message overhead; the authors, however, do not mention this optimization.
+
+\subsection{INSERT ANOTHER}
+\section{Block-Level Data Replication}
+\section{Trust Models}
+\section{Probabilistic Swarm Guidance}
 %%%%%%%%%%%%%%%%%%%%

### 2019-12-12 docs/thesis-project/article.tex
Update

@@
 %%%%%%%%%%%%%%%%%
+\RequirePackage{amsmath}
 \documentclass[runningheads]{llncs}
 \usepackage{indentfirst}
-\usepackage{graphicx}
 \usepackage{cite}
@@
 \newpage\section{P2P Overlays}
+
+\begin{figure}[t]
+    \centering
+    \subfigure{\label{fig:iplayer}\includegraphics[width=0.35\textwidth, clip=true]{images/ip-layer.png}}
+    \hspace{0.2\textwidth}
+    \subfigure{\label{fig:overlay}\includegraphics[width=0.35\textwidth, clip=true]{images/overlay-layer.png}}
+    \caption{P2P Network Diagrams: IP layer \textit{vs.} One possible overlay abstraction}
+    \label{fig:network_diagrams}
+\end{figure}
+
+In large-scale P2P systems, it is unfeasible and even undesirable for each peer in the network to know and explicitly interact with every other node connecting it to other peers.
+In this regard, a P2P overlay is a logical network that abstracts the physical network underneath it, typically IP-based, and where, single-hop edges represent links between pairs of peers, as depicted in Fig.
+\ref{fig:network_diagrams}.
+Before we survey related work in the field of P2P overlays, we first list some essential properties and technical jargon related to overlays and graph modeling, that helps us understand better the pros and cons of the research we are presenting in the next subsections of this thesis proposal.
+
+An overlay is said to be \textbf{\textit{connected}} if there is a path that allows any pair of peers \textit{i, j} to communicate with each other, i.e., there is no partitioning and no peer is isolated in the system.
+
+The \textbf{\textit{degree}} of a peer is the total number of connections established between some peer \textit{i} and other peers.
+It is possible to divide the degree into \textit{in-degree} and \textit{out-degree}, i.e., the number of channels a peer \textit{i} has, through which it can only receive data or send data, respectively.
+When all peers in an overlay have the same degree, the overlay is said to be \textbf{\textit{regular}} or \textbf{\textit{symmetric}}, but this is not usual.
+We can use the \textbf{\textit{degree distribution}} to evaluate the robustness of an overlay since through it; it is possible to detect \textit{weakly connected peers} and \textit{massively connected hubs}, i.e., peers with very high degree.
+In general, we want to minimize the existence of hubs.
+In P2P networks, hubs may provide the overlay with resistance against random damage and possibly speed-up lookup operations in unstructured overlays, using biased searching; however, targeted-attacks to hubs, or natural hub failures will have more significant impacts on the connectivity of the overlay, possibly leading to partitioning or halting of the system \cite{webdragons, controlling_the_hubs}.
+Peer degrees can also allow evaluation of fair work distribution among peers of the system, e.g., in terms of processing or bandwidth.
+
+\textbf{\textit{Network diameter}} is the number of links that compose the longest of all the shortest paths between any two peers \textit{i}, \textit{j}.
+Knowing the length of all shortest paths between all possible pairs of peers \textit{i}, \textit{j} one can extract the \textbf{\textit{average path length}} and hence derive how messages, on average, will be relayed on the overlay to contact a stranger.
+Still regarding distances, the path stretch is the ratio between the number of links a message has to traverse, between any two peers, in the underlying network and the overlay network.
+
+\textbf{\textit{Clustering}} directly affects connectivity and network diameter; we can measure it using the \textbf{\textit{clustering coefficient}}.
+This coefficient represents the percentage of neighbors a peer \textit{i} has, who are also neighbors among themselves.
+Like peers with high degrees, high clustering coefficients indicate the increased chance of partitioning in the system in the event of massive peer failures or high degrees of \textit{churn}.
+Furthermore, high clustering coefficients may indicate, in gossip-based protocols, bandwidth overhead due to redundant communications in the cluster, but has the advantage of making lookups faster even when the network is experiencing heavy load, e.g., flash crowd scenarios.
+
+Another two terms that often appear in P2P literature are the small-world and the power-law properties.
+The small-world property is an umbrella term for a system with a high clustering coefficient and low average path length, yet most peers are not direct neighbors of each other.
+The power-law property means that there is non-uniform distribution influence in the overlay, e.g., some peers have more responsibilities or perks because they have higher capacities than the rest.
+
 \subsection{Structured Overlays}
@@
 
-\subsection{INSERT ANOTHER}
+\subsection{CYCLON-CLOUDCAST: Adaptive Replica Management}
 \section{Block-Level Data Replication}

### 2019-12-13 docs/thesis-project/article.tex
thesis project update

@@
 \newpage\section{Distributed File Systems}\label{sec:relatedwork}
+Distributed file system (DFS) enables individuals and organizations to store, access, and secure data on possibly remote locations, precisely as they would their local machines.
+It is common practice to replicate data at the file-level, block-level, or both, depending on the requirements of each system.
+For example, block-level replication may allow parallel execution of applications, such as is the case of Hadoop, Google, and Spark MapReduce algorithms.
+Here we present two DFS, which follow entirely different approaches when it comes to architecture and protocols.
+We focus only on the architectural implementation, their advantages or disadvantages.
+We do not consider parallel computation because our primary objective, in this thesis proposal, is the continuous availability of data in a system, to clients.
+Before we dive into them, it is essential to understand some desired characteristics of a DFS.
+In general, distributed systems of any kind should be \textbf{\textit{dependable}}, meaning that they should be able to avoid service failures that are more frequent and more severe than is acceptable.
+To be dependable is to be ready to provide the correct service (\textbf{\textit{availability}}), in a continuous manner (\textbf{\textit{reliability}}), with absence of catastrophic consequences to clients (\textbf{\textit{safety}}), e.g., losing a file, with no improper system alterations (\textbf{\textit{integrity}}) and, finally, to be able to undergo modifications and repairs at any time (\textbf{\textit{maintainability}}).
+Features that contribute to the above characteristics are summarized below.
+
+\textbf{\textit{Load-Balancing}} is the process of balancing communication, processing, and storage overhead among all nodes of the system.
+It helps to avoid bottlenecks, increasing operational performance, and helps to prevent catastrophic failures.
+Prevention of failures is also achieved through \textbf{\textit{fault tolerance}}, which enables a system to continue operating, possibly at a reduced level,  rather  than failing completely,  when some part of it fails.
+As the amount of data passing in the network, being processed or stored in the system grows, or the number of clients using it increases, the capacity of a system to maintain its service without, or with little degradation, is called \textbf{\textit{scalability}}, it depends on the implementation of the system, and as a rule of thumb, centralized solutions are less scalable.
+
 \subsection{Google File System}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.8\textwidth]{images/gfs.png}
+\caption{GFS Architecture}\label{fig:gfs_arch}
+\end{figure}
+
 Google File System (GFS) \cite{gfs} is designed to provide efficient and reliable access to data using clusters of commodity machines located in geographically spread data-centers.
@@
 For writes, the client splits the file into 64MB chunks and asks the Master to assign them an identifier, and then the client uploads his chunks to the chunk-servers indicated by the Master.
+See Fig.\ref{fig:gfs_arch} for a simplified architecture diagram.
 
@@
 Ceph aims to deliver reliable, autonomous, distributed object, block, and file storage that is self-healing, self-managing, and has no single point of failure.
-The base architectural component of Ceph is the RADOS object store, and all content a client reads and writes to the system are represented as objects, irrespective of the objects' original data type.
-RADOS is for all purposes a Ceph cluster, and its storage space originates from the disks of multiple machines.
-Ceph architecture is slightly more complicated than that of GFS, so we will summarize the components that are of interest to us.
-Inside a cluster, we find three types of machines, storage devices, abstracted with a layer called Object Software Daemon (OSD), Monitors, and Metadata Servers.
-A small count of Monitors maintain cluster membership and state using PAXOS \cite{paxos} to make consensual decisions.
-Metadata Servers only required sharing the distributed file system, are used for POSIX semantics, e.g., directory hierarchy and file metadata such as owner and last accessed timestamps.
-Monitors can detect if metadata servers are down and launch new ones as required.
-Finally, OSDs store clients' data, and there are anywhere between dozens to thousands of these in a Ceph cluster.
+The base architectural component of Ceph is the \textit{RADOS} object store.
+All content clients read and write to the system is represented as an object, irrespective of the object's original data type.
+\textit{RADOS} is a single logical entity composed of many \textit{Ceph Storage clusters}.
+Ceph architecture, Fig.
+\ref{fig:cfs_arch} is slightly more sophisticated than that of GFS, so we will summarize the components that are of interest to us.
+Inside a \textit{Ceph Storage cluster}, we find four types of logical machines, storage devices, abstracted with a layer called \textit{Object Software Daemon} (OSD), \textit{Monitors}, \textit{Managers}, and \textit{Metadata Servers}.
+\textit{Metadata Servers} are only required when a client his sharing his DFS with other clients, these servers contain information used for POSIX semantics, e.g., file metadata such as owner and last accessed timestamps.
+A small amount of \textit{Monitors} maintain a master copy of the storage cluster map with its current state.
+Monitors require high consistency and use Paxos \cite{paxos} to decide on the state of the cluster.
+\textit{Managers} maintains detailed information about placement groups, process metadata and host metadata.
+Finally, OSD store data on behalf of clients.
+Additionally, OSD utilize their nodes' resources to perform data replication, erasure coding, load-balancing, recovery, monitoring and reporting functions.
+Each Storage Cluster as anywhere from dozens to thousands of these OSDs.
 
-The innovation Ceph brings is that the OSDs within a cluster are organized in a P2P fashion and are capable of autonomously carrying out replication and recovery tasks.
-When one OSD suspects a peer as died or joined the cluster, it consults the Monitors to obtain an updated cluster map and possibly offloads some of his data to some other living OSD.
-The way an OSD knows based on the cluster map if he needs to send data to another OSD is the same process a client goes through to discover to which OSD he should write or read his data.
-In all cases, a client or OSD process runs the Ceph.CRUSH(placement groups, cluster state, rule set) to obtain the Target OSDs of the named object.
-The last two parameters are obtained directly from the monitors.
-The first parameter is the result of the operation: hash(object name) mod (size of object pool).
-CRUSH \cite{crush} is a light-weight pseudo-random placement algorithm, that provides deterministic object quantization without lookups to central directories.
+The innovation Ceph brings is that the OSD within the same \textit{placement group}, within a cluster, are organized in a P2P fashion and are capable of autonomously carrying out replication and recovery tasks.
+\textit{Placement groups} are fragments of a logical \textit{pool}, and in turn, \textit{pools} are logical, dynamic partitions that define among other things, how much and which type of replication is done to objects that belong to it.
+When one OSD in a placement group suspects another as died or joined the cluster, it consults a \textit{Monitor} to obtain an up to date cluster map.
+Based on the cluster map and the \textit{CRUSH}\cite{crush} algorithm, it is possible that this OSD will have to offload some of its content to other OSD.
+Clients also use \textit{CRUSH} to perform reads and writes to the system.
+To perform a write operation, for example, a client connects to a \textit{pool} in the storage cluster and runs \textit{CRUSH} with a cluster map retrieved from a Monitor.
+The result of the operation gives the client information on the placement group of the data he wishes to upload as well as the OSD he should contact to complete his request.
+When the target OSD receives the request, it runs \textit{CRUSH} to find the number of replicas that it should store.
+Then, it takes the object identifier, pool name, and the cluster map and uses \textit{CRUSH}, again, to discover the secondary OSDs to where he will replicate the write.
+The operation is completed when the primary OSD, receives acknowledgments from the remaining secondary OSD and acknowledges back to the client.
+\textit{CRUSH} is a light-weight pseudo-random placement algorithm that provides deterministic object quantization without lookups to central directories.
 Because it runs independently on clients and peers, it avoids the need to have the monitor constantly updating cluster to object mappings.
+Ceph is proof that distributed file systems can reliably be implemented using P2P behavior.
+Ceph OSDs perform a lot of self monitoring and management operations automatically and in a decentralized fashion, while being highly dependable.
+\cite{ceph, ceph_benchmarks}
 
-Ceph is proof that cloud systems can leverage P2P behavior in order to dynamically adjust replication in a decentralized fashion while meeting high dependability standards.
-\cite{ceph, ceph_benchmarks}
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.8\textwidth]{images/cephfs.png}
+\caption{CFS Architecture \cite{configure-ceph}}
+\label{fig:cfs_arch}
+\end{figure}
 
@@
 
-An overlay is said to be \textbf{\textit{connected}} if there is a path that allows any pair of peers \textit{i, j} to communicate with each other, i.e., there is no partitioning and no peer is isolated in the system.
+An overlay is said to be \textbf{\textit{connected}} if there is a path that allows any pair of peers \textit{i}, \textit{j} to communicate with each other, i.e., there is no partitioning and no peer is isolated in the system.
 
@@
 
-\subsection{Unstructured Overlays}
+\newpage\subsection{Unstructured Overlays}
 \subsubsection{Gia} \cite{gia} is a decentralized overlay protocol whose purpose is to overcome the drawbacks of Gnutella\cite{gnutella-rfc, gnutella-cs}.
@@
 
-\newpage\section{Membership Management Protocols}
+\section{Membership Management Protocols}
 \subsection{CYCLON} CYCLON\cite{cyclon} is a protocol that emerges as an inexpensive solution to the maintenance of gossip-based unstructured overlays.

### 2019-12-14 docs/thesis-project/article.tex
Thesis Project upddate

@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=0.8\textwidth]{images/gfs.png}
-\caption{GFS Architecture}\label{fig:gfs_arch}
-\end{figure}
-
 Google File System (GFS) \cite{gfs} is designed to provide efficient and reliable access to data using clusters of commodity machines located in geographically spread data-centers.
@@
 Concerning our proposal, Hives, since convergence to a desired steady-state using Markov Chains theory requires a large number of chunks to exist in a P2P Hive, we might have to do file-level replication rather than the desired block-level when client files are too small.\newline
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=\textwidth]{images/gfs_paper.png}
+\caption{GFS Architecture \cite{gfs}}
+\label{fig:gfs_arch}
+\end{figure}
 
@@
 
-\begin{figure}[t]
+\begin{figure}[!ht]
 \centering
@@
 
-\begin{figure}[t]
+In large-scale P2P systems, it is unfeasible and even undesirable for each peer in the network to know and explicitly interact with every other node connecting it to other peers.
+In this regard, a P2P overlay is a logical network that abstracts the physical network underneath it, typically IP-based, and where, single-hop edges represent links between pairs of peers, as depicted in Fig.
+\ref{fig:network_diagrams}.
+Before we survey related work in the field of P2P overlays, we first list some essential properties and technical jargon related to overlays and graph modeling, that helps us understand better the pros and cons of the research we are presenting in the next subsections of this thesis proposal.
+
+\begin{figure}[ht]
     \centering
@@
 \end{figure}
-
-In large-scale P2P systems, it is unfeasible and even undesirable for each peer in the network to know and explicitly interact with every other node connecting it to other peers.
-In this regard, a P2P overlay is a logical network that abstracts the physical network underneath it, typically IP-based, and where, single-hop edges represent links between pairs of peers, as depicted in Fig.
-\ref{fig:network_diagrams}.
-Before we survey related work in the field of P2P overlays, we first list some essential properties and technical jargon related to overlays and graph modeling, that helps us understand better the pros and cons of the research we are presenting in the next subsections of this thesis proposal.
 
@@
 \subsection{CYCLON-CLOUDCAST: Adaptive Replica Management}
+Following the suggestions of CLOUDCAST authors, H.
+Kavalionak developed a decentralized algorithm \cite{marriage_of_convinience} that aims to do replica management in cloud-based, peer-assisted applications.
+Because the proposed work has no running title, we shall refer to it as ARM.
+The goal of the ARM is to maintain an appropriate replication level, despite churn, such that, at all times, at least one replica is available and ensuring that replicas survive failures.
+Furthermore, the proposal acknowledges that replicas should be synchronized, i.e., the system provides at least some consistency guarantees.
+In simple terms, this is perhaps the main difference between the proposed work and CLOUDCAST, which stores all updates in an always-on passive cloud server, helping with information diffusion and overlay connectivity and overall reliability of the system in extreme conditions.
+CLOUDCAST monetary cost reductions are primarily associated with limiting access to the cloud server.
+ARM effectively reduces those costs even further by performing timely switches between CLOUDCAST and regular CYCLON protocols, depending on the size of the network.
+To estimate the network size without centralization or broadcasting algorithms that would ultimately lead to either bottlenecks or overhead, ARM leverages a gossip-based aggregation method \cite{gossip-based_aggregation} that allows each peer to infer the approximate size of the network autonomously, at the end of each epoch.
+Estimation accuracy is dependent only on the size of the epoch.
+Likewise, the execution cost depends on the length of the epochs and the network churn rate, but not on the number of peers in the system, which means that ARM keeps the scale-free properties of the unstructured overlays that used in CLOUDCAST.
+The peers randomly organize themselves according to the CYCLON peer-sampling protocol, but depending on the results of the aggregation, the passive cloud-server used in CLOUDCAST is removed or kept in the overlay.
+In ARM, each data object to be replicated as its overlay, which facilitates synchronization and cloud-usage decisions.
+
 \section{Block-Level Data Replication}
-\section{Trust Models}
 \section{Probabilistic Swarm Guidance}

### 2019-12-14 docs/thesis-project/article.tex
Update article.tex

@@
 \usepackage{subfigure}
-\usepackage{algorithmic}
 \usepackage{algorithm}
+\usepackage{algpseudocode}
 \usepackage{todonotes}
@@
 \subsection{CYCLON-CLOUDCAST: Adaptive Replica Management}
+
+\begin{algorithm}[ht]
+\caption{MakeDecision(N\textsubscript{e}), replicaOverlay}
+\label{alg:makedecision}
+\begin{algorithmic}
+    \If {$N\textsubscript{e}\geq S$}
+        \State $replicaOverlay.Neighbors.Remove(CloudReference)$\
+        \If {$N\textsubscript{e} > R$}
+            \State $p\textsubscript{yes} \gets \dfrac{N\textsubscript{e} - R}{N\textsubscript{e}}$
+            \State $this.LeaveOverlay(replicaOverlay, p\textsubscript{yes})$
+        \State $replicaOverlay.Protocol \gets CYCLON$
+        \EndIf
+    \ElsIf {$C < N\textsubscript{e} < S$}
+        \State $N\textsubscript{a} \gets [\dfrac{S(1 + k)}{N\textsubscript{e}} - 1]$
+        \State $newNeighbors \gets GetNewNeighbors(this.UnderlyingOverlay, N\textsubscript{a})$
+        \State $newNeighbors.Send(this.RandomView(replicaOverlay))$
+        \State $replicaOverlay.Neighbors.Add(newNeighbors)$
+        \State $replicaOverlay.Protocol \gets CLOUDCAST$
+    \ElsIf{$Ne \leq C$}
+        \State $N\textsubscript{a} \gets [\dfrac{S(1 + k)}{N\textsubscript{e}} - 1]$
+        \State $newNeighbors \gets GetNewNeighbors(this.UnderlyingOverlay, N\textsubscript{a})$
+        \State $newNeighbors.Send(this.RandomView(replicaOverlay) \cup CloudReference)$
+        \State $replicaOverlay.Neighbors.Add(newNeighbors \cup CloudReference)$
+        \State $replicaOverlay.Protocol \gets CLOUDCAST$
+    \EndIf
+\end{algorithmic}
+\end{algorithm}
+
 Following the suggestions of CLOUDCAST authors, H.
@@
 The peers randomly organize themselves according to the CYCLON peer-sampling protocol, but depending on the results of the aggregation, the passive cloud-server used in CLOUDCAST is removed or kept in the overlay.
-In ARM, each data object to be replicated as its overlay, which facilitates synchronization and cloud-usage decisions.
+In ARM, each data object to be replicated as its own \textit{replica-overlay} on top of the regular unstructured overlay, which facilitates synchronization and cloud-usage decisions.
+The ARM algorithm is designed to consider four replica-overlay states, separated by three size thresholds: \textit{Critical (C)}, \textit{Sufficient (S)} and \textit{Redundant (S)}.
+\textit{C} depends on the chosen redundancy method, i.e., file-level \textit{vs.} block-level replication and is the sum of the minimum number of replicas required for data recovery plus the number of peers that will leave during the cloud backup replication phase.
+\textit{S} depends on the churn-rate and is the sum of \textit{C} plus the number of peers expected between two successive recovery phases.
+Finally, \textit{R} is a system parameter dependent also on the chosen redundancy method and is application requirements specific.
+The aggregation method mentioned earlier is the foundation \textit{monitoring phase}, which consists of a fixed number of gossip and idle rounds, allowing a peer to infer the \textit{expected size (N\textsubscript{e})} of the replica-overlay and decide if he should add or remove the cloud or peers to the replica-overlay as well as the protocol he uses, based on Algorithm \ref{alg:makedecision}.
+After each monitoring phase, each peer calculates how long it should wait before executing the \textit{monotoring phase} again.
+That time depends mostly on \textit{N\textsubscript{e}}, e.g., if the state is below the critical threshold the monitoring phase will initiate earlier than if it is above the redundant threshold, allowing the system to respond to overlay changes promptly and to keep the overlay size in a range that minimizes overhead but ensures that data does not become unavailable.
 

### 2019-12-14 docs/thesis-project/article.tex
Thesis Project updates

@@
             \State $this.LeaveOverlay(replicaOverlay, p\textsubscript{yes})$
-        \State $replicaOverlay.Protocol \gets CYCLON$
+        \State $replicaOverlay.Protocol \gets cyclon$
         \EndIf
@@
         \State $replicaOverlay.Neighbors.Add(newNeighbors)$
-        \State $replicaOverlay.Protocol \gets CLOUDCAST$
+        \State $replicaOverlay.Protocol \gets cloudcast$
     \ElsIf{$Ne \leq C$}
@@
         \State $replicaOverlay.Neighbors.Add(newNeighbors \cup CloudReference)$
-        \State $replicaOverlay.Protocol \gets CLOUDCAST$
+        \State $replicaOverlay.Protocol \gets cloudcast$
     \EndIf
@@
 Following the suggestions of CLOUDCAST authors, H.
-Kavalionak developed a decentralized algorithm \cite{marriage_of_convinience} that aims to do replica management in cloud-based, peer-assisted applications.
+Kavalionak \textit{et al.}, developed a decentralized algorithm \cite{marriage_of_convinience} that aims to do replica management in cloud-based, peer-assisted applications.
 Because the proposed work has no running title, we shall refer to it as ARM.
@@
 The peers randomly organize themselves according to the CYCLON peer-sampling protocol, but depending on the results of the aggregation, the passive cloud-server used in CLOUDCAST is removed or kept in the overlay.
-In ARM, each data object to be replicated as its own \textit{replica-overlay} on top of the regular unstructured overlay, which facilitates synchronization and cloud-usage decisions.
+In ARM, each data object to be replicated as its \textit{replica-overlay} on top of the regular unstructured overlay, which facilitates synchronization and cloud-usage decisions.
 The ARM algorithm is designed to consider four replica-overlay states, separated by three size thresholds: \textit{Critical (C)}, \textit{Sufficient (S)} and \textit{Redundant (S)}.
@@
 The aggregation method mentioned earlier is the foundation \textit{monitoring phase}, which consists of a fixed number of gossip and idle rounds, allowing a peer to infer the \textit{expected size (N\textsubscript{e})} of the replica-overlay and decide if he should add or remove the cloud or peers to the replica-overlay as well as the protocol he uses, based on Algorithm \ref{alg:makedecision}.
-After each monitoring phase, each peer calculates how long it should wait before executing the \textit{monotoring phase} again.
-That time depends mostly on \textit{N\textsubscript{e}}, e.g., if the state is below the critical threshold the monitoring phase will initiate earlier than if it is above the redundant threshold, allowing the system to respond to overlay changes promptly and to keep the overlay size in a range that minimizes overhead but ensures that data does not become unavailable.
+After each monitoring phase, each peer calculates how long it should wait before executing the \textit{monitoring phase} again.
+That time depends mostly on \textit{N\textsubscript{e}}, e.g., if the state is below the critical threshold, the monitoring phase will initiate earlier than if it is above the redundant threshold, allowing the system to respond to overlay changes promptly and to keep the overlay size in a range that minimizes overhead but ensures that data does not become unavailable.
+The authors do not specify how much savings the system was capable of achieving using ARM; however, we can assume it is a substantial value; In Fig.
+\ref{fig:arm_results}, the plot on top displays the number of peers in the underlying network as a function of time, and the bottom plot shows the in-degree to the cloud server as a function of time, comparing CLOUDCAST and ARM (labeled as "our protocol").
 
-\section{Block-Level Data Replication}
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=\textwidth]{images/armresults.PNG}
+\caption{Cloud-Server in-degree, CLOUDCAST \textit{vs.} ARM \cite{marriage_of_convinience}}
+\label{fig:arm_results}
+\end{figure}
+
+\newpage\section{Block-Level Data Replication}
 \section{Probabilistic Swarm Guidance}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage\section{Solution Proposal}\label{sec:proposal}
+\section{Solution Proposal}\label{sec:proposal}
 %%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-12-15 docs/thesis-project/article.tex
Update thesis

@@
 %%%%%%%%%%%%%%%%%%%%%%
-\newpage\section{Distributed File Systems}\label{sec:relatedwork}
+\newpage\section{Distributed File Systems}\label{sec:distributed-file-systems}
 Distributed file system (DFS) enables individuals and organizations to store, access, and secure data on possibly remote locations, precisely as they would their local machines.
@@
 
-\newpage\section{P2P Overlays}
+\newpage\section{P2P Overlays}\label{sec:p2p-overlays}
 
@@
 
-\section{Membership Management Protocols}
-\subsection{CYCLON} CYCLON\cite{cyclon} is a protocol that emerges as an inexpensive solution to the maintenance of gossip-based unstructured overlays.
-CYCLON is useable even in highly dynamic environments, providing peer degree symmetry, low diameter, low clustering, resilience to churn and massive node failures, in a fully decentralized manner.
-One fundamental issue of unstructured networks is to avoid partitioning.
-One naive solution would be for each peer to keep a view of all the peers in the network and create connections has needed, but that solution is not scalable.
+\section{Membership Management Frameworks}\label{sec:membership-management}
 
-The protocol is simple.
+Gossip-based protocols, also know has epidemic protocols, gained popularity in the nineties for their usability to solve problems like database replication, failure detection, and resource-monitoring.
+In this section, we present two such protocols used for membership management, i.e., used to keep current partial views of large-scale, dynamic networks, built on unstructured overlays.
+Just like we mentioned in section \ref{sec:p2p-overlays}, keeping a view of all peers in the system is not a solution.
+Another fundamental issue of unstructured networks is to avoid partitioning.
+The following alternatives are both scalable, robust, and decentralized solutions to both these problems.
+Likewise, peers act autonomously in order to provide self-organization or self-healing capabilities to the overlays in which the protocols are applied.
+
+\subsection{CYCLON} CYCLON\cite{cyclon} is useable even in highly dynamic environments, providing peer degree symmetry, low diameter, low clustering, resilience to churn, and massive node failures.
+The focus of the authors was to provide a lightweight and simple protocol.
 It consists of having each peer keeping an ever-changing, partial view, with fixed-size \textit{c}, of the network in a structure called \textit{neighbors}.
-Periodically and independently, a contacts a random neighbor to exchange acquaintances.
+Periodically, independently and asynchronously, a peer contacts a random neighbor to exchange acquaintances.
 This process is called \textit{shuffling}.
@@
 CYCLON simulations show that the average path length and clustering coefficient, for various configurations, converged to values similar to those found in random graphs, i.e., converged to small-diameter topologies with low clustering.
-Gossip-based protocols often have high clustering, which is undesirable, since that may indicate an increased amount of redundant messages passing through the network and a higher chance of network partitioning.
+Gossip-based protocols often have high clustering, which is undesirable both in terms of robustness and overhead.
 Another exciting result of CYCLON is that the average path length increased logarithmically, and clustering decreased exponentially, as the number of peers in the network increased, hence demonstrating scalability and increased robustness.
@@
 
-We conclude by reinforcing that CYCLON is a scalable and robust solution for peer-sampling, providing good self-healing and self-organization capabilities without complex algorithms or apparent downsides other than the one where too few peers in the system may cause it to be under performant.
+We conclude by reinforcing that CYCLON provides good self-healing and self-organization capabilities without complex algorithms or apparent downsides other than the one where too few peers in the system may cause it to be under performant due to clustering effects.
 In a way, it relates to BlatAnt since the proposal passively bounds the diameter of the network.
 
-\subsection{Newscast}
-Newscast \cite{newscast} is an alternative peer-sampling protocol
+\newpage\subsection{Newscast}
+In Newscast \cite{newscast-computing}, all peers proactively exchange, timestamped, information with each other, including not only neighboring information but also app-specific information, allowing for simplified implementation of aggregation algorithms, i.e., finding statistics regarding for example network size.
+Newscast is responsible for membership-management and a second function, which does not exist in CYCLON, which is information dissemination.
+
+In Newscast, there is a separation of responsibilities.
+Within each peer, there are two types of entities.
+Those who run the newscast protocol, are called \textit{corespondents}.
+Those who run processes of a distributed application are called \textit{agents}.
+\textit{Correspondents} run the same instance of the newscast protocol.
+\textit{Agents} are not required to run the same application.
+Each \textit{agent} has one or more \textit{correspondents}.
+This separation does not explicitly exist in CYCLON.
+Each \textit{agent} implements an interface with two functions \textit{getNews()} and \textit{newsUpdate(news[])}.
+The \textit{correspondent} uses the former to request news from their \textit{agent}, and the latter to deliver app-specific news collected from other peers' \textit{correspondents} in the network.
+Each \textit{correspondent} as a fixed-size news cache of size \textit{c} and periodically exchange news with each other has followed: request news to the \textit{agent}, timestamp them along with the local time, and IP-address of the \textit{agent}, then storing them in a cache entry.
+Afterward, it selects another peer's IP address running the same \textit{correspondent} aggregation function, as available in his cache.
+Both \textit{correspondents} exchange the entirety of their caches and pass the merged cache with $2c+1$ entries to their respective \textit{agents} before discarding the oldest entries, after timestamp normalization, down to \textit{c} and storing the results themselves.
+The timestamp normalization is not a clock-synchronization process and has some inaccuracy, but as long as transmission times between two peers are not too big, there should be no problem.
+During the discarding process, the algorithm ensures that there is at most one news item per \textit{agent}.
+
+It is debatable if Newscast has an advantage over CYCLON since news exchange messages already include peer addresses, leading to a transparent resolution of the membership problem.
+CYCLON, however, has several advantages over Newscast.
+These include: the resulting overlays of CYCLON have better randomness due to its shuffling-based algorithm, whereas Newscast and other gossip-based management frameworks, seem to have small-world properties according to previous research \cite{eval-gossip-based}.
+Like in CYCLON, disconnected peers are eventually forgotten due to the timestamp property of the newscasting process.
+Newly joined peers' cache initialization is based on a regular news exchange with one or more peers who are already members of the overlay.
+In contrast, CYCLON uses \textit{c} random-walks exchanging exactly one cache entry with possibly different peers, which in turn helps to maintain the randomness properties; finally, while Newscast managed to achieve their goals of providing a robust, scalable, adaptive and lightweight solution for simultaneous membership management and aggregation CYCLON exhibits less network overhead\cite{cyclon}.
 

### 2019-12-16 docs/thesis-project/article.tex
Thesis updates

@@
 The focus of this thesis is to create a structured P2P overlay to be used on a distributed backup system, where clients upload their files to remote peers, who are selling their storage space, without persistence guarantees.
-Probabilistic Swarm Guidance (PSG) and Markov Chain theories \cite{markovchain_approach_to_pga} will be used, along with block-based file replication using erasure code algorithms and a reputation system, to hopefully create a network that is able to keep a file available at any given time, regardless of presence of attackers, network partitions and amount of churn, i.e., no matter how many peers leave/join an overlay.
+Probabilistic Swarm Guidance (PSG) and Markov Chain theories will be used, along with block-based file replication using erasure code algorithms and a reputation system, to hopefully create a network that is able to keep a file available at any given time, regardless of presence of attackers, network partitions and amount of churn, i.e., no matter how many peers leave/join an overlay.
 
@@
 \newpage\section{Block-Level Data Replication}
-\section{Probabilistic Swarm Guidance}
+\newpage\section{Probabilistic Swarm Guidance}
+Markov chains describe a sequence of possible events in which the probability of each event depends only on the previous event states, often used to model real-world processes, such as queues of customers arriving at an airport and population growths.
+Markov processes are the basis for stochastic simulation methods known as Markov Chain Monte Carlo (MCMC), which allow one to perform random sampling over vast state spaces, i.e., computing expectations concerning complex, high dimensional probability distributions.
+The idea behind MCMC is creating a Markov Chain, which will, over time, tend to an equilibrium that is close to desired density distribution.
+In robotics and control areas Markov Chains and MCMC are used as the basis foundation for a process known as probabilistic swarm guidance, which gifts autonomous robots with the capability of generating their trajectories independently of each other, in decentralized fashion and without ahead-of-time position allocation, so that sets of robots, known as swarm, converges to the desired distribution shape.
+Swarm guidance has the advantage over other methods for this problem because robots do not necessarily need to communicate with each other to perform their tasks, even if they could gain from it; their overall distribution can be adapted dynamically to the environment, and any damage to their desired formation is eventually self-repaired.
+Demonstration of these properties can be found in recent research by B.
+Açikmeçe \textit{et al.} \cite{psg-mca, psg-caa, psg-onoff}, which inspired us to apply swarm guidance principles to P2P networks and file survivability.
+Boyd et al.
+\cite{fast_mixing_mc} argue that the most popular MCMC methods, Metropolis-Hastings and Maximum-Degree Chain, that allow the creation of transition matrices, that can then be followed independently by autonomous robots in order to practice swarm guidance behavior are not optimal with respect to mixing rates, i.e.,  these algorithms may not result in matrices that converge in small amount of steps towards the desired density distribution.
+They propose that the problem of obtaining such matrices can be formulated as a convex optimization problem and prove that, for most cases, the resulting matrices have higher mixing rates.
+It is reasonable to assume that, concerning our problem in P2P networks, this might open the possibility of imposing additional constraints to the creation of the transition matrix, e.g., bounding the probability of a peer sending a file to some other peer in his hive to increase bandwidth balancing.
+
 %%%%%%%%%%%%%%%%%%%%

### 2019-12-26 docs/thesis-project/article.tex
updt

@@
 \begin{abstract}
-Peer-to-Peer (P2P) systems have emerged as a potential choice to build large scale distributed storage systems, at a fraction of the cost of the alternative cloud approaches, which have a better quality of service guarantees.
+Peer-to-Peer (P2P) systems have emerged as a potential choice to build large scale distributed storage systems, at a fraction of the cost of the alternative cloud approaches, which have better quality of service guarantees.
 We survey general-purpose P2P protocols, distributed file systems, cloud-assisted P2P protocols, and erasure-code algorithms.
@@
 Cloud-based systems are centralized architectures, in which a large number of computers are clustered and managed by master entities, which possibly, become bottlenecks.
-Both paradigms can be used to deploy distributed file-storage or file-backup applications; despite the fact that P2P implementations are cheaper for both companies and their clients, these have a hard-time achieving 99.9\% availability seen Cloud-based implementations, furthermore, depending on the P2P application there may also exist the risk of permanent file loss, making them somewhat unappealing for clients who want to store or backup their files remotely.
+Both paradigms can be used to deploy distributed file-storage or file-backup applications; despite the fact that P2P implementations are cheaper for both companies and their clients, these have a hard-time achieving 99.9\% availability seen in Cloud-based implementations, furthermore, depending on the P2P application there may also exist the risk of permanent file loss, making them somewhat unappealing for clients who want to store or backup their files remotely.
 
@@
 To be dependable is to be ready to provide the correct service (\textbf{\textit{availability}}), in a continuous manner (\textbf{\textit{reliability}}), with absence of catastrophic consequences to clients (\textbf{\textit{safety}}), e.g., losing a file, with no improper system alterations (\textbf{\textit{integrity}}) and, finally, to be able to undergo modifications and repairs at any time (\textbf{\textit{maintainability}}).
-Features that contribute to the above characteristics are summarized below.
+Some of the features that contribute to the above characteristics are summarized below.
 
@@
 It helps to avoid bottlenecks, increasing operational performance, and helps to prevent catastrophic failures.
-Prevention of failures is also achieved through \textbf{\textit{fault tolerance}}, which enables a system to continue operating, possibly at a reduced level,  rather  than failing completely,  when some part of it fails.
+Prevention of failures is also achieved through \textbf{\textit{fault tolerance}}, which enables a system to continue operating, possibly at a reduced level,  rather  than failing completely, when some part of it fails, to this end, some sort of \textbf{\textit{machine redundancy}} and \textbf{\textit{data redundancy}} may be important.
 As the amount of data passing in the network, being processed or stored in the system grows, or the number of clients using it increases, the capacity of a system to maintain its service without, or with little degradation, is called \textbf{\textit{scalability}}, it depends on the implementation of the system, and as a rule of thumb, centralized solutions are less scalable.
@@
 
-\newpage\section{Block-Level Data Replication}
-\newpage\section{Probabilistic Swarm Guidance}
+\newpage\section{Data Redundancy and File Survivability}\label{sec:dataredundancy}
+In distributed storage systems, not losing files is of the utmost importance.
+Particularly, since peer availability in P2P networks is ever-changing and not guaranteed, the availability and even durability of a file may be endangered.
+A \textbf{\textit{durable file}} is one that, once put into the system, is not lost due to failures, whereas availability means that a user can access it promptly.
+For clarity, an available file is durable, but the other way is not necessarily correct.
+Distributed systems as a whole, usually ensure durability by employing file-level or block-level \textbf{\textit{replication}}; alternatively, some systems adopt \textbf{\textit{erasure coding}} techniques (\textit{coding} for short) or a mixture of both approaches.
+Replication has the advantage of having a more straightforward and less resource-intensive recovery model, on the other hand, coding has a better \textit{expansion factor}, i.e., can provide the same or even increased durability at a lower storage overhead.
+Examples already mentioned in this thesis proposal using replication include GFS and Hadoop, but there are many others; examples of modern systems using coding redundancy include Microsoft Azure and Facebook Analytics Hadoop Cluster \cite{lrc}.
+Because our proposal focuses on the survivability of files in P2P networks, using swarm guidance, where resources may be limited, we care only about coding in the survey that follows, as it allows more files to be kept by the system.
+We will briefly mention some of the raised concerns and inherent disadvantages associated with this technique and some of the suggested approaches to mitigate them over the two decades.
+
+Two parameters specify some of the most widely used coding algorithms, the number of data symbols \textit{k} to be encoded, and the number of coded symbols \textit{n} to be produced, both assumed to be in a finite field $GF(2^w)$ and these are referred as \textit{(n, k) erasure-codes}.
+The fundamental idea, is that a file of size $S$ is divided into $k$ equally sized fragments and encode them into $n$ encoded fragments, allowing the original file to recovered from any set of encoded-fragment with count $k$.
+Reed-Solomon codes are one such example.
+Systems that use these and other more sophisticated coding are evaluated under some of the following metrics, the \textbf{\textit{repair-bandwidth}}, i.e., the number of bits communicated in the network during normal and repair operations, the \textbf{\textit{disk I/O}} which is the number of bits read from disk during each repair, the \textbf{\textit{repair locality}} which is the number of peers that would be involved in the repair process; Since coding algorithms are slow processes CPU cycles, XOR operation counts and cache misses are metrics often researched as well \cite{fast_coding}, when the main interest is tied with load-balancing and cost savings.
+We take particular interest in the first metric because, in general, coding leads to bursts in bandwidth consumption.
+Whenever a peer fails, all files he held need to be quickly recovered by some peer who needs to gather enough fragments from possibly many different peers.
+Such burst in consumption may lead to delays or even failure of the repair process due to physical link capacity between peers, who are involved in the repair process, or even throughput constraints enforced by each peer.
+
+\newpage\section{Probabilistic Swarm Guidance}\label{sec:psg}
 Markov chains describe a sequence of possible events in which the probability of each event depends only on the previous event states, often used to model real-world processes, such as queues of customers arriving at an airport and population growths.
@@
 They propose that the problem of obtaining such matrices can be formulated as a convex optimization problem and prove that, for most cases, the resulting matrices have higher mixing rates.
-It is reasonable to assume that, concerning our problem in P2P networks, this might open the possibility of imposing additional constraints to the creation of the transition matrix, e.g., bounding the probability of a peer sending a file to some other peer in his hive to increase bandwidth balancing.
+It is reasonable to assume that, concerning our problem in P2P networks, this might open the possibility of imposing additional constraints on the creation of the transition matrix, e.g., bounding the probability of a peer sending a file to some other peer in his hive to increase bandwidth balancing.
 

### 2019-12-26 docs/thesis-project/article.tex
updt

@@
 The fundamental idea, is that a file of size $S$ is divided into $k$ equally sized fragments and encode them into $n$ encoded fragments, allowing the original file to recovered from any set of encoded-fragment with count $k$.
-Reed-Solomon codes are one such example.
 Systems that use these and other more sophisticated coding are evaluated under some of the following metrics, the \textbf{\textit{repair-bandwidth}}, i.e., the number of bits communicated in the network during normal and repair operations, the \textbf{\textit{disk I/O}} which is the number of bits read from disk during each repair, the \textbf{\textit{repair locality}} which is the number of peers that would be involved in the repair process; Since coding algorithms are slow processes CPU cycles, XOR operation counts and cache misses are metrics often researched as well \cite{fast_coding}, when the main interest is tied with load-balancing and cost savings.
-We take particular interest in the first metric because, in general, coding leads to bursts in bandwidth consumption.
+We take particular interest in the first metric because, in general, coding leads may lead to unreasonable bursts in bandwidth consumption when availability requirements, in very dynamic systems, are high \cite{coding-problems}, even for today's standards.
 Whenever a peer fails, all files he held need to be quickly recovered by some peer who needs to gather enough fragments from possibly many different peers.
 Such burst in consumption may lead to delays or even failure of the repair process due to physical link capacity between peers, who are involved in the repair process, or even throughput constraints enforced by each peer.
+When it comes to redundancy in a distributed storage systems, one fundamental question is deciding when to replicate or recover data; this is often called the replica-control problem.
+
+There are two main approaches, reactive and proactive.
+Reactive approaches are the least complex, and consist of recovering data whenever a failure is detected; they suffer the most from bandwidth bursts, and mitigations to this problem include in using prediction models and thresholds to avoid unnecessary recoveries \cite{lifetime-reactive, efficient-replica-mng-reactive}.
+Proactive approaches seek to smooth bandwidth consumption by injecting new redundancy data into the network overtime at a fixed-rate \cite{proactive-rep, reliability-without-availability}.
+When it comes to coding techniques, the most widely known algorithms belong to the \textit{Maximum Distance Separable} (MDS) class, this class is likely to offer the best reliability-redundancy trade-off while consuming the least amount of storage space at the cost of increased repair-bandwidth, disk I/O and CPU time.
+Regardless of the chosen approach, research \cite{network-coding-for-dss} suggests that for coding to be competitive in distributed storage systems, one should seek to minimize these costs, while maximizing redundancy to only the amount necessary to ensure a specific service guarantee.
+They identify \textit{Minimum Bandwidth Regenerating} (MBS), and all codes in between these two ends of the spectrum as \textit{regenerating codes}, the differences between codes on the spectrum depend on the values such has \textit{n}, \textit{k} and on the implementation of the algorithms themselves.
+Shokrollahi \textit{et al.} introduces raptor codes \cite{raptor-codes}, which are a linear-time coding and decoding extension of LT codes \cite{lt-codes}.
+LT codes belong to the class of rateless codes, i.e., a potentially infinite sequence of encoded fragments can be generated from a given data source.
+Furthermore, the original data is recoverable from any subset of the encoded fragments of size equal to or slightly larger than original data size; These rateless code implementations are of particular importance because they are scalable with regards to the machine and network resources,  and fault-tolerant concerning dynamic and heterogeneous network environments.
+Finally, Zhou \textit{et al.} \cite{fast_coding} perform a study on state of the art techniques that improve coding efficiency through bitmatrix optimization, vectorization, and scheduling and prove that when used in conjunction much better levels of performance can be delivered, e.g., 552.27\% when compared to simple MDS implementations, in particular, the Reed-Solomon algorithm.
 

### 2019-12-27 docs/thesis-project/article.tex
Update article.tex

@@
 Replication has the advantage of having a more straightforward and less resource-intensive recovery model, on the other hand, coding has a better \textit{expansion factor}, i.e., can provide the same or even increased durability at a lower storage overhead.
-Examples already mentioned in this thesis proposal using replication include GFS and Hadoop, but there are many others; examples of modern systems using coding redundancy include Microsoft Azure and Facebook Analytics Hadoop Cluster \cite{lrc}.
-Because our proposal focuses on the survivability of files in P2P networks, using swarm guidance, where resources may be limited, we care only about coding in the survey that follows, as it allows more files to be kept by the system.
-We will briefly mention some of the raised concerns and inherent disadvantages associated with this technique and some of the suggested approaches to mitigate them over the two decades.
+Because our proposal focuses on the survivability of files in P2P networks, using swarm guidance, where resources may be limited, we care only about coding in the survey that follows, briefly mentioning some of the raised concerns and inherent disadvantages of this technique.
 
@@
 Systems that use these and other more sophisticated coding are evaluated under some of the following metrics, the \textbf{\textit{repair-bandwidth}}, i.e., the number of bits communicated in the network during normal and repair operations, the \textbf{\textit{disk I/O}} which is the number of bits read from disk during each repair, the \textbf{\textit{repair locality}} which is the number of peers that would be involved in the repair process; Since coding algorithms are slow processes CPU cycles, XOR operation counts and cache misses are metrics often researched as well \cite{fast_coding}, when the main interest is tied with load-balancing and cost savings.
-We take particular interest in the first metric because, in general, coding leads may lead to unreasonable bursts in bandwidth consumption when availability requirements, in very dynamic systems, are high \cite{coding-problems}, even for today's standards.
+We take particular interest in the first metric because, in general, coding may lead to unreasonable bursts in bandwidth consumption when availability requirements, in very dynamic systems, are high \cite{coding-problems}, even for today's standards.
 Whenever a peer fails, all files he held need to be quickly recovered by some peer who needs to gather enough fragments from possibly many different peers.
@@
 There are two main approaches, reactive and proactive.
-Reactive approaches are the least complex, and consist of recovering data whenever a failure is detected; they suffer the most from bandwidth bursts, and mitigations to this problem include in using prediction models and thresholds to avoid unnecessary recoveries \cite{lifetime-reactive, efficient-replica-mng-reactive}.
+Reactive approaches are the least complex, and consist of recovering data whenever a failure is detected; they suffer the most from bandwidth bursts, and mitigations to this problem include use of predictive models and thresholds to avoid unnecessary recoveries \cite{lifetime-reactive, efficient-replica-mng-reactive}.
 Proactive approaches seek to smooth bandwidth consumption by injecting new redundancy data into the network overtime at a fixed-rate \cite{proactive-rep, reliability-without-availability}.
-When it comes to coding techniques, the most widely known algorithms belong to the \textit{Maximum Distance Separable} (MDS) class, this class is likely to offer the best reliability-redundancy trade-off while consuming the least amount of storage space at the cost of increased repair-bandwidth, disk I/O and CPU time.
+
+The most widely known coding algorithms belong to the \textit{Maximum Distance Separable} (MDS) class, this class is likely to offer the best reliability-redundancy trade-off.
 Regardless of the chosen approach, research \cite{network-coding-for-dss} suggests that for coding to be competitive in distributed storage systems, one should seek to minimize these costs, while maximizing redundancy to only the amount necessary to ensure a specific service guarantee.

### 2019-12-28 docs/thesis-project/article.tex
Update article.tex

@@
 
-\section{Cloud-Assisted P2P Networks}
+\section{Cloud-Assisted P2P Networks}\label{sec:ca-networks}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Solution Proposal}\label{sec:proposal}
+\newpage\section{Solution Proposal}\label{sec:proposal}
+In this section, we introduce a prototype architecture that targets the use case that inspired to explore the theme of this thesis. We explain it by decomposition.
+
+\subsection{Use Case}
+Fundamentally, the idea is that a company, Hives, with dedicated web servers called Hiveminds, provides a platform for registered users to sell storage belonging to their personal computers directly to other registered users.
+Clients pay a monthly fee directly to Hives based on how much storage space they need.
+Hiveminds then put together a set of sellers, which constitute a Hive, that serves adequate storage space to the clients.
+Sellers are selected according to the reputation they earn over time based on their reliability, with a bias towards more reputable sellers.
+The collected fee is distributed proportionately among Hive members according to their overall space contribution and up time, calculated weekly, for example.
+The company's revenue comes from keeping a small percentage of the monthly fees.
+If the client requires continuous availability, rather than durability, he may pay a premium, in which case, inspired on the related work mentioned in section \ref{sec:ca-networks}, cloud storage may be used to help the Hive in meeting the required service guarantees.
+Hive revenue shares are reduced if their reliability degrades too much.
+
+\subsection{Key notes and Assumptions}
+\begin{itemize}
+    \item Session handling and payment models are out of the scope of this proposal.
+    \item Reputation is static, i.e., they are simulation parameters.\footnote{See goals in Section \ref{sec:intro}}
+    \item Message integrity and confidentiality are not at risk.
+    \item Sellers are not malicious, e.g., do not perform DoS on other sellers.
+    \item Sellers may fail without warning, and some absences may be transient.
+    \item Cloud service is abstract, i.e., ranges from local to a cloud storage server.
+    \item The storage space purpose is backups, thus write operations are rare.\footnote{Write operations possibly cause old versions to be overwritten}
+    \item The storage space is not shared, i.e., there are no concurrent writes.
+\end{itemize}
+
+\subsection{Hiveminds - Master Servers}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Evaluation Methodology}\label{sec:methodology}
+\newpage\section{Evaluation Methodology}\label{sec:methodology}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-12-29 docs/thesis-project/article.tex
Update article.tex

@@
 \newpage\section{Solution Proposal}\label{sec:proposal}
-In this section, we introduce a prototype architecture that targets the use case that inspired to explore the theme of this thesis. We explain it by decomposition.
+In this section, we introduce a prototype architecture that targets the use case that inspired to explore the theme of this thesis.
+We explain it by decomposition.
+See Annex section for UML, UML Sequence and other diagrams regarding the following subsections.
 
@@
 \subsection{Hiveminds - Master Servers}
+The Hivemind servers are the backbone of the system; they have to handle registration, login, subscription and payment procedures.
+When a client wishes to subscribe to a monthly plan, he contacts one Hivemind server at random by accessing his desktop application, and the Hivemind then sets a \textit{usedSpace} variable indicating how much data the Client uploaded into the system to zero.
+This variable is updated according to the Client's future requests.
+In Hives File System (Hives), users' files are uniquely identified and have no hierarchy.
+Hiveminds keep a mapping between the \textit{fileId} and the Hive\footnote{Set of peers S responsible for holding one or more files from one or more clients} that is responsible for storing them.
+Hive membership is dynamic, and a mapping between the \textit{HiveId} and the Hive members (Workers) is also kept.
+Note that it is not relevant to know which fragments each Worker in the Hive has; this is important because it diminishes the number of messages sent to Hiveminds and also reduces space overhead required to keep track of file locations.
+When a Hivemind receives a file upload request from a Client, it finds or creates a suitable Hive to host the data and calculates, if needed, the transition matrix used for PSG within it, before returning all relevant data to the Client, including Workers' addresses.
+When the Hivemind receives a delete request, it drops all mapping data regarding the respective \textit{fileId}.
 
+\subsection{Clients}
+After registration, logging in, and subscription procedures, Clients, can perform four main procedures, which we name and implement using HTTP semantics.
+In all cases, the Client first retrieves the Hive Workers' addresses.
+In a GET request, the Client downloads enough fragments from the Hive Workers and reconstructs the file himself.
+In a DELETE request, the Client first asks the Hivemind to stop tracking efforts, and once he receives an acknowledge, he broadcasts in a best-effort manner a DELETE message to any online worker within that file's Hive.
+In a POST operation, the Client first obtains permission to upload from the Hivemind; after that, he divides his file into parts, applies an erasure coding algorithm over each them, and uploads the encoded fragments to the workers at random.
+PUT updates existing files and operate similarly to POST, but first, try to do the upload to the existing Hive and increment the \textit{versionId}.
 %%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-12-29 docs/thesis-project/article.tex
Update article.tex

@@
     \item Sellers may fail without warning, and some absences may be transient.
-    \item Cloud service is abstract, i.e., ranges from local to a cloud storage server.
+    \item Cloud service is abstract, i.e., ranges from local to cloud storage servers.
     \item The storage space purpose is backups, thus write operations are rare.\footnote{Write operations possibly cause old versions to be overwritten}
@@
 In a POST operation, the Client first obtains permission to upload from the Hivemind; after that, he divides his file into parts, applies an erasure coding algorithm over each them, and uploads the encoded fragments to the workers at random.
-PUT updates existing files and operate similarly to POST, but first, try to do the upload to the existing Hive and increment the \textit{versionId}.
+PUT updates existing files and operate similarly to POST, but whenever possible, try to send the new version to the Hive that contains the previous \textit{versionId}.
+
+\subsection{Workers - Sellers}
+Apart from adequately dealing with clients' requests, Workers organize themselves in a P2P fashion within each Hive.
+One Worker can be in zero or more Hives, and within each Hive, they continuously redistribute encoded file part fragments according to the transition matrix that is calculated by the Hivemind and delivered to them by a Client.
+In the long run, we expect that a Hive achieves the desired fragment distribution that fairly distributes load while ensuring optimal durability.
+Constantly exchanging fragments is expected to cause a bandwidth overhead but has the advantage that Clients can upload their fragments to any online peer or peers in the Hive instead of one single entity that may go offline during the upload.
+Workers will then redistribute the fragment using simple routing rules.
+This approach does not sacrifice fast lookup operations due to the structured nature of the system using a Hivemind that behaves somewhat like a BitTorrent tracker.
+The never-ending fragments exchange between peers also facilitates the timely detection of non-transient faults in the Hive, which helps to ensure file durability.
+Workers piggyback aggregate information to the fragments, including but not limited to, how many parts of a file they have, allowing us to implement a hybrid between a reactive and a proactive system concerning replica control, in hopes of minimizing bandwidth burst on top of already high bandwidth requirements.
+Fragment exchange may eventually be replaced with ping-like messages to minimize bandwidth cost further, hence the importance of having a fast mixing rate transition matrix.
+When a Worker suspects that some other Worker on a Hive is faulty for too long, he first contacts a dedicated server, called Keeper, who decides the new Hive membership.
+The suspecting Worker then individually decides based on previous aggregate information if he should inject new fragments regarding the file into the network if he believes durability is at risk.
+Whenever a Worker who has been temporarily offline reconnects, he first contacts the master to know what \textit{fileId} is he is no longer responsible for maintaining, allowing him to free space.
+They also perform this procedure whenever they do not receive a fragment or a ping regarding a \textit{fileId} for a sufficiently long period.
+
+\subsection{Keepers}
+Keepers' sole responsibilities are: performing limited hop random walks on the network to gather data that might be used on a Reputation System and to decide new Hive membership on behalf of the Workers who detect faults and possibly on behalf of the Hiveminds.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-12-31 docs/thesis-project/article.tex
updt

@@
 On the one hand, we have Peer-to-Peer (P2P) networking, which can be defined as a group of equally privileged peers, who contribute with a portion of their resources, to achieve common goals.
-These networks are popular among file sharing and streaming applications \cite{ssaroiu:msp2pfss}, due to their self-organized behavior, lack of centralization and, low-cost.
+These networks are popular among file sharing and streaming applications, due to their self-organized behavior, lack of centralization and, low-cost.
 On the other hand, Cloud platforms offer unmatched, on-demand, self-service, availability, and reliability, at a higher cost, and are currently trendy, with companies such as Amazon, Google, and Microsoft offering various ITaaS products to individuals and organizations alike.
@@
 \subsection
-{Frequent approach}There has been much research on the topic of P2P overlays.
+{Frequent approach} There has been much research on the topic of P2P overlays.
 There are two significant categories of overlays, Structured and Unstructured.
 In the former, the overlay has rigid rules regarding peer placement; they favor lookup speed but have higher maintenance overhead under churn, i.e., the process of peers leaving or joining the overlay.
-Some popular P2P Structured overlays include Chord \cite{chord}, Tapestry \cite{tapestry}, Kademlia \cite{kademlia}, and P-Grid \cite{pgrid}.
+Some popular P2P Structured overlays include Chord, Tapestry, Kademlia, and P-Grid.
 In the latter, peer placement is random, which provides the overlay network with a higher degree of resilience and robustness in the advent of failures.
 Unstructured overlays disseminate information in a gossip-like fashion; thus, lookup operations are slower, resource management is less efficient, but overlay maintenance is often straightforward.
-Popular  Unstructured P2P overlays include FreeNet \cite{freenet}, Gnutella \cite{gnutella-rfc} and BitTorrent \cite{bittorrent}.
+Popular  Unstructured P2P overlays include FreeNet, Gnutella and BitTorrent.
 Research on multi-layer overlays has not yielded satisfactory results, and implementation is difficult \cite{sotart}, but research on bio-inspired overlays did have promising results, some of these include Self-Chord \cite{selfchord}.
-Finally, there have also been some attempts of transparently integrating Cloud Systems into P2P networks, e.g., CLOUDCAST \cite{cloudcast}, and vice-versa, e.g., Spotify \cite{spotify} and Wuala \cite{wuala}.
-From all the mentioned research, perhaps the one that borrows most similarity with our proposal is CLOUDCAST; however, neither it nor any of the remaining literature we present here, try to create a structured overlay composed of completely autonomous peers, who work together to keep the file at all points in time using PSG.
+Finally, there have also been some attempts of transparently integrating Cloud Systems into P2P networks, e.g., CLOUDCAST \cite{cloudcast}, and vice-versa, e.g., Spotify and Wuala.
+To the best of our knowledge, there is no P2P, nor Cloud system using PSG as a way of routing data across the network nor as a way of ensuring file survivability.
 
@@
         \SubItem{Hive is resilient to massive node failures and resistant to churn;}
-    \item Evaluating the system using PeerSim \cite{peersim} or equivalent platform;
     \item Extra miles may include:
@@
 In particular, GFS aims to provide high data throughput, low latency, and reliability in the face of individual server failures.
-Because GFS runs over commodity hardware, the proposed model considers machine and network faults as being the norm rather than the exception \cite{gfs}.
+Because GFS runs over commodity hardware, the proposed model considers machine and network faults as being the norm rather than the exception.
 This assumption is typical in modern distributed systems of any kind.
 GFS is the basis for an open-source project, which has seen continuous improvements over the years, called Hadoop Distributed File System (HDFS) \cite{hadoop}; we will not discuss it here because it holds many similarities to GFS with key differences being the approach to security management.
-One advantage of GFS over other architectures is that it leverages a distributed file system that tailors to the designers observation that, once written, files on google provided services, are only read, often sequentially, with random writes being practically non-existent, shifting optimization focus from latency to throughput, from random writes to appends when it comes to atomicity and no caching of data blocks \cite{gfs}.
+One advantage of GFS over other architectures is that it leverages a distributed file system that tailors to the designers observation that, once written, files on google provided services, are only read, often sequentially, with random writes being practically non-existent, shifting optimization focus from latency to throughput, from random writes to appends when it comes to atomicity and no caching of data blocks.
 In the GFS architecture cluster consists of a single master and multiple chunk-servers.
@@
 GFS has proven its qualities in supporting large-scale data processing.
-However, small files have shown \cite{gfs} to lead to overloads on chunk-servers holding them because chunks have a fixed large value size of 64MB.
+However, small files have shown to lead to overloads on chunk-servers holding them because chunks have a fixed large value size of 64MB.
 The reasoning for GFS to use such substantial value is to reduce network overload, reduce the number of accesses to the Master, and reduce memory used at the Master due to mappings.
@@
 \includegraphics[width=\textwidth]{images/gfs_paper.png}
-\caption{GFS Architecture \cite{gfs}}
+\caption{GFS Architecture}
 \label{fig:gfs_arch}
@@
 A small amount of \textit{Monitors} maintain a master copy of the storage cluster map with its current state.
-Monitors require high consistency and use Paxos \cite{paxos} to decide on the state of the cluster.
+Monitors require high consistency and use Paxos to decide on the state of the cluster.
 \textit{Managers} maintains detailed information about placement groups, process metadata and host metadata.
@@
 When one OSD in a placement group suspects another as died or joined the cluster, it consults a \textit{Monitor} to obtain an up to date cluster map.
-Based on the cluster map and the \textit{CRUSH}\cite{crush} algorithm, it is possible that this OSD will have to offload some of its content to other OSD.
+Based on the cluster map and the \textit{CRUSH} algorithm, it is possible that this OSD will have to offload some of its content to other OSD.
 Clients also use \textit{CRUSH} to perform reads and writes to the system.
@@
 Ceph OSDs perform a lot of self monitoring and management operations automatically and in a decentralized fashion, while being highly dependable.
-\cite{ceph, ceph_benchmarks}
+\cite{ceph_benchmarks}
 
@@
 \includegraphics[width=0.8\textwidth]{images/cephfs.png}
-\caption{CFS Architecture \cite{configure-ceph}}
+\caption{CFS Architecture}
 \label{fig:cfs_arch}
@@
 In general, we want to minimize the existence of hubs.
-In P2P networks, hubs may provide the overlay with resistance against random damage and possibly speed-up lookup operations in unstructured overlays, using biased searching; however, targeted-attacks to hubs, or natural hub failures will have more significant impacts on the connectivity of the overlay, possibly leading to partitioning or halting of the system \cite{webdragons, controlling_the_hubs}.
+In P2P networks, hubs may provide the overlay with resistance against random damage and possibly speed-up lookup operations in unstructured overlays, using biased searching; however, targeted-attacks to hubs, or natural hub failures will have more significant impacts on the connectivity of the overlay, possibly leading to partitioning or halting of the system \cite{controlling_the_hubs}.
 Peer degrees can also allow evaluation of fair work distribution among peers of the system, e.g., in terms of processing or bandwidth.
@@
 \subsection{Structured Overlays}
-\subsubsection{Kademlia} \cite{kademlia} is structured overlay, yet it is fully decentralized, designed for key-value storage systems, which, according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology \cite{kademlia}.
+\subsubsection{Kademlia} \cite{kademlia} is structured overlay, yet it is fully decentralized, designed for key-value storage systems, which, according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology.
 Kademlia peer-identifiers and data-keys are built as in Chord\cite{chord}, using consistent hashing.
@@
 Self-chord foundations are that of the standard Chord, both in terms of identifiers and the way peers organize themselves on the overlay, i.e., in ring-like shapes.
-However, it has a few benefits, which are the result of having mobile agents, called ants, independently roaming the overlay network, but collaboratively re-arranging resource identifiers so that similar data items are clustered around neighboring peers with relatively proportional workloads, meaning that data item placement is not based solely on the numerical-space shared between them and possible placement peers \cite{sotart, selfchord}.
+However, it has a few benefits, which are the result of having mobile agents, called ants, independently roaming the overlay network, but collaboratively re-arranging resource identifiers so that similar data items are clustered around neighboring peers with relatively proportional workloads, meaning that data item placement is not based solely on the numerical-space shared between them and possible placement peers.
 Hence advantages of Self-Chord over the standard implementation include lack of need to assign data to well-specified peers since keys are independently defined, leading to better behavior under high churn conditions.
@@
 \newpage\subsection{Unstructured Overlays}
-\subsubsection{Gia} \cite{gia} is a decentralized overlay protocol whose purpose is to overcome the drawbacks of Gnutella\cite{gnutella-rfc, gnutella-cs}.
+\subsubsection{Gia} \cite{gia} is a decentralized overlay protocol whose purpose is to overcome the drawbacks of Gnutella.
 Gnutella is the first-ever decentralized, unstructured P2P protocol.
@@
 \subsubsection{BlatAnt} \cite{blatant} is a bio-inspired, unstructured overlay, designed for grid computing architectures where resource discovery needs to be efficient; to this end, BlatAnt focuses solely on bounded-diameter optimization in order to minimize network overhead and lookup latency, by using swarm intelligence to support flood-like discovery protocols.
-Like in Self-Chord\cite{selfchord}, ant agents are used to optimize and maintain a self-structured P2P overlay, conversely, in BlatAnt the ants do not change the location of data items from peer to peer, but instead build and optimize the overlay such that the network uses a minimal number of connections between peers and such that the diameter of the overlay is bounded.
+Like in Self-Chord, ant agents are used to optimize and maintain a self-structured P2P overlay, conversely, in BlatAnt the ants do not change the location of data items from peer to peer, but instead build and optimize the overlay such that the network uses a minimal number of connections between peers and such that the diameter of the overlay is bounded.
 The idea is that by doing so, the average number of hops, i.e., average path length, required to obtain a data item from a lookup will be smaller, thus achieving the sought goals.
-BlatAnt also uses local index caching to improve efficiency further; The authors argue that local caches,  as an alternative to semantic clustering such as the ones used in Antares\cite{antares} and Self-Chord, are better for self-organized overlays where a stable topology can not be guaranteed.
+BlatAnt also uses local index caching to improve efficiency further; The authors argue that local caches,  as an alternative to semantic clustering such as the ones used in Self-Chord, are better for self-organized overlays where a stable topology can not be guaranteed.
 The overlay management can be viewed in three major components.
@@
 
-BlatAnt proved to accomplish its goals, quickly converging to stable overlays with bounded-diameter, even under very dynamic conditions \cite{sotart}, resulting in faster lookups due to the reduction of flood messages but also due to caching.
+BlatAnt proved to accomplish its goals, quickly converging to stable overlays with bounded-diameter, even under very dynamic conditions, resulting in faster lookups due to the reduction of flood messages but also due to caching.
 Note that BlatAnt can leverage random walks such as the ones used in Gia for even less network overhead.
@@
 They identify \textit{Minimum Bandwidth Regenerating} (MBS), and all codes in between these two ends of the spectrum as \textit{regenerating codes}, the differences between codes on the spectrum depend on the values such has \textit{n}, \textit{k} and on the implementation of the algorithms themselves.
-Shokrollahi \textit{et al.} introduces raptor codes \cite{raptor-codes}, which are a linear-time coding and decoding extension of LT codes \cite{lt-codes}.
+Shokrollahi \textit{et al.} introduces raptor codes \cite{raptor-codes}, which are a linear-time coding and decoding extension of LT codes.
 LT codes belong to the class of rateless codes, i.e., a potentially infinite sequence of encoded fragments can be generated from a given data source.
@@
 Demonstration of these properties can be found in recent research by B.
-Açikmeçe \textit{et al.} \cite{psg-mca, psg-caa, psg-onoff}, which inspired us to apply swarm guidance principles to P2P networks and file survivability.
+Açikmeçe \textit{et al.} \cite{psg-mca, psg-caa}, which inspired us to apply swarm guidance principles to P2P networks and file survivability.
 Boyd et al.
@@
 \newpage\section{Solution Proposal}\label{sec:proposal}
-In this section, we introduce a prototype architecture that targets the use case that inspired to explore the theme of this thesis.
-We explain it by decomposition.
-See Annex section for UML, UML Sequence and other diagrams regarding the following subsections.
+In this section, we introduce a prototype architecture that targets the use case that inspired to explore the theme of this thesis. We explain it by decomposition.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage\section{Evaluation Methodology}\label{sec:methodology}
+\section{Evaluation Methodology}\label{sec:methodology}
+We will first evaluate the performance of different transition matrix generating algorithms by comparing their theoretical fast mixing property against the number of discrete time steps the matrices take to converge on a simple, cycle-based, locally ran, custom-built simulator.
+We will select one of these algorithms based on how fast convergence occurs, computing time vs.
+space complexity, and the possibility of allowing each peer to update their transition matrix autonomously when faced with transient neighbor faults.
+After this initial step, we will implement a reactive replica-control policy using erasure-coding at each peer and test the durability of files on our custom-built simulator, where we have perfect failure detectors.
+We are concerned with discovering a good trade-off between storage overhead and durability by using different (n, k) values without caring for any networking aspects.
+Finally, we will measure bandwidth consumption as well as the robustness of our solution under more realistic simulations by either implement our protocol on well-known P2P Simulation test-bed such has PeerfactSim.KOM\cite{peerfact}, or if time permits it, deploy proof of concept applications on the university's computer cluster for distributed testing.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%
-\section{Work Schedule}\label{sec:workschedule}
+\newpage\section{Work Schedule}\label{sec:workschedule}
 %%%%%%%%%%%%%%%%%%%%%

### 2019-12-31 docs/thesis-project/article.tex
Update article.tex

@@
 Peer-to-Peer (P2P) systems have emerged as a potential choice to build large scale distributed storage systems, at a fraction of the cost of the alternative cloud approaches, which have better quality of service guarantees.
-We survey general-purpose P2P protocols, distributed file systems, cloud-assisted P2P protocols, and erasure-code algorithms.
-We then draft a solution for the problem of creating a bio-inspired P2P overlay to be used on a cloud-assisted P2P storage and backup system using probabilistic swarm guidance.
-We anticipate that our solution will have high network overhead, in order to provide self-healing properties to the files stored in the system.
+We survey  distributed file systems, P2P overlays, cloud-assisted P2P protocols, and erasure-code algorithms.
+We then draft a solution for the problem of creating a P2P overlay to be used on a cloud-assisted P2P backup system that uses probabilistic swarm guidance and aims to ensure file durability.
+We anticipate that our solution will have high network overhead, in order to provide good self-healing properties to the files stored in the system.
 
@@
 On the one hand, we have Peer-to-Peer (P2P) networking, which can be defined as a group of equally privileged peers, who contribute with a portion of their resources, to achieve common goals.
-These networks are popular among file sharing and streaming applications, due to their self-organized behavior, lack of centralization and, low-cost.
+These networks are popular among file sharing and streaming applications due to their self-organized behavior, lack of centralization and, low-cost.
 On the other hand, Cloud platforms offer unmatched, on-demand, self-service, availability, and reliability, at a higher cost, and are currently trendy, with companies such as Amazon, Google, and Microsoft offering various ITaaS products to individuals and organizations alike.
 Cloud-based systems are centralized architectures, in which a large number of computers are clustered and managed by master entities, which possibly, become bottlenecks.
-Both paradigms can be used to deploy distributed file-storage or file-backup applications; despite the fact that P2P implementations are cheaper for both companies and their clients, these have a hard-time achieving 99.9\% availability seen in Cloud-based implementations, furthermore, depending on the P2P application there may also exist the risk of permanent file loss, making them somewhat unappealing for clients who want to store or backup their files remotely.
+Both paradigms can be used to deploy distributed file-storage or file-backup applications; although P2P implementations are cheaper for both companies and their clients, these have a hard-time achieving availability levels seen in Cloud implementations.
+P2P systems also have a higher inherent risk of permanent file loss, making them somewhat unappealing for clients.
+
+\subsection{Frequent approach}
+P2P research often focuses on optimizing the overlay topology to provide load-balancing, read/write efficiency, and partition avoidance, implicitly assuming that reliable overlays provide reliable systems, in the case of distributed storage systems, replication or erasure coding is usually written on top of the P2P protocol to ensure durability.
+The most common approach for distributed storage systems, however, is done on the cloud using highly available machines, with expensive file or block-level replication, continuously monitored by failure detection entities to ensure the replication level is always above some threshold; While effective, the latest architecture is very costly and inefficient.
+Finally, there have also been some attempts of transparently integrating Cloud Systems into P2P networks, e.g., Cloudcast \cite{cloudcast}, and vice-versa, e.g., Spotify\cite{spotify} and Wuala\cite{wuala}.
+To the best of our knowledge, there is no system using PSG as a way of simultaneously deciding data routes, failure-detection, load-balancing, and, thus, file survivability in distributed storage systems.
 
 \subsection{Our approach}
-The focus of this thesis is to create a structured P2P overlay to be used on a distributed backup system, where clients upload their files to remote peers, who are selling their storage space, without persistence guarantees.
-Probabilistic Swarm Guidance (PSG) and Markov Chain theories will be used, along with block-based file replication using erasure code algorithms and a reputation system, to hopefully create a network that is able to keep a file available at any given time, regardless of presence of attackers, network partitions and amount of churn, i.e., no matter how many peers leave/join an overlay.
-
-\subsection
-{Frequent approach} There has been much research on the topic of P2P overlays.
-There are two significant categories of overlays, Structured and Unstructured.
-In the former, the overlay has rigid rules regarding peer placement; they favor lookup speed but have higher maintenance overhead under churn, i.e., the process of peers leaving or joining the overlay.
-Some popular P2P Structured overlays include Chord, Tapestry, Kademlia, and P-Grid.
-In the latter, peer placement is random, which provides the overlay network with a higher degree of resilience and robustness in the advent of failures.
-Unstructured overlays disseminate information in a gossip-like fashion; thus, lookup operations are slower, resource management is less efficient, but overlay maintenance is often straightforward.
-Popular  Unstructured P2P overlays include FreeNet, Gnutella and BitTorrent.
-Research on multi-layer overlays has not yielded satisfactory results, and implementation is difficult \cite{sotart}, but research on bio-inspired overlays did have promising results, some of these include Self-Chord \cite{selfchord}.
-Finally, there have also been some attempts of transparently integrating Cloud Systems into P2P networks, e.g., CLOUDCAST \cite{cloudcast}, and vice-versa, e.g., Spotify and Wuala.
-To the best of our knowledge, there is no P2P, nor Cloud system using PSG as a way of routing data across the network nor as a way of ensuring file survivability.
+The focus of this thesis is to create a distributed backup system.
+Probabilistic Swarm Guidance (PSG) and Markov Chain theories will be used, along with erasure coding.
+We hope not to lose the files regardless of the amount of churn\footnote{Property of P2P systems, which regards peers joining or leaving a P2P network} or other experienced faults.
+Furthermore, to boost the availability and reliability files, cloud servers might be sporadically used to assist the P2P networks when these are facing critical conditions.
+The main difference between our approach and the frequent ones is that we use a layered solution, where a typical client-server approach facilitates lookup operations and provides system-wide but file durability, load-balancing, and failure detection is the responsibility of multiple P2P overlays specifically tailored for backing up files.
 
 \subsection{Proposal Goals}\label{subsec:intro}
-This work seeks to address the problem of file survivability in P2P networks; our primary goal is to study, design, and evaluate a P2P overlay that uses probabilistic swarm guidance as a way to maintain a file available to service clients at all times.
-Steps will include:
+This proposal seeks to address the problem of file survivability in P2P networks; our primary goal is to design and evaluate a P2P overlay that uses probabilistic swarm guidance and erasure-coding as a mean to provide file survivability.
+Our secondary goal is to integrate instances of this overlay on a broader centralized system that provides fast look-up operations and increases availability and reliability by making connections to the cloud when needed.
+Steps include:
 
 \begin{itemize}
-    \item Implement a Tracker server with light-weight responsibilities:
-        \SubItem{Act as peer discovery service for clients;}
-        \SubItem{Act as an indexer for clusters of peers (Hives);}
-        \SubItem{Track peer reputations;}
-    \item Implement an erasure-code algorithm used by Hives to replicate clients' files;
-    \item Calculate the ideal steady-state distribution of file blocks, for a Hive, when:
-        \SubItem{No history information on a peer is available;}
-        \SubItem{Peer availability, reliability and other trust factors are known;}
-    \item Hives converge to ideal block distribution in bounded-times using PSG;
-    \item Efficiently update routing as Hives' topologies change over time;
-    \item Balance the degree of each peer such that:
-        \SubItem{Message exchange overhead is minimal;}
-        \SubItem{Any changes to the probability of file loss are negligible;}
-        \SubItem{Hive is resilient to massive node failures and resistant to churn;}
-    \item Extra miles may include:
-        \SubItem{Implement a trust system managed by the Tracker}
-        \SubItem{Peers can reconstruct Hives without Tracker intervention;}
-        \SubItem{Peers trust their neighbors independently of Tracker;}
-        \SubItem{Hives can leverage dedicated Cloud-Storage if they become unreliable;}
+    \item Identify a steady-state, $Q$, providing optimal robustness and load-balancing
+    \item Identify a transition matrix, $M$, that allows peers to achieve $Q$ collectively
+    \item Optimize $M$ with respect to peer connectivity degree
+    \item Optimize updates to $M$ as overlay topology changes over time
+    \item Implement a replication model that exploits the properties of the $M$
+    \item Evaluate solution performance in a Simulation or on in real life scenario
 \end{itemize}
@@
 \newpage\section{Distributed File Systems}\label{sec:distributed-file-systems}
-Distributed file system (DFS) enables individuals and organizations to store, access, and secure data on possibly remote locations, precisely as they would their local machines.
-It is common practice to replicate data at the file-level, block-level, or both, depending on the requirements of each system.
-For example, block-level replication may allow parallel execution of applications, such as is the case of Hadoop, Google, and Spark MapReduce algorithms.
-Here we present two DFS, which follow entirely different approaches when it comes to architecture and protocols.
-We focus only on the architectural implementation, their advantages or disadvantages.
-We do not consider parallel computation because our primary objective, in this thesis proposal, is the continuous availability of data in a system, to clients.
-Before we dive into them, it is essential to understand some desired characteristics of a DFS.
+Distributed file system (DFS) enables individuals and organizations to store, access, and secure data on possibly remote locations, as they would their local machines.
+It is common practice to do file-level or block-level, but it is rare to use erasure-coding.
+The choice depends on the solution requirements and the broadness of the system.
+For example, block replication allows parallel execution of applications using MapReduce algorithms, and because of that, Google File System\cite{gfs} (GFS) uses block-level replication.
+We present two different implementations running on commodity hardware and focus on their architecture, advantages, or disadvantages.
+Before that, it is essential to understand some desired characteristics of a DFS.
 In general, distributed systems of any kind should be \textbf{\textit{dependable}}, meaning that they should be able to avoid service failures that are more frequent and more severe than is acceptable.
 To be dependable is to be ready to provide the correct service (\textbf{\textit{availability}}), in a continuous manner (\textbf{\textit{reliability}}), with absence of catastrophic consequences to clients (\textbf{\textit{safety}}), e.g., losing a file, with no improper system alterations (\textbf{\textit{integrity}}) and, finally, to be able to undergo modifications and repairs at any time (\textbf{\textit{maintainability}}).
-Some of the features that contribute to the above characteristics are summarized below.
-
-\textbf{\textit{Load-Balancing}} is the process of balancing communication, processing, and storage overhead among all nodes of the system.
+Some of the features that contribute to the above characteristics are the following: \textbf{\textit{load-Balancing}} is the process of balancing communication, processing, and storage overhead among all nodes of the system.
 It helps to avoid bottlenecks, increasing operational performance, and helps to prevent catastrophic failures.
@@
 \subsection{Google File System}
-
-Google File System (GFS) \cite{gfs} is designed to provide efficient and reliable access to data using clusters of commodity machines located in geographically spread data-centers.
+GFS wants to provide efficient and reliable access to data using clusters located in geographically spread data-centers.
 In particular, GFS aims to provide high data throughput, low latency, and reliability in the face of individual server failures.
-Because GFS runs over commodity hardware, the proposed model considers machine and network faults as being the norm rather than the exception.
-This assumption is typical in modern distributed systems of any kind.
-GFS is the basis for an open-source project, which has seen continuous improvements over the years, called Hadoop Distributed File System (HDFS) \cite{hadoop}; we will not discuss it here because it holds many similarities to GFS with key differences being the approach to security management.
-One advantage of GFS over other architectures is that it leverages a distributed file system that tailors to the designers observation that, once written, files on google provided services, are only read, often sequentially, with random writes being practically non-existent, shifting optimization focus from latency to throughput, from random writes to appends when it comes to atomicity and no caching of data blocks.
-In the GFS architecture cluster consists of a single master and multiple chunk-servers.
+Because GFS model considers machine and network faults as being the norm, as is, nowadays typical.
+GFS serves as inspiration for a massively adopted open-source project called Hadoop Distributed File System\cite{hadoop}; we will not discuss it here because it holds many similarities to GFS with key differences being the approach to security management.
+Based on the author's observation that once written, files on google services, are only read, often sequentially, with random writes being practically non-existent, GFS chooses to optimize throughput over latency, appends over random writes and deliberately skips data block caching.
+GFS clusters consists of a single master and multiple chunk-servers.
 The Master maintains a mapping from files to chunks, the current location of chunks, and is responsible for failure detection over the chunk-servers as well as chunk-migration between them.
 Clients communicate, for reads and writes, with the Master for metadata operations and directly with the chunk-servers for data-bearing operations.
-Neither the client nor the chunk-servers cache file data because most of google's client applications stream through huge files.
 A read operation involves obtaining chunk handles and locations from the Master and then retrieving the chunks from one of the many available chunk-severs.
 For writes, the client splits the file into 64MB chunks and asks the Master to assign them an identifier, and then the client uploads his chunks to the chunk-servers indicated by the Master.
-See Fig.\ref{fig:gfs_arch} for a simplified architecture diagram.
-
 GFS has proven its qualities in supporting large-scale data processing.
 However, small files have shown to lead to overloads on chunk-servers holding them because chunks have a fixed large value size of 64MB.
-The reasoning for GFS to use such substantial value is to reduce network overload, reduce the number of accesses to the Master, and reduce memory used at the Master due to mappings.
-Concerning our proposal, Hives, since convergence to a desired steady-state using Markov Chains theory requires a large number of chunks to exist in a P2P Hive, we might have to do file-level replication rather than the desired block-level when client files are too small.\newline
+The reasoning such substantial size value is to reduce network overhead, reduce the number of accesses to the Master, and reduce memory used at the Master due to mappings.
+Concerning our proposal, since convergence to a desired steady-state is likely better achieved when more parts exist on the P2P network, we might consider using file replication rather than erasure-coding when files are small.
 
-\begin{figure}[!ht]
+\begin{figure}[t]
 \centering
@@
 
-\newpage\subsection{Ceph File System}
-Ceph \cite{ceph} is an open-source, storage system that, like GFS, runs over commodity hardware.
-Ceph aims to deliver reliable, autonomous, distributed object, block, and file storage that is self-healing, self-managing, and has no single point of failure.
-The base architectural component of Ceph is the \textit{RADOS} object store.
-All content clients read and write to the system is represented as an object, irrespective of the object's original data type.
-\textit{RADOS} is a single logical entity composed of many \textit{Ceph Storage clusters}.
-Ceph architecture, Fig.
-\ref{fig:cfs_arch} is slightly more sophisticated than that of GFS, so we will summarize the components that are of interest to us.
-Inside a \textit{Ceph Storage cluster}, we find four types of logical machines, storage devices, abstracted with a layer called \textit{Object Software Daemon} (OSD), \textit{Monitors}, \textit{Managers}, and \textit{Metadata Servers}.
-\textit{Metadata Servers} are only required when a client his sharing his DFS with other clients, these servers contain information used for POSIX semantics, e.g., file metadata such as owner and last accessed timestamps.
-A small amount of \textit{Monitors} maintain a master copy of the storage cluster map with its current state.
-Monitors require high consistency and use Paxos to decide on the state of the cluster.
-\textit{Managers} maintains detailed information about placement groups, process metadata and host metadata.
-Finally, OSD store data on behalf of clients.
-Additionally, OSD utilize their nodes' resources to perform data replication, erasure coding, load-balancing, recovery, monitoring and reporting functions.
-Each Storage Cluster as anywhere from dozens to thousands of these OSDs.
+\subsection{Ceph File System}
 
-The innovation Ceph brings is that the OSD within the same \textit{placement group}, within a cluster, are organized in a P2P fashion and are capable of autonomously carrying out replication and recovery tasks.
-\textit{Placement groups} are fragments of a logical \textit{pool}, and in turn, \textit{pools} are logical, dynamic partitions that define among other things, how much and which type of replication is done to objects that belong to it.
-When one OSD in a placement group suspects another as died or joined the cluster, it consults a \textit{Monitor} to obtain an up to date cluster map.
-Based on the cluster map and the \textit{CRUSH} algorithm, it is possible that this OSD will have to offload some of its content to other OSD.
-Clients also use \textit{CRUSH} to perform reads and writes to the system.
-To perform a write operation, for example, a client connects to a \textit{pool} in the storage cluster and runs \textit{CRUSH} with a cluster map retrieved from a Monitor.
-The result of the operation gives the client information on the placement group of the data he wishes to upload as well as the OSD he should contact to complete his request.
-When the target OSD receives the request, it runs \textit{CRUSH} to find the number of replicas that it should store.
-Then, it takes the object identifier, pool name, and the cluster map and uses \textit{CRUSH}, again, to discover the secondary OSDs to where he will replicate the write.
-The operation is completed when the primary OSD, receives acknowledgments from the remaining secondary OSD and acknowledges back to the client.
-\textit{CRUSH} is a light-weight pseudo-random placement algorithm that provides deterministic object quantization without lookups to central directories.
-Because it runs independently on clients and peers, it avoids the need to have the monitor constantly updating cluster to object mappings.
-Ceph is proof that distributed file systems can reliably be implemented using P2P behavior.
-Ceph OSDs perform a lot of self monitoring and management operations automatically and in a decentralized fashion, while being highly dependable.
-\cite{ceph_benchmarks}
-
-\begin{figure}[!ht]
+\begin{figure}[t]
 \centering
-\includegraphics[width=0.8\textwidth]{images/cephfs.png}
+\includegraphics[width=0.6\textwidth]{images/cephfs.png}
 \caption{CFS Architecture}
@@
 
-\newpage\section{P2P Overlays}\label{sec:p2p-overlays}
+Ceph \cite{ceph} is an open-source system that aims to deliver reliable, autonomous, distributed object, block, and file storage that is self-healing, self-managing, and has no single point of failure.
+The base architectural component of Ceph is the RADOS object store.
+All clients' reads and writes to the system are represented as objects, irrespective of the objects' original data types.
+RADOS is a single logical entity composed of many storage clusters.
+Ceph architecture is sophisticated, so, we summarize the main components.
+Inside a storage cluster, we find four types of logical machines, storage devices, abstracted with a layer called \textit{object software daemon} (OSD), \textit{monitors}, \textit{managers}, and \textit{metadata servers}.
+Metadata servers contain information used for POSIX semantics, e.g., file owner or last accessed timestamps.
+A small number of monitors maintain a master copy of the storage cluster map with its current state.
+Monitors use Paxos to decide on the state of the cluster.
+Managers maintain detailed information about placement groups, process and host metadata.
+Finally, OSD store data on behalf of clients.
+Additionally, OSD utilize their nodes' resources to perform replication, erasure coding, load-balancing, recovery, monitoring, and reporting functions.
+Each Storage Cluster as anywhere from dozens to thousands of these OSDs.
+Ceph's innovation is that OSD within the same \textit{placement group}, within a cluster, are organized in a P2P fashion and are capable of autonomously carrying out recovery tasks.
+Placement groups are fragments of a logical \textit{pool}, and in turn, pools are logical, dynamic partitions that define among other things, how much and which type of replication is done to objects that belong to it.
+When one OSD in a placement group suspects another as left or joined the cluster, it consults a Monitor to obtain an up to date cluster map.
+After running the CRUSH algorithm\cite{crush} over the cluster map, it is possible that the OSD will offload some of its content to another OSD.
+Clients also use CRUSH to perform reads and writes to the system.
+To perform a write operation, for example, a client connects to a pool in the storage cluster and runs CRUSH with a cluster map retrieved from a Monitor.
+The result of the operation gives the client information on the placement group of the data he wishes to upload as well as the OSD he should contact to complete his request.
+When the target OSD receives the request, it takes the object identifier, pool name, and the cluster map and uses CRUSH, to discover the number of replicas that it should store and to which secondary OSDs he must replicate the write.
+The operation is completed when the primary OSD, receives acknowledgments from the remaining secondary OSD and acknowledges back to the client.
+Because operations runs independently on clients and OSDs, there is no need to have the monitor constantly updating cluster to object mappings.
+Ceph is proof that DFS can reliably be implemented using P2P behavior.
+Ceph OSDs perform a lot of self-monitoring and management operations and in a decentralized fashion while being highly dependable\cite{ceph_benchmarks}; we hope to be able to replicate these achievements in our approach.
 
+\section{P2P Overlays}\label{sec:p2p-overlays}
 In large-scale P2P systems, it is unfeasible and even undesirable for each peer in the network to know and explicitly interact with every other node connecting it to other peers.
@@
 \subsection{Structured Overlays}
-\subsubsection{Kademlia} \cite{kademlia} is structured overlay, yet it is fully decentralized, designed for key-value storage systems, which, according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology.
+Structured overlay ensures peer placement follows rigid rules in order to speed up lookup operations, commonly using Distributed Hash Tables or Tree implementations.
+However, for this reason, these overlays also have higher maintenance costs due to churn, making them less ideal for highly dynamic environments.
+
+\subsubsection{Kademlia} \cite{kademlia} is fully decentralized and designed for key-value storage systems, which, according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology.
 Kademlia peer-identifiers and data-keys are built as in Chord\cite{chord}, using consistent hashing.
@@
 
-\newpage\subsection{Unstructured Overlays}
+\subsection{Unstructured Overlays}
+Unstructured overlays usually disseminate information in a gossip-like fashion; thus, lookup operations are slower than structured overlays, resource management is less efficient, but overlay maintenance is often straightforward.
+Because peer placement is arbitrary, they tend to provide better scalability and a higher degree of resilience in the advent of failures.
+
 \subsubsection{Gia} \cite{gia} is a decentralized overlay protocol whose purpose is to overcome the drawbacks of Gnutella.
@@
 
-\subsection{CYCLON} CYCLON\cite{cyclon} is useable even in highly dynamic environments, providing peer degree symmetry, low diameter, low clustering, resilience to churn, and massive node failures.
+\subsection{Cyclon} Cyclon\cite{cyclon} is useable even in highly dynamic environments, providing peer degree symmetry, low diameter, low clustering, resilience to churn, and massive node failures.
 The focus of the authors was to provide a lightweight and simple protocol.
@@
 Note that, after \textit{i} initiates a shuffle with \textit{j}, \textit{i} becomes \textit{j}'s neighbor, but \textit{j} is no longer a neighbor to \textit{i}, i.e., the neighboring relation between \textit{i} and \textit{j} is reversed.
-Unlike many protocols, CYCLON links between peers are not bidirectional.
+Unlike many protocols, Cyclon links between peers are not bidirectional.
 
@@
 Timely removal is fundamental for the robustness of the overlay.
-CYCLON uses passive detection.
+Cyclon uses passive detection.
 Whenever a peer \textit{i} contacts \textit{j} for shuffling and obtains no response it assumes \textit{j} is disconnected and removes it from his \textit{neighbors}.
@@
 
-CYCLON simulations show that the average path length and clustering coefficient, for various configurations, converged to values similar to those found in random graphs, i.e., converged to small-diameter topologies with low clustering.
+Cyclon simulations show that the average path length and clustering coefficient, for various configurations, converged to values similar to those found in random graphs, i.e., converged to small-diameter topologies with low clustering.
 Gossip-based protocols often have high clustering, which is undesirable both in terms of robustness and overhead.
-Another exciting result of CYCLON is that the average path length increased logarithmically, and clustering decreased exponentially, as the number of peers in the network increased, hence demonstrating scalability and increased robustness.
+Another exciting result of Cyclon is that the average path length increased logarithmically, and clustering decreased exponentially, as the number of peers in the network increased, hence demonstrating scalability and increased robustness.
 
@@
 
-We conclude by reinforcing that CYCLON provides good self-healing and self-organization capabilities without complex algorithms or apparent downsides other than the one where too few peers in the system may cause it to be under performant due to clustering effects.
+We conclude by reinforcing that Cyclon provides good self-healing and self-organization capabilities without complex algorithms or apparent downsides other than the one where too few peers in the system may cause it to be under performant due to clustering effects.
 In a way, it relates to BlatAnt since the proposal passively bounds the diameter of the network.
 
-\newpage\subsection{Newscast}
+\subsection{Newscast}
 In Newscast \cite{newscast-computing}, all peers proactively exchange, timestamped, information with each other, including not only neighboring information but also app-specific information, allowing for simplified implementation of aggregation algorithms, i.e., finding statistics regarding for example network size.
-Newscast is responsible for membership-management and a second function, which does not exist in CYCLON, which is information dissemination.
+Newscast is responsible for membership-management and a second function, which does not exist in Cyclon, which is information dissemination.
 
@@
 Each \textit{agent} has one or more \textit{correspondents}.
-This separation does not explicitly exist in CYCLON.
+This separation does not explicitly exist in Cyclon.
 Each \textit{agent} implements an interface with two functions \textit{getNews()} and \textit{newsUpdate(news[])}.
@@
 
-It is debatable if Newscast has an advantage over CYCLON since news exchange messages already include peer addresses, leading to a transparent resolution of the membership problem.
-CYCLON, however, has several advantages over Newscast.
-These include: the resulting overlays of CYCLON have better randomness due to its shuffling-based algorithm, whereas Newscast and other gossip-based management frameworks, seem to have small-world properties according to previous research \cite{eval-gossip-based}.
-Like in CYCLON, disconnected peers are eventually forgotten due to the timestamp property of the newscasting process.
+It is debatable if Newscast has an advantage over Cyclon since news exchange messages already include peer addresses, leading to a transparent resolution of the membership problem.
+Cyclon, however, has several advantages over Newscast.
+These include: the resulting overlays of Cyclon have better randomness due to its shuffling-based algorithm, whereas Newscast and other gossip-based management frameworks, seem to have small-world properties according to previous research \cite{eval-gossip-based}.
+Like in Cyclon, disconnected peers are eventually forgotten due to the timestamp property of the newscasting process.
 Newly joined peers' cache initialization is based on a regular news exchange with one or more peers who are already members of the overlay.
-In contrast, CYCLON uses \textit{c} random-walks exchanging exactly one cache entry with possibly different peers, which in turn helps to maintain the randomness properties; finally, while Newscast managed to achieve their goals of providing a robust, scalable, adaptive and lightweight solution for simultaneous membership management and aggregation CYCLON exhibits less network overhead\cite{cyclon}.
+In contrast, Cyclon uses \textit{c} random-walks exchanging exactly one cache entry with possibly different peers, which in turn helps to maintain the randomness properties; finally, while Newscast managed to achieve their goals of providing a robust, scalable, adaptive and lightweight solution for simultaneous membership management and aggregation Cyclon exhibits less network overhead\cite{cyclon}.
 
@@
 
-\subsection{CLOUDCAST}
-CLOUDCAST \cite{cloudcast} is a P2P-assisted cloud architecture that differs from other proposals because it does not offload the provided service to peers only when service availability is not affected.
+\subsection{Cloudcast}
+Cloudcast \cite{cloudcast} is a P2P-assisted cloud architecture that differs from other proposals because it does not offload the provided service to peers only when service availability is not affected.
 Furthermore, the architecture does not augment P2P systems with elastic computing nodes that perform particular tasks beyond the reach of the P2P network, such has bulk-synchronous content distribution.
-CLOUDCAST functionality depends solely on a passive storage service that transparently integrates into a gossip-based P2P system.
+Cloudcast functionality depends solely on a passive storage service that transparently integrates into a gossip-based P2P system.
 The reasoning behind not using elastic computing instances is to eliminate fixed renting costs.
@@
 Two problems are identified, \textit{membership management} and \textit{information diffusion}.
-To solve the first problem, CLOUDCAST uses the CYCLON protocol to do peer-sampling.
-From other alternatives, CYCLON was chosen not for its particular inexpensiveness, but because the protocol maintains a random overlay topology that works even during high churn periods and even in the presence of catastrophic failures where up to 80\% of the peers fail.
-CYCLON properties also guarantee that references to the cloud server are not too many keeping usage cost under control.
-Finally, the cloud server can easily integrate in the protocol, in particular, CYCLON does not require the cloud server to initiate any communications, i.e., its active roles can be carried out by peers interacting with it; Finally, the cloud server can easily integrate into the protocol, in particular, CYCLON does not require the cloud server to initiate any communications, i.e., its active roles can be carried out by peers interacting with it; In this regard, the dummy process carried by the peers occasionally refreshes pointers to the cloud to avoid losing all references to it in the system.
+To solve the first problem, Cloudcast uses the Cyclon protocol to do peer-sampling.
+From other alternatives, Cyclon was chosen not for its particular inexpensiveness, but because the protocol maintains a random overlay topology that works even during high churn periods and even in the presence of catastrophic failures where up to 80\% of the peers fail.
+Cyclon properties also guarantee that references to the cloud server are not too many keeping usage cost under control.
+Finally, the cloud server can easily integrate in the protocol, in particular, Cyclon does not require the cloud server to initiate any communications, i.e., its active roles can be carried out by peers interacting with it; Finally, the cloud server can easily integrate into the protocol, in particular, Cyclon does not require the cloud server to initiate any communications, i.e., its active roles can be carried out by peers interacting with it; In this regard, the dummy process carried by the peers occasionally refreshes pointers to the cloud to avoid losing all references to it in the system.
 To deal with the case where all references cloud are lost due to the dynamic characteristics of the environment, e.g.
@@
 Peers never contact the cloud server for rumor-mongering purposes.
-The anti-entropy approach can be piggybacked to the CYCLON peer-sampling step, thus decreasing message overhead; the authors, however, do not mention this optimization.
+The anti-entropy approach can be piggybacked to the Cyclon peer-sampling step, thus decreasing message overhead; the authors, however, do not mention this optimization.
 
-\subsection{CYCLON-CLOUDCAST: Adaptive Replica Management}
+\subsection{Cyclon-Cloudcast: Adaptive Replica Management}
 
@@
 
-Following the suggestions of CLOUDCAST authors, H.
+Following the suggestions of Cloudcast authors, H.
 Kavalionak \textit{et al.}, developed a decentralized algorithm \cite{marriage_of_convinience} that aims to do replica management in cloud-based, peer-assisted applications.
@@
 Furthermore, the proposal acknowledges that replicas should be synchronized, i.e., the system provides at least some consistency guarantees.
-In simple terms, this is perhaps the main difference between the proposed work and CLOUDCAST, which stores all updates in an always-on passive cloud server, helping with information diffusion and overlay connectivity and overall reliability of the system in extreme conditions.
-CLOUDCAST monetary cost reductions are primarily associated with limiting access to the cloud server.
-ARM effectively reduces those costs even further by performing timely switches between CLOUDCAST and regular CYCLON protocols, depending on the size of the network.
+In simple terms, this is perhaps the main difference between the proposed work and Cloudcast, which stores all updates in an always-on passive cloud server, helping with information diffusion and overlay connectivity and overall reliability of the system in extreme conditions.
+Cloudcast monetary cost reductions are primarily associated with limiting access to the cloud server.
+ARM effectively reduces those costs even further by performing timely switches between Cloudcast and regular Cyclon protocols, depending on the size of the network.
 To estimate the network size without centralization or broadcasting algorithms that would ultimately lead to either bottlenecks or overhead, ARM leverages a gossip-based aggregation method \cite{gossip-based_aggregation} that allows each peer to infer the approximate size of the network autonomously, at the end of each epoch.
 Estimation accuracy is dependent only on the size of the epoch.
-Likewise, the execution cost depends on the length of the epochs and the network churn rate, but not on the number of peers in the system, which means that ARM keeps the scale-free properties of the unstructured overlays that used in CLOUDCAST.
-The peers randomly organize themselves according to the CYCLON peer-sampling protocol, but depending on the results of the aggregation, the passive cloud-server used in CLOUDCAST is removed or kept in the overlay.
+Likewise, the execution cost depends on the length of the epochs and the network churn rate, but not on the number of peers in the system, which means that ARM keeps the scale-free properties of the unstructured overlays that used in Cloudcast.
+The peers randomly organize themselves according to the Cyclon peer-sampling protocol, but depending on the results of the aggregation, the passive cloud-server used in Cloudcast is removed or kept in the overlay.
 In ARM, each data object to be replicated as its \textit{replica-overlay} on top of the regular unstructured overlay, which facilitates synchronization and cloud-usage decisions.
@@
 The authors do not specify how much savings the system was capable of achieving using ARM; however, we can assume it is a substantial value; In Fig.
-\ref{fig:arm_results}, the plot on top displays the number of peers in the underlying network as a function of time, and the bottom plot shows the in-degree to the cloud server as a function of time, comparing CLOUDCAST and ARM (labeled as "our protocol").
+\ref{fig:arm_results}, the plot on top displays the number of peers in the underlying network as a function of time, and the bottom plot shows the in-degree to the cloud server as a function of time, comparing Cloudcast and ARM (labeled as "our protocol").
 
@@
 \includegraphics[width=\textwidth]{images/armresults.PNG}
-\caption{Cloud-Server in-degree, CLOUDCAST \textit{vs.} ARM \cite{marriage_of_convinience}}
+\caption{Cloud-Server in-degree, Cloudcast \textit{vs.} ARM \cite{marriage_of_convinience}}
 \label{fig:arm_results}

### 2019-12-31 docs/thesis-project/article.tex
Update article.tex

@@
 \section{P2P Overlays}\label{sec:p2p-overlays}
-In large-scale P2P systems, it is unfeasible and even undesirable for each peer in the network to know and explicitly interact with every other node connecting it to other peers.
-In this regard, a P2P overlay is a logical network that abstracts the physical network underneath it, typically IP-based, and where, single-hop edges represent links between pairs of peers, as depicted in Fig.
-\ref{fig:network_diagrams}.
-Before we survey related work in the field of P2P overlays, we first list some essential properties and technical jargon related to overlays and graph modeling, that helps us understand better the pros and cons of the research we are presenting in the next subsections of this thesis proposal.
 
-\begin{figure}[ht]
+\begin{figure}[t]
     \centering
@@
 
-An overlay is said to be \textbf{\textit{connected}} if there is a path that allows any pair of peers \textit{i}, \textit{j} to communicate with each other, i.e., there is no partitioning and no peer is isolated in the system.
-
-The \textbf{\textit{degree}} of a peer is the total number of connections established between some peer \textit{i} and other peers.
-It is possible to divide the degree into \textit{in-degree} and \textit{out-degree}, i.e., the number of channels a peer \textit{i} has, through which it can only receive data or send data, respectively.
-When all peers in an overlay have the same degree, the overlay is said to be \textbf{\textit{regular}} or \textbf{\textit{symmetric}}, but this is not usual.
+In large-scale P2P systems, it is unfeasible and even undesirable for each peer in the network to know and explicitly interact with every other node connecting it to other peers.
+In this regard, a P2P overlay is a logical network that abstracts the physical network underneath it, typically IP-based, and where, single-hop edges represent links between pairs of peers, as depicted in Fig.
+\ref{fig:network_diagrams}.
+We define some properties and technical jargon related to overlays and graphs that help us understand the research we survey in the next subsections.
+An overlay is said to be \textbf{\textit{connected}} if there is a path that allows any pair of peers $(i, j)$ to communicate with each other, i.e., there is no partitioning and no peer is isolated in the system.
+The \textbf{\textit{degree}} of a peer is the total number of connections established between a peer \textit{i} and other peers.
+We can divide the degree into \textit{in-degree} and \textit{out-degree}, i.e., the number of channels a peer \textit{i} has, through which it can only receive data or send data, respectively.
+When all peers have the same degree, the overlay is said to be \textbf{\textit{regular}} or \textbf{\textit{symmetric}}.
 We can use the \textbf{\textit{degree distribution}} to evaluate the robustness of an overlay since through it; it is possible to detect \textit{weakly connected peers} and \textit{massively connected hubs}, i.e., peers with very high degree.
-In general, we want to minimize the existence of hubs.
-In P2P networks, hubs may provide the overlay with resistance against random damage and possibly speed-up lookup operations in unstructured overlays, using biased searching; however, targeted-attacks to hubs, or natural hub failures will have more significant impacts on the connectivity of the overlay, possibly leading to partitioning or halting of the system \cite{controlling_the_hubs}.
-Peer degrees can also allow evaluation of fair work distribution among peers of the system, e.g., in terms of processing or bandwidth.
-
-\textbf{\textit{Network diameter}} is the number of links that compose the longest of all the shortest paths between any two peers \textit{i}, \textit{j}.
-Knowing the length of all shortest paths between all possible pairs of peers \textit{i}, \textit{j} one can extract the \textbf{\textit{average path length}} and hence derive how messages, on average, will be relayed on the overlay to contact a stranger.
+Hubs may provide the overlay with resistance against random damage and possibly speed-up lookup operations in unstructured overlays that use biased searching; however, hub failures will have significant impacts on the connectivity of the overlay, severely damaging performance or even halting the system \cite{webdragons, controlling_the_hubs}.
+Peer degrees are also indicators of fair work distribution among peers.
+\textbf{\textit{Network diameter}} is the number of links that compose the longest of all the shortest paths between any two peers $(i, j)$.
+Knowing the length of all shortest paths between all possible pairs of peers allows us to extract the \textbf{\textit{average path length}} and derive how messages, on average, will be relayed on the overlay to contact a stranger.
 Still regarding distances, the path stretch is the ratio between the number of links a message has to traverse, between any two peers, in the underlying network and the overlay network.
-
 \textbf{\textit{Clustering}} directly affects connectivity and network diameter; we can measure it using the \textbf{\textit{clustering coefficient}}.
 This coefficient represents the percentage of neighbors a peer \textit{i} has, who are also neighbors among themselves.
-Like peers with high degrees, high clustering coefficients indicate the increased chance of partitioning in the system in the event of massive peer failures or high degrees of \textit{churn}.
-Furthermore, high clustering coefficients may indicate, in gossip-based protocols, bandwidth overhead due to redundant communications in the cluster, but has the advantage of making lookups faster even when the network is experiencing heavy load, e.g., flash crowd scenarios.
-
-Another two terms that often appear in P2P literature are the small-world and the power-law properties.
+Like peers with high degrees, high clustering coefficients indicate the increased chance of partitioning in the system in the event of massive peer failures and \textit{churn}.
+Furthermore, it may indicate, in gossip-based protocols, bandwidth overhead due to redundant communications in the cluster, but has the advantage of making lookups faster even when the network is experiencing heavy load, e.g., flash crowd scenarios.
+Another two terms that often appear in literature are the \textbf{\textit{small-world}} and the \textbf{\textit{power-law}} properties.
 The small-world property is an umbrella term for a system with a high clustering coefficient and low average path length, yet most peers are not direct neighbors of each other.

### 2019-12-31 docs/thesis-project/article.tex
Update article.tex

@@
 \subsection{Structured Overlays}
-Structured overlay ensures peer placement follows rigid rules in order to speed up lookup operations, commonly using Distributed Hash Tables or Tree implementations.
-However, for this reason, these overlays also have higher maintenance costs due to churn, making them less ideal for highly dynamic environments.
+Structured overlay ensures peer placement follows rigid rules in order to speed up lookup operations, commonly using Distributed Hash Tables (DHT) or Tree implementations.
+However, for this reason, these overlays also have higher maintenance costs, making them less ideal for highly dynamic environments.
 
-\subsubsection{Kademlia} \cite{kademlia} is fully decentralized and designed for key-value storage systems, which, according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology.
+\subsubsection{Kademlia} \cite{kademlia} is fully decentralized and designed for key-value storage systems, and according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology.
 Kademlia peer-identifiers and data-keys are built as in Chord\cite{chord}, using consistent hashing.
@@
 The unidirectionality allows caching to be done along lookup paths, alleviating hot spots.
-Each peer in the overlay maintains a binary tree, routing table, whose leaves are lists of peers, called k-buckets, at distances $2^{x}$ to $2^{x+1}$, $x \in\hspace{1mm}\mathclose[0,160\mathclose]$ from itself, sorted according to a least recently seen policy.
-As a peer receives messages if the sender is a new acquaintance, that sender is placed on the proper k-bucket if there is enough space in it; otherwise, it will replace the least seen peer only if that peer fails to respond to a ping.
+Each peer in the overlay maintains a binary tree, routing table, whose leaves are lists of peers, called k-buckets, at distances $2^{x}$ to $2^{x+1}$, $x \in\hspace{1mm}\mathclose[0,160\mathclose]$ from itself, sorted according to the least recently seen policy.
+As a peer receives messages, if the sender is a new acquaintance, that sender is placed on the proper k-bucket.
+If a replacement is required, then the least seen peer will be evicted, but only if that peer fails to respond to a ping.
+This policy helps in the defense against attacks where malicious entities flood the network with new dummy peers.
 By preferring old acquaintances, Kademlia seeks to enhance the stability of the overlay based on the observation that the longer a peer remains connected to the network, the higher the probability that it will remain connected at least another hour\cite{ssaroiu:msp2pfss}.
-This policy also prevents some denial of service attacks, where an enemy floods the network with new peers.
-The XOR metric gives Kademlia an advantage over other DHT-based implementations by making routing tables less rigid and by allowing peers in the network to learn useful routing information from discovery and lookup messages.
-Because of the existence of k-buckets, routing can also be done around failed nodes, providing some resistance to churn and partitioning.
-
-The standard Kademlia implementation, however, does not implement an effective mechanism to keep dated entries out of a peer's k-buckets, meaning that the longer a peer stays offline, the worse is k-buckets are.
-Dated entries removed when the peer contacts a peer and obtains no response, meaning that insert and lookup operations may take longer than they could.
-The take-away from Kademlia protocol is that both the liveness and the freshness of an overlay can be improved by passively or actively making peers monitor each other.\newline
+The XOR metric gives Kademlia an advantage over other DHT-based implementations by making routing tables less rigid and by allowing peers to learn useful routing information from messages they receive.
+K-buckets allow routing to be done around failed nodes, providing increased churn resistance.
+We note that Kademlia does not implement an effective mechanism to keep dated entries out of a peer's k-buckets, meaning that the longer a peer stays offline, the worse is k-buckets are when he returns.
+The standard implementation only removes dated entries when the peer contacts another and obtains no response; While our proposed solution differs from Kademlia's solution, but it raises awareness to the fact that our networks rely upon active peers, while not keeping new interested peers locked out of the system.
 

### 2020-01-01 docs/thesis-project/article.tex
Update article.tex

@@
 We note that Kademlia does not implement an effective mechanism to keep dated entries out of a peer's k-buckets, meaning that the longer a peer stays offline, the worse is k-buckets are when he returns.
-The standard implementation only removes dated entries when the peer contacts another and obtains no response; While our proposed solution differs from Kademlia's solution, but it raises awareness to the fact that our networks rely upon active peers, while not keeping new interested peers locked out of the system.
+The standard implementation only removes dated entries when the peer contacts another and obtains no response;
 
-\subsubsection{Self-Chord} \cite{selfchord} is a bio-inspired, structured overlay, specifically designed to be used in cloud computing or grid computing architectures.
-Self-chord foundations are that of the standard Chord, both in terms of identifiers and the way peers organize themselves on the overlay, i.e., in ring-like shapes.
-However, it has a few benefits, which are the result of having mobile agents, called ants, independently roaming the overlay network, but collaboratively re-arranging resource identifiers so that similar data items are clustered around neighboring peers with relatively proportional workloads, meaning that data item placement is not based solely on the numerical-space shared between them and possible placement peers.
-Hence advantages of Self-Chord over the standard implementation include lack of need to assign data to well-specified peers since keys are independently defined, leading to better behavior under high churn conditions.
-Less overlay maintenance overhead since when a peer joins or leaves the network, ants will autonomously and eventually, distribute data items appropriately; Lastly, since load-balancing is continuously being carried out by ants, the scalability and robustness of the system increases since individual peer failures are less likely to lead to massive file loss.
+\subsubsection{Self-Chord} \cite{selfchord} is a bio-inspired overlay designed for cloud or grid computing architectures.
+Its foundations are that of the standard Chord, both in terms of identifier generation and the way peers connect, i.e., in ring-like shapes.
+Self-Chord a few key differences, however, which result from having \textit{ant} agents independently roaming the overlay but collaboratively re-arranging resource location, i.e., data placement is not based solely on the numerical-space relation between data and placement peers.
+The rearrangement improves lookup operation speed by clustering similar items around neighboring nodes and provides some churn resistance.
+Other advantages include less overlay maintenance overhead since when a peer joins or leaves the network, ants will autonomously and eventually distribute data appropriately; While redistributing data, Ants also ensure load-balancing, thus improving scalability and robustness since individual peer failures are less likely to lead to massive file loss.
 According to the authors, the behavior of the ants ensures that even under highly dynamic, unfavorable conditions, rearrangement, and discovery of data items takes only logarithmic time.
-
-Bio-inspired systems like the one above, diminish or eliminate inherent disadvantages typically associated with structured or unstructured overlays, such as the ones mentioned in the introduction of this thesis proposal.
-In these, agents with simple behavior often accomplish complex behavior that would otherwise require time or space consuming algorithms, sometimes with high code complexity.
-In our proposal, we, too, seek to create a system that, just like Self-Chord, has self-healing and self-organizing behaviors.
+Bio-inspired systems like Self-Chord, diminish or eliminate some of the inherent disadvantages associated with structured or unstructured overlays.
+In these, simple agents often accomplish complex behavior that would otherwise require time or space consuming algorithms, often with high code complexity.
+The downside of bio-inspired overlays is that the results of agent actions are not always easy to evaluate or predict.
+In our proposal, we, too, seek to create a system that employs autonomous, self-healing, and self-organizing behaviors.
 
 \subsection{Unstructured Overlays}
-Unstructured overlays usually disseminate information in a gossip-like fashion; thus, lookup operations are slower than structured overlays, resource management is less efficient, but overlay maintenance is often straightforward.
+Unstructured overlays usually disseminate information in a gossip or broadcasting fashion; lookup operations are slower than on structured overlays, but overlay maintenance is often straightforward.
 Because peer placement is arbitrary, they tend to provide better scalability and a higher degree of resilience in the advent of failures.
 
-\subsubsection{Gia} \cite{gia} is a decentralized overlay protocol whose purpose is to overcome the drawbacks of Gnutella.
-Gnutella is the first-ever decentralized, unstructured P2P protocol.
-Gnutella's main issues are lack of scalability, as resource discovery bases itself on simple message flooding mechanisms and how easily some peers become overloaded when faced with high rates of aggregate queries, leading to system degradation.
-Gia proposes the use of super-nodes, which are now also part of the Gnutella protocol.
-The super-nodes receive and route queries to peers holding data, but these super-nodes and the construction of the topology around them is dynamic and adaptive.
-Gia recognizes the heterogeneity and resource constraints of the peers and also uses random walks to alter the topology adaptively as well as lookup operations.
-The results show that Gia is three to five orders of magnitude better than that of Gnutella.
-Random-walks by themselves, while minimizing the network overhead, are less likely to find appropriate responses to the performed lookup, unless they are biased to high-degree peers, but this can make peer overloading more frequent.
-Thus, to tackle both problems, Gia implements a topology adaptation that puts most peers within reach of high capacity peers, while simultaneously ensuring that they can handle the likely incoming requests.
-Lastly, all peers keep pointers to data items kept by their immediate one-hop neighbors.
-
-The algorithm Gia uses to ensure high-capacity-peers are also high-degree ones depends on a satisfaction function that returns a value in $\mathclose[0,1\mathclose]$.
-As long as the returned value is not one, the peer takes the initiative of trying to connect itself to another peer at random, preferring those whose capacity is better than his.
+\subsubsection{Gia} \cite{gia} seeks to overcome the drawbacks of Gnutella\cite{gnutella-rfc}, which was the first decentralized, unstructured P2P protocol.
+Gnutella's main issues were the lack of scalability, as resource discovery was made with flooding, and how easily some peers became overloaded when faced with high rates of aggregate queries.
+Gia proposes the use of super-nodes, which are nowadays also part of the Gnutella protocol.
+The super-nodes receive and route queries to peers holding data.
+Super-node selection and the construction of the topology around them is dynamic.
+Gia uses random walks, both to alter the topology as well as perform lookup operations.
+Random-walks alone are less likely to find appropriate lookup responses unless they are biased to high-degree peers, which in turn can make peer overloading frequent.
+To solve the issue, Gia implements a topology adaptation that puts most peers within reach of high capacity peers, while simultaneously ensuring that they can handle the incoming requests.
+Lastly, all peers keep pointers to data items kept by their immediate one-hop neighbors to help guiding queries in the right direction.
+The algorithm Gia uses to ensure high-capacity-peers are also high-degree ones depends on a satisfaction function autonomously executed by each peer, which returns a value $s\in\mathclose[0,1\mathclose]$.
+As long as $s<1$, the peer takes the initiative of calling another random peer, preferring those whose capacity is better than his.
 During the handshake, any of the intervenients may abort, based on their capacity and the degrees of their neighbors.
-During this process, the target of the handshake request accepts the requester whenever he does not know enough peers or the requester as more capacity than at least one of his neighbors.
-To avoid disconnecting poorly connected peers, the target of the handshake request will replace the highest capacity peer whose capacity is smaller than that of the requester.
-In order to avoid hot-spots or overloading of peers, a peer is only allowed to send requests to a neighbor if that neighbor has explicitly given him one Token; a token represents one request the neighbor is willing to accept; Each peer allocates tokens at the rate at which they can answer requests.
-
-Gia successfully promotes optimal performance and longevity on the systems overlay topology through continuous optimization.
-Unlike many other unstructured overlays, Gia does not use message flooding to disseminate information, minimizing overhead.
-There is, however, one fundamental problem with Gia, related to the promotion of fairness; according to their capacity, peers, receive tokens from their neighbors, hence decreasing the usability of the system by low capacity peers.
-Either way, free riding is not a concept that directly applies to our proposal.\newline
+The callee accepts the caller whenever he does not know enough peers or the caller as more capacity than at least one of his neighbors; to avoid disconnecting poorly connected peers, the callee will replace the highest capacity peer whose capacity is smaller than that of the caller.
+In order to avoid hot-spots or overloading of peers, a peer is only allowed to send requests to a neighbor if that neighbor has explicitly given him one Token representing one request that neighbor is willing to accept; Each peer allocates tokens at the rate at which they can answer requests.
+The results show that Gia manages to reduce network overhead, as expected since it does not use flooding, but is also up to five orders of magnitude faster than Gnutella performing lookups.
+Gia successfully promotes optimal performance and longevity of the system through continuous, peer carried optimizations.
+However, we identify one fundamental problem with Gia, related to the systems attempt to load-balance super-peers and implicit free-rider avoidance\footnote{Peers who consume more community resources than they contribute with.}; According to their capacity, peers receive tokens from their neighbors, hence decreasing the number of tokens received by low-capacity peers and, thus, system usability.
 

### 2020-01-01 docs/thesis-project/article.tex
Update article.tex

@@
 
-\subsubsection{Kademlia} \cite{kademlia} is fully decentralized and designed for key-value storage systems, and according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology.
+\subsubsection{Kademlia}\cite{kademlia} is fully decentralized and designed for key-value storage systems, and according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology.
 Kademlia peer-identifiers and data-keys are built as in Chord\cite{chord}, using consistent hashing.
@@
 
-\subsubsection{Self-Chord} \cite{selfchord} is a bio-inspired overlay designed for cloud or grid computing architectures.
+\subsubsection{Self-Chord}\cite{selfchord} is a bio-inspired overlay designed for cloud or grid computing architectures.
 Its foundations are that of the standard Chord, both in terms of identifier generation and the way peers connect, i.e., in ring-like shapes.
@@
 
-\subsubsection{Gia} \cite{gia} seeks to overcome the drawbacks of Gnutella\cite{gnutella-rfc}, which was the first decentralized, unstructured P2P protocol.
+\subsubsection{Gia}\cite{gia} seeks to overcome the drawbacks of Gnutella\cite{gnutella-rfc}, which was the first decentralized, unstructured P2P protocol.
 Gnutella's main issues were the lack of scalability, as resource discovery was made with flooding, and how easily some peers became overloaded when faced with high rates of aggregate queries.
@@
 
-\subsubsection{BlatAnt} \cite{blatant} is a bio-inspired, unstructured overlay, designed for grid computing architectures where resource discovery needs to be efficient; to this end, BlatAnt focuses solely on bounded-diameter optimization in order to minimize network overhead and lookup latency, by using swarm intelligence to support flood-like discovery protocols.
-Like in Self-Chord, ant agents are used to optimize and maintain a self-structured P2P overlay, conversely, in BlatAnt the ants do not change the location of data items from peer to peer, but instead build and optimize the overlay such that the network uses a minimal number of connections between peers and such that the diameter of the overlay is bounded.
-The idea is that by doing so, the average number of hops, i.e., average path length, required to obtain a data item from a lookup will be smaller, thus achieving the sought goals.
+\subsubsection{BlatAnt}\cite{blatant} is a bio-inspired overlay designed for grid computing architectures where resource discovery needs to be efficient; BlatAnt focuses solely on network bounded-diameter optimization as means of minimizing network overhead and lookup latency, using swarm intelligence to support flood-like discovery protocols.
+Like in Self-Chord, ant agents optimize and maintain the overlay; conversely, the ants do not change the location of data items, but instead reorganize peer connections in such a way that the global overlay uses as little of them as possible.
+The idea is that by doing so, the average number of hops, i.e., average path length, required to obtain a data item from a lookup will be smaller.
 BlatAnt also uses local index caching to improve efficiency further; The authors argue that local caches, as an alternative to semantic clustering such as the ones used in Self-Chord, are better for self-organized overlays where a stable topology can not be guaranteed.
 The overlay management can be viewed in three major components.
-
-The first component are the peers themselves.
+The first component are the \textit{peers} themselves.
 Each peer keeps two structures, a fixed-size $\alpha$-table, i.e., a partial view of the overlay retaining neighborhood information used to evaluate the redundant connections or the need for new ones, and, a neighborhood-set \textit{N}, containing peer identifiers.
 Peers contribute to the optimization by rearranging local connections according to the connection rule, which reduces the diameter of the overlay, and disconnection rule, which discards redundant connections.
-
-The second component are the ant agents, which have multiple species; \textit{discovery ants} are randomly spawned by peers and live for a limited time.
+The second component are the ants, which have multiple species; \textit{discovery ants} are randomly spawned by peers and live for a limited time.
 They wander across the overlay, collecting information about its topology and update the $\alpha$-tables of peers they meet along the way.
@@
 Finally, \textit{Ping Ants} are periodically exchanged between peers to keep their connections alive in low traffic situations.
-
-The last component are pheromone trails, which evaporate overtime.
-A trail is a value assigned to a connection between peers, whenever an ant agent, or a query, walks over that connection, they increase the pheromone concentration increases on both ends.
+The last component are \textit{pheromone trails}, which evaporate overtime.
+A trail is a value assigned to a connection between peers, whenever an ant or query, walks over that connection, they increase the pheromone concentration on both ends.
 Discovery ants will tend to follow paths with less pheromone concentration, thus incentivizing full network coverage during the exploration.
@@
 Thus pheromone trails provide seamless error resolution.
-
-BlatAnt proved to accomplish its goals, quickly converging to stable overlays with bounded-diameter, even under very dynamic conditions, resulting in faster lookups due to the reduction of flood messages but also due to caching.
-Note that BlatAnt can leverage random walks such as the ones used in Gia for even less network overhead.
-BlatAnt also introduces optimizations over caching procedures; we do not review them here, because they are based on complex profile similarities and because caching is out of the scope of our proposal.\newline
+BlatAnt quickly converges to stable overlays with bounded-diameter, even in very dynamic conditions, resulting in faster lookups due to a decrease in flood messages and also due to caching.
+Note that BlatAnt could also leverage random walks such as the ones used in Gia for even less network overhead.
+BlatAnt also introduces optimizations over caching procedures; we do not review them here, because they are based on complex profile similarities and caching is out of the scope of our proposal.
 

### 2020-01-01 docs/thesis-project/article.tex
Update article.tex

@@
 \section{Membership Management Frameworks}\label{sec:membership-management}
-
-Gossip-based protocols, also know has epidemic protocols, gained popularity in the nineties for their usability to solve problems like database replication, failure detection, and resource-monitoring.
-In this section, we present two such protocols used for membership management, i.e., used to keep current partial views of large-scale, dynamic networks, built on unstructured overlays.
-Just like we mentioned in section \ref{sec:p2p-overlays}, keeping a view of all peers in the system is not a solution.
-Another fundamental issue of unstructured networks is to avoid partitioning.
+Gossip-based protocols are well known for their usability to solve problems like database replication, failure detection, and resource-monitoring.
+We present two frameworks used for membership management that ensure peers keep up to date partial views of large-scale, dynamic networks, built on unstructured overlays using such protocols.
+Just like we mentioned in section \ref{sec:p2p-overlays}, keeping a view of all peers in the system is often not a solution.
+Another fundamental issue in P2P is how to avoid partitioning.
 The following alternatives are both scalable, robust, and decentralized solutions to both these problems.
-Likewise, peers act autonomously in order to provide self-organization or self-healing capabilities to the overlays in which the protocols are applied.
+They gift peers with autonomous capabilities regarding self-organization and overlay healing.
 
@@
 Upon reception of subsets \textit{s} and \textit{s'}, \textit{j} and \textit{i}, respectively, update their \textit{neighbors} by first using any empty slots in their partial views and then by replacing entries which they sent to the other, i.e., if \textit{i} has no more slots, he replaces, in its \textit{neighbors} structure, the entries that he sent in \textit{s} to \textit{j}, with the ones he received in \textit{s'} from \textit{j}.
-Note that, after \textit{i} initiates a shuffle with \textit{j}, \textit{i} becomes \textit{j}'s neighbor, but \textit{j} is no longer a neighbor to \textit{i}, i.e., the neighboring relation between \textit{i} and \textit{j} is reversed.
-Unlike many protocols, Cyclon links between peers are not bidirectional.
-
+Note that, after \textit{i} initiates a shuffle with \textit{j}, \textit{i} becomes \textit{j}'s neighbor, but \textit{j} is no longer a neighbor to \textit{i}, effectively reversing neighboring relation between both peers\footnote{Cyclon links are not bidirectional, conversely to the majority of protocols.}.
 The age is essential in the shuffling algorithm, for two reasons.
-First, it limits the lifetime of each peer in \textit{neighbors} structures, which globally bounds the number of existing pointers to them, it also limits the time each peers' addresses are passed around until they are selected as shuffling targets, resulting on a more up-to-date overlay as well as uniform distribution of each peers' addresses over the network.
-
+First, it limits the lifetime of each peer in neighboring structures, which globally bounds the number of existing pointers to them.
+Secondly, it also limits the time each peers' addresses are passed around until they are selected as shuffling targets, resulting in a more up-to-date overlay as well as uniform distribution of each peers' addresses over the network.
 When a peer leaves the overlay for any reason, the remaining peers may have to remove him from their \textit{neighbors} structure.
-Timely removal is fundamental for the robustness of the overlay.
-Cyclon uses passive detection.
-Whenever a peer \textit{i} contacts \textit{j} for shuffling and obtains no response it assumes \textit{j} is disconnected and removes it from his \textit{neighbors}.
-Since the time for \textit{i} to contact \textit{j} is bounded by the age property in \textit{neighbors}, detection is accelerated.
+Many authors argue that timely removal is fundamental for the robustness of the overlay, Cyclon uses a reactive pessimistic policy that works as follows: Whenever a peer \textit{i} contacts \textit{j} for shuffling and obtains no response it assumes \textit{j} is disconnected and removes it from his \textit{neighbors}.
+Since the time for \textit{i} to contact \textit{j} is bounded by the age property, detection is accelerated, which is a property Gia does not have.
 
@@
 Another exciting result of Cyclon is that the average path length increased logarithmically, and clustering decreased exponentially, as the number of peers in the network increased, hence demonstrating scalability and increased robustness.
-
-The authors also propose a join method that does not disrupt the randomness of the obtained overlays, while simultaneously making new peers indistinguishable from old ones.
+Noting this property, the authors propose a join method that does not disrupt the randomness of the obtained overlays.
 Whenever \textit{i} wishes to join the overlay, he contacts one peer \textit{j} already in the overlay who initiates \textit{c} random-walks with time-to-live (TTL) equal to the average path length of the overlay.
@@
 Note that even if some of the random walks do not complete due to byzantine failures, \textit{i} will remain connected to the overlay and will eventually find new acquaintances.
-
-We conclude by reinforcing that Cyclon provides good self-healing and self-organization capabilities without complex algorithms or apparent downsides other than the one where too few peers in the system may cause it to be under performant due to clustering effects.
-In a way, it relates to BlatAnt since the proposal passively bounds the diameter of the network.
+Cyclon as no apparent downsides other than the one where too few peers in the system may cause it to be under performant due to clustering effects, but this is expected in other solutions too.
+We note that in a certain way, this framework relates to BlatAnt since it implicitly bounds the diameter of the network.
 
@@
 In Newscast \cite{newscast-computing}, all peers proactively exchange, timestamped, information with each other, including not only neighboring information but also app-specific information, allowing for simplified implementation of aggregation algorithms, i.e., finding statistics regarding for example network size.
-Newscast is responsible for membership-management and a second function, which does not exist in Cyclon, which is information dissemination.
-
-In Newscast, there is a separation of responsibilities.
-Within each peer, there are two types of entities.
-Those who run the newscast protocol, are called \textit{corespondents}.
-Those who run processes of a distributed application are called \textit{agents}.
-\textit{Correspondents} run the same instance of the newscast protocol.
-\textit{Agents} are not required to run the same application.
-Each \textit{agent} has one or more \textit{correspondents}.
-This separation does not explicitly exist in Cyclon.
+Newscast is thus responsible for both membership management as well as information dissemination.
+Within each Newscast peer, there are two types of entities.
+Those who run the actual newscast protocol, are called \textit{corespondents}.
+Those who run the distributed application are called \textit{agents}.
+\textit{Correspondents} all run the same instance of the newscast protocol.
+\textit{Agents} are not required to run the same distributed application.
+Each \textit{agent} has one or more \textit{correspondents}; this responsibility separation does not explicitly exist in Cyclon.
 Each \textit{agent} implements an interface with two functions \textit{getNews()} and \textit{newsUpdate(news[])}.
 The \textit{correspondent} uses the former to request news from their \textit{agent}, and the latter to deliver app-specific news collected from other peers' \textit{correspondents} in the network.
-Each \textit{correspondent} as a fixed-size news cache of size \textit{c} and periodically exchange news with each other has followed: request news to the \textit{agent}, timestamp them along with the local time, and IP-address of the \textit{agent}, then storing them in a cache entry.
-Afterward, it selects another peer's IP address running the same \textit{correspondent} aggregation function, as available in his cache.
+Each \textit{correspondent} as a fixed-size news cache of size \textit{c} and periodically exchange news with each other has followed: request news to its \textit{agent}, timestamp them along with the local time, and IP-address of the peer, then store the news in a cache entry.
+Afterward, it selects another peer's IP address running the same type of \textit{correspondent}, as available in his cache.
 Both \textit{correspondents} exchange the entirety of their caches and pass the merged cache with $2c+1$ entries to their respective \textit{agents} before discarding the oldest entries, after timestamp normalization, down to \textit{c} and storing the results themselves.
 The timestamp normalization is not a clock-synchronization process and has some inaccuracy, but as long as transmission times between two peers are not too big, there should be no problem.
-During the discarding process, the algorithm ensures that there is at most one news item per \textit{agent}.
+During the discarding process, the algorithm ensures that there is at most one news item per \textit{agent} in the cache.
 
-It is debatable if Newscast has an advantage over Cyclon since news exchange messages already include peer addresses, leading to a transparent resolution of the membership problem.
-Cyclon, however, has several advantages over Newscast.
+Newscast as a slight edge over Cyclon, due to graciously accounting for aggregation functions in its native protocol.
+Similarly, inactive peers are eventually forgotten due to the timestamp property, but we believe that Cyclon has several significant advantages.
 These include: the resulting overlays of Cyclon have better randomness due to its shuffling-based algorithm, whereas Newscast and other gossip-based management frameworks, seem to have small-world properties according to previous research \cite{eval-gossip-based}.
-Like in Cyclon, disconnected peers are eventually forgotten due to the timestamp property of the newscasting process.
+Also, Cyclon uses logical timestamps; hence, no clock synchronization is required, not even a relaxed one.
 Newly joined peers' cache initialization is based on a regular news exchange with one or more peers who are already members of the overlay.
-In contrast, Cyclon uses \textit{c} random-walks exchanging exactly one cache entry with possibly different peers, which in turn helps to maintain the randomness properties; finally, while Newscast managed to achieve their goals of providing a robust, scalable, adaptive and lightweight solution for simultaneous membership management and aggregation Cyclon exhibits less network overhead\cite{cyclon}.
+In contrast, Cyclon uses \textit{c} random-walks exchanging exactly one cache entry with possibly different peers, which in turn helps to maintain overlay randomness; finally, while Newscast managed to achieve their goals of providing a robust, scalable, adaptive and lightweight solution, Cyclon exhibits less network overhead and can easily be extended to perform aggregation.
 
@@
 
-Many academic and commercial projects have been developed over the years where cloud, grid, and other centralized architectures are partly assisted by P2P networks.
-In these systems, peers are usually machines belonging to the clients of the service, and they help to decrease the load on the system, hence reducing its cost.
-This approach is especially prevalent in streaming and online gaming services, for example, Blizzard Activision uses their player base to help distribute installers and patches in a P2P fashion, and Digital Extremes uses, in their free-to-play game Warframe, P2P networks in particular, but not all, group missions where individuals are in instances of the game inaccessible by players outside the group.
-However, only recently, was the alternative explored, where P2P networks serve the vast majority of the service and are assisted by stable cloud services only to help guarantee some quality of service property of the system under critical situations.
-We present two of them.
+Many academic and commercial projects have been suggested and used over the years where cloud and other centralized architectures are partly assisted by P2P networks; In these systems, peers are usually clients of the service, who help to decrease the load on the central system, hence reducing its cost.
+Examples include Spotify and Wuala until approximately five years ago.
+However, these approaches are still prevalent in streaming and online gaming services.
+Blizzard Activision uses their player base to help distribute installers and patches in a P2P fashion, and Digital Extremes uses, in their free-to-play game Warframe, P2P networks in particular, but not all, group missions where individuals are in instances of the game inaccessible by players outside the group.
+However, only recently, was the alternative experimented, where P2P networks serve the vast majority of the service and are assisted by stable cloud services only to help guarantee some quality of service property of the system under critical situations.
 
 \subsection{Cloudcast}
-Cloudcast \cite{cloudcast} is a P2P-assisted cloud architecture that differs from other proposals because it does not offload the provided service to peers only when service availability is not affected.
-Furthermore, the architecture does not augment P2P systems with elastic computing nodes that perform particular tasks beyond the reach of the P2P network, such has bulk-synchronous content distribution.
-Cloudcast functionality depends solely on a passive storage service that transparently integrates into a gossip-based P2P system.
+Cloudcast \cite{cloudcast} is a P2P-assisted cloud architecture that differs from previous approaches because it does not offload the provided service to peers only when service availability is not at risk of being affected.
+Furthermore, the architecture does not augment P2P systems with elastic computing nodes that perform particular tasks beyond the reach of the P2P network, e.g., bulk-synchronous content distribution.
+Cloudcast functionality depends only on a passive storage service that transparently integrates into a gossip-based P2P system.
 The reasoning behind not using elastic computing instances is to eliminate fixed renting costs.
@@
 The architecture uses two gossip protocols, one for membership management and another for information dissemination.
-Key results of the proposal include: accesses to the cloud service are bounded regardless of the number of clients in the system; the cloud service only distributes content when few peers are available; the overlay does not suffer from partitioning regardless of the number of peers in the system; the two illustrated use cases have an absurdly low yearly cost when compared to pure cloud approaches, and finally the authors recognize that the vast part of their bill is tied to membership management and not to actual storage costs, thus using strategies that adapt to the size of the network could yield even better savings.
+Key results of the proposal include: i) accesses to the cloud service are bounded regardless of the number of clients in the system; ii) the cloud server only distributes content when few peers are available; iii) the overlay does not suffer from partitioning regardless of the number of peers in the system; iv) the two illustrated use cases have an absurdly low yearly cost when compared to pure cloud approaches; iv) authors imply that most of the expected bill is tied to membership management and not to actual storage costs, suggesting that strategies that adapt to the size of the network could yield better savings.
 
@@
 From other alternatives, Cyclon was chosen not for its particular inexpensiveness, but because the protocol maintains a random overlay topology that works even during high churn periods and even in the presence of catastrophic failures where up to 80\% of the peers fail.
-Cyclon properties also guarantee that references to the cloud server are not too many keeping usage cost under control.
-Finally, the cloud server can easily integrate in the protocol, in particular, Cyclon does not require the cloud server to initiate any communications, i.e., its active roles can be carried out by peers interacting with it; Finally, the cloud server can easily integrate into the protocol, in particular, Cyclon does not require the cloud server to initiate any communications, i.e., its active roles can be carried out by peers interacting with it; In this regard, the dummy process carried by the peers occasionally refreshes pointers to the cloud to avoid losing all references to it in the system.
-To deal with the case where all references cloud are lost due to the dynamic characteristics of the environment, e.g.
-churn and lost messages, all peers keep a logical timestamp of the last round they heard about the cloud server updating it whenever they successfully contact it or because one of their a \textit{neighbor} heard about it more recently then them, if after \textit{t} rounds peer has not updated this timestamp he will create a new cloud reference with probability $p\textsubscript{recovery}$.
-
-The second problem is solved using two different implementations of gossip algorithms explained in depth in \cite{epidemic_algorithms}.
-They are a \textit{rumor-mongering, coin and blind} approach which favors fast news spreading and consists in having a peer \textit{i} periodically forwarding updates he learns about to some other peer \textit{j}, then deciding with probability $p\textsubscript{rumor}$ if he should, permanently, stop gossiping the update; And a \textit{anti-entropy, push-pull} approach which guarantees that news are eventually known by all participants of the system and consists in having each peer \textit{i} periodically contacting a random neighbor \textit{j} to exchange update they know about.
-All updates to the data are recorded in the cloud server through a PUT operation before becoming hot-topics, thus triggering rumor-mongering.
+Cyclon properties also guarantee that cloud server has a bounded in-degree, keeping usage cost under control.
+Finally, the cloud server can easily integrate into Cyclon because the latter does not require the cloud server to initiate any communications, i.e., its active roles can be carried out by peers interacting with it; In this regard, the dummy process carried by the peers occasionally refreshes pointers to the cloud to avoid losing all references to it in the system.
+To deal with the case where all cloud references are lost due to the dynamicity of the environment and unreliability of the network layer, all peers keep a logical timestamp of the last round they heard about the cloud server updating it whenever they successfully contact it or because one of their a \textit{neighbor} heard about it more recently then them.
+If after \textit{t} rounds a peer has not updated this timestamp he will create a new cloud reference with probability $p\textsubscript{recovery}$.
+The second problem is tackled by implementing two gossip algorithms.
+They are a \textit{rumor-mongering, coin and blind} approach which favors fast news spreading and consists in having a peer \textit{i} periodically forwarding updates he learns about to some other peer \textit{j}, then deciding with probability $p\textsubscript{rumor}$ if he should stop rumoring; And a \textit{anti-entropy, push-pull} approach which guarantees that news are eventually known by all peers by having each peer \textit{i} periodically contacting a random neighbor \textit{j} to exchange updates they know about.
+All updates to the data are recorded in the cloud server through PUT operations before being rumored.
 Peers never contact the cloud server for rumor-mongering purposes.
-The anti-entropy approach can be piggybacked to the Cyclon peer-sampling step, thus decreasing message overhead; the authors, however, do not mention this optimization.
+We suggest that anti-entropy messages can be piggybacked to the peer-sampling messages, decreasing the number of messages passed in the network; the authors, however, do not implement this optimization.
 
@@
 
-\begin{algorithm}[ht]
+\begin{algorithm}[t]
 \caption{MakeDecision(N\textsubscript{e}), replicaOverlay}
@@
 Following the suggestions of Cloudcast authors, H.
-Kavalionak \textit{et al.}, developed a decentralized algorithm \cite{marriage_of_convinience} that aims to do replica management in cloud-based, peer-assisted applications.
+Kavalionak \textit{et al.}, developed a decentralized algorithm\cite{marriage_of_convinience} that aims to do replica management in cloud-based, peer-assisted applications.
 Because the proposed work has no running title, we shall refer to it as ARM.
 The goal of the ARM is to maintain an appropriate replication level, despite churn, such that, at all times, at least one replica is available and ensuring that replicas survive failures.
-Furthermore, the proposal acknowledges that replicas should be synchronized, i.e., the system provides at least some consistency guarantees.
-In simple terms, this is perhaps the main difference between the proposed work and Cloudcast, which stores all updates in an always-on passive cloud server, helping with information diffusion and overlay connectivity and overall reliability of the system in extreme conditions.
-Cloudcast monetary cost reductions are primarily associated with limiting access to the cloud server.
-ARM effectively reduces those costs even further by performing timely switches between Cloudcast and regular Cyclon protocols, depending on the size of the network.
-To estimate the network size without centralization or broadcasting algorithms that would ultimately lead to either bottlenecks or overhead, ARM leverages a gossip-based aggregation method \cite{gossip-based_aggregation} that allows each peer to infer the approximate size of the network autonomously, at the end of each epoch.
+Furthermore, the proposal acknowledges that replicas should be synchronized, i.e., the system should support session guarantees.
+The main difference between ARM and Cloudcast is that in the latter, the cloud server stores all updates in an always-on passive cloud server helping with information dissemination when needed, whereas, in the former, the cloud does not participate in the system at all, unless conditions are extreme.
+Cloudcast monetary cost reductions are thus, primarily associated with limiting access to the cloud server, and ARM effectively reduces those costs further by performing timely switches between Cloudcast and regular Cyclon protocols, depending on the size of the network.
+To estimate the network size without centralization or broadcasting that would ultimately lead to either bottlenecks or overhead, ARM leverages a gossip aggregation method that allows each peer to infer the approximate size of the network autonomously, at the end of each epoch.
 Estimation accuracy is dependent only on the size of the epoch.
@@
 
-\newpage\section{Probabilistic Swarm Guidance}\label{sec:psg}
+\section{Probabilistic Swarm Guidance}\label{sec:psg}
 Markov chains describe a sequence of possible events in which the probability of each event depends only on the previous event states, often used to model real-world processes, such as queues of customers arriving at an airport and population growths.

### 2020-01-02 docs/thesis-project/article.tex
Update article.tex

@@
 
-Following the suggestions of Cloudcast authors, H.
-Kavalionak \textit{et al.}, developed a decentralized algorithm\cite{marriage_of_convinience} that aims to do replica management in cloud-based, peer-assisted applications.
-Because the proposed work has no running title, we shall refer to it as ARM.
-The goal of the ARM is to maintain an appropriate replication level, despite churn, such that, at all times, at least one replica is available and ensuring that replicas survive failures.
-Furthermore, the proposal acknowledges that replicas should be synchronized, i.e., the system should support session guarantees.
-The main difference between ARM and Cloudcast is that in the latter, the cloud server stores all updates in an always-on passive cloud server helping with information dissemination when needed, whereas, in the former, the cloud does not participate in the system at all, unless conditions are extreme.
-Cloudcast monetary cost reductions are thus, primarily associated with limiting access to the cloud server, and ARM effectively reduces those costs further by performing timely switches between Cloudcast and regular Cyclon protocols, depending on the size of the network.
-To estimate the network size without centralization or broadcasting that would ultimately lead to either bottlenecks or overhead, ARM leverages a gossip aggregation method that allows each peer to infer the approximate size of the network autonomously, at the end of each epoch.
-Estimation accuracy is dependent only on the size of the epoch.
-Likewise, the execution cost depends on the length of the epochs and the network churn rate, but not on the number of peers in the system, which means that ARM keeps the scale-free properties of the unstructured overlays that used in Cloudcast.
-The peers randomly organize themselves according to the Cyclon peer-sampling protocol, but depending on the results of the aggregation, the passive cloud-server used in Cloudcast is removed or kept in the overlay.
-In ARM, each data object to be replicated as its \textit{replica-overlay} on top of the regular unstructured overlay, which facilitates synchronization and cloud-usage decisions.
-The ARM algorithm is designed to consider four replica-overlay states, separated by three size thresholds: \textit{Critical (C)}, \textit{Sufficient (S)} and \textit{Redundant (S)}.
-\textit{C} depends on the chosen redundancy method, i.e., file-level \textit{vs.} block-level replication and is the sum of the minimum number of replicas required for data recovery plus the number of peers that will leave during the cloud backup replication phase.
-\textit{S} depends on the churn-rate and is the sum of \textit{C} plus the number of peers expected between two successive recovery phases.
-Finally, \textit{R} is a system parameter dependent also on the chosen redundancy method and is application requirements specific.
-The aggregation method mentioned earlier is the foundation \textit{monitoring phase}, which consists of a fixed number of gossip and idle rounds, allowing a peer to infer the \textit{expected size (N\textsubscript{e})} of the replica-overlay and decide if he should add or remove the cloud or peers to the replica-overlay as well as the protocol he uses, based on Algorithm \ref{alg:makedecision}.
-After each monitoring phase, each peer calculates how long it should wait before executing the \textit{monitoring phase} again.
+Following Cloudcast suggestions, H.
+Kavalionak \textit{et al.}, developed a solution\cite{marriage_of_convinience} that we shall refer to as ARM, with the goal of maintaining an appropriate data replication level, despite churn and, in such a way that peers sharing the same data object are synchronized, i.e., the system should support session guarantees.
+The main difference between ARM and Cloudcast is that in the latter, the cloud server stores all updates in an always-on passive cloud server that helps with information dissemination, whereas, in the former, the cloud does not participate in the system at all, unless conditions are extreme.
+Cloudcast monetary cost reductions result from limiting cloud accesses, and ARM effectively reduces those costs further by performing timely switches between Cloudcast and regular Cyclon-based protocols, depending on the size of the network.
+To estimate that size without centralized or broadcast solutions that could, respectively, lead to bottlenecks or bandwidth waste, ARM periodically runs a monitoring phase whose accuracy dependents only on its duration and whose execution cost depends on the network churn rate, but not on the number of peers in the system.
+Peer neighboring relations are built naturally from Cyclon peer-sampling, but depending on the results of the aggregation, the cloud server is either removed or kept in the overlay.
+The decision to add or remove the cloud depends on the following threshold: \textit{Critical (C)} equals the minimum number of replicas required for data recovery plus the number of peer faults the system must support during cloud replication.
+\textit{Sufficient (S)} depends on the churn-rate and equals \textit{C} plus the number of peers expected between two successive recovery phases.
+Finally, \textit{Redundant (S)} is an application-specific system parameter.
+The monitoring phase thus, consists of a fixed number of gossip and idle rounds, allowing a peer to infer the \textit{expected size (N\textsubscript{e})} of the overlay based on Algorithm \ref{alg:makedecision}.
+After each monitoring phase, peers independently calculate how long they should wait before re-executing the phase.
 That time depends mostly on \textit{N\textsubscript{e}}, e.g., if the state is below the critical threshold, the monitoring phase will initiate earlier than if it is above the redundant threshold, allowing the system to respond to overlay changes promptly and to keep the overlay size in a range that minimizes overhead but ensures that data does not become unavailable.
 The authors do not specify how much savings the system was capable of achieving using ARM; however, we can assume it is a substantial value; In Fig.
-\ref{fig:arm_results}, the plot on top displays the number of peers in the underlying network as a function of time, and the bottom plot shows the in-degree to the cloud server as a function of time, comparing Cloudcast and ARM (labeled as "our protocol").
+\ref{fig:arm_results}, the plot on top displays the number of peers in the underlying network as a function of time, and the bottom plot shows the in-degree to the cloud server as a function of time, comparing Cloudcast and ARM\footnote{ARM is labeled as "our protocol"}.
 
@@
 
-\newpage\section{Data Redundancy and File Survivability}\label{sec:dataredundancy}
+\section{Data Redundancy and File Survivability}\label{sec:dataredundancy}
 In distributed storage systems, not losing files is of the utmost importance.

### 2020-01-02 docs/thesis-project/article.tex
Update article.tex

@@
 \includegraphics[width=\textwidth]{images/armresults.PNG}
-\caption{Cloud-Server in-degree, Cloudcast \textit{vs.} ARM \cite{marriage_of_convinience}}
+\caption{Cloud-Server in-degree, Cloudcast \textit{vs.} ARM}
 \label{fig:arm_results}
@@
 
-\section{Data Redundancy and File Survivability}\label{sec:dataredundancy}
-In distributed storage systems, not losing files is of the utmost importance.
-Particularly, since peer availability in P2P networks is ever-changing and not guaranteed, the availability and even durability of a file may be endangered.
+\section{Data Redundancy}\label{sec:dataredundancy}
+In DFS, not losing data is of the utmost importance.
+Particularly, peer availability in P2P networks is ever-changing and not guaranteed, and thus, availability, and even durability of files are at risk.
 A \textbf{\textit{durable file}} is one that, once put into the system, is not lost due to failures, whereas availability means that a user can access it promptly.
-For clarity, an available file is durable, but the other way is not necessarily correct.
-Distributed systems as a whole, usually ensure durability by employing file-level or block-level \textbf{\textit{replication}}; alternatively, some systems adopt \textbf{\textit{erasure coding}} techniques (\textit{coding} for short) or a mixture of both approaches.
-Replication has the advantage of having a more straightforward and less resource-intensive recovery model, on the other hand, coding has a better \textit{expansion factor}, i.e., can provide the same or even increased durability at a lower storage overhead.
-Because our proposal focuses on the survivability of files in P2P networks, using swarm guidance, where resources may be limited, we care only about coding in the survey that follows, briefly mentioning some of the raised concerns and inherent disadvantages of this technique.
+Durability is accomplished by employing file or block-level \textbf{\textit{replication}} or alternatively, using \textbf{\textit{erasure coding}} techniques (\textit{coding} for short).
+Hybrid approaches are also possible.
+Replication has the advantage of having a straightforward and less resource-intensive recovery model, on the other hand, coding has a better \textit{expansion factor}, i.e., can provide the same or even increased durability at a lower storage overhead.
+Because our proposal focuses on the survivability of files in P2P networks built on top of personal computers, using swarm guidance, where resources may be limited, we focus on surveying coding, mentioning raised concerns and inherent disadvantages of this technique.
 

### 2020-01-02 docs/thesis-project/article.tex
Update article.tex

@@
 Replication has the advantage of having a straightforward and less resource-intensive recovery model, on the other hand, coding has a better \textit{expansion factor}, i.e., can provide the same or even increased durability at a lower storage overhead.
-Because our proposal focuses on the survivability of files in P2P networks built on top of personal computers, using swarm guidance, where resources may be limited, we focus on surveying coding, mentioning raised concerns and inherent disadvantages of this technique.
+Because our proposal focuses on durability of files in P2P networks built on top of personal computers, where resources may be limited, we survey only coding, mentioning raised concerns and inherent disadvantages of this technique.
 
-Two parameters specify some of the most widely used coding algorithms, the number of data symbols \textit{k} to be encoded, and the number of coded symbols \textit{n} to be produced, both assumed to be in a finite field $GF(2^w)$ and these are referred as \textit{(n, k) erasure-codes}.
-The fundamental idea, is that a file of size $S$ is divided into $k$ equally sized fragments and encode them into $n$ encoded fragments, allowing the original file to recovered from any set of encoded-fragment with count $k$.
-Systems that use these and other more sophisticated coding are evaluated under some of the following metrics, the \textbf{\textit{repair-bandwidth}}, i.e., the number of bits communicated in the network during normal and repair operations, the \textbf{\textit{disk I/O}} which is the number of bits read from disk during each repair, the \textbf{\textit{repair locality}} which is the number of peers that would be involved in the repair process; Since coding algorithms are slow processes CPU cycles, XOR operation counts and cache misses are metrics often researched as well \cite{fast_coding}, when the main interest is tied with load-balancing and cost savings.
-We take particular interest in the first metric because, in general, coding may lead to unreasonable bursts in bandwidth consumption when availability requirements, in very dynamic systems, are high \cite{coding-problems}, even for today's standards.
-Whenever a peer fails, all files he held need to be quickly recovered by some peer who needs to gather enough fragments from possibly many different peers.
-Such burst in consumption may lead to delays or even failure of the repair process due to physical link capacity between peers, who are involved in the repair process, or even throughput constraints enforced by each peer.
-When it comes to redundancy in a distributed storage systems, one fundamental question is deciding when to replicate or recover data; this is often called the replica-control problem.
-
-There are two main approaches, reactive and proactive.
-Reactive approaches are the least complex, and consist of recovering data whenever a failure is detected; they suffer the most from bandwidth bursts, and mitigations to this problem include use of predictive models and thresholds to avoid unnecessary recoveries \cite{lifetime-reactive, efficient-replica-mng-reactive}.
-Proactive approaches seek to smooth bandwidth consumption by injecting new redundancy data into the network overtime at a fixed-rate \cite{proactive-rep, reliability-without-availability}.
-
-The most widely known coding algorithms belong to the \textit{Maximum Distance Separable} (MDS) class, this class is likely to offer the best reliability-redundancy trade-off.
-Regardless of the chosen approach, research \cite{network-coding-for-dss} suggests that for coding to be competitive in distributed storage systems, one should seek to minimize these costs, while maximizing redundancy to only the amount necessary to ensure a specific service guarantee.
-They identify \textit{Minimum Bandwidth Regenerating} (MBS), and all codes in between these two ends of the spectrum as \textit{regenerating codes}, the differences between codes on the spectrum depend on the values such has \textit{n}, \textit{k} and on the implementation of the algorithms themselves.
-Shokrollahi \textit{et al.} introduces raptor codes \cite{raptor-codes}, which are a linear-time coding and decoding extension of LT codes.
-LT codes belong to the class of rateless codes, i.e., a potentially infinite sequence of encoded fragments can be generated from a given data source.
-Furthermore, the original data is recoverable from any subset of the encoded fragments of size equal to or slightly larger than original data size; These rateless code implementations are of particular importance because they are scalable with regards to the machine and network resources,  and fault-tolerant concerning dynamic and heterogeneous network environments.
-Finally, Zhou \textit{et al.} \cite{fast_coding} perform a study on state of the art techniques that improve coding efficiency through bitmatrix optimization, vectorization, and scheduling and prove that when used in conjunction much better levels of performance can be delivered, e.g., 552.27\% when compared to simple MDS implementations, in particular, the Reed-Solomon algorithm.
+Commonly coding algorithms receive two parameters, the number of data symbols \textit{k} to be encoded, and the number of coded symbols \textit{n} to be produced; these are referred to as \textit{(n, k) erasure-codes}.
+The fundamental idea, is that a file of size $S$ is divided into $k$ equally sized fragments then turned into $n$ encoded fragments, allowing the original file to recovered from any set of encoded-fragment with count $k$.
+Systems that use (n, k) or more sophisticated codes are evaluated under some of the following metrics: the \textbf{\textit{repair-bandwidth}}, i.e., the number of bits communicated in the network during normal and repair operations, the \textbf{\textit{disk I/O}} which is the number of bits read from disk during each repair, the \textbf{\textit{repair locality}} which is the number of peers that would be involved in the repair process; Since coding is a slow process CPU cycles, XOR counts and cache misses are often evaluated as well\cite{fast_coding}.
+We take particular interest in the first metric because, in general, coding may lead to unreasonable bursts in bandwidth consumption when availability requirements, in very dynamic systems, are high\cite{coding-problems}, even for today's standards.
+Whenever a peer fails, all files he held need to be promptly recovered by some peer who needs to gather enough fragments from possibly many different peers.
+Such burst in consumption may lead to delays or even failure of the repair process due to physical link capacity between peers, who are involved in the repair process, or even throughput constraints enforced by each peer; this leads us to the second key question, which is: at what time should data be replicated or recovered?\footnote{Literature often refers to this question as the replica-control problem} Reactive approaches are the least complex and consist of recovering data whenever a failure is detected; they suffer the most from bandwidth bursts, and burst mitigation includes the use of predictive models and thresholds to avoid or delay recovery\cite{lifetime-reactive, efficient-replica-mng-reactive}.
+Proactive approaches seek to smooth bandwidth consumption by injecting new redundancy data into the network over time at a fixed or variable rate \cite{proactive-rep, reliability-without-availability}.
+Concerning algorithm classification, the most widely known belong to the \textit{Maximum Distance Separable} (MDS) class, which likely offers the best reliability-redundancy trade-off.
+However, research \cite{network-coding-for-dss} suggests that for coding to be competitive in DFS, we need to minimize costs inherent to the MDS class, while keeping redundancy levels to only the amount necessary to ensure specific service guarantees.
+The same research identifies a \textit{Minimum Bandwidth Regenerating} (MBS) class and refers to codes in between these two as \textit{regenerating codes}.
+There is no standard for classification, and it often depends on the obtained trade-offs and thus depends on chosen $n, k$ values, possibly some extra parameters, and actual algorithmic implementation.
+Shokrollahi \textit{et al.} introduces raptor codes \cite{raptor-codes}, which are a linear-time coding and decoding extension of LT codes, belonging to the rateless code class.
+Rateless codes are characterized by the property that a potentially limitless sequence of encoded fragments can be generated from a given source, and the original data is recoverable from any subset of the encoded fragments of size equal to or slightly larger than original data size; Rateless codes differ from MDS codes because they are scalable with regards to both machine and network resources, and fault-tolerant concerning dynamic and heterogeneous network environments.
+Finally, Zhou \textit{et al.} \cite{fast_coding} carry out a study on state of the art techniques that improve coding efficiency through bitmatrix optimization, vectorization, and scheduling, proving that when used in conjunction, fragment generation and data reconstruction can be up to 552.27\% faster.
 

### 2020-01-02 docs/thesis-project/article.tex
Update article.tex

@@
 \section{Probabilistic Swarm Guidance}\label{sec:psg}
-Markov chains describe a sequence of possible events in which the probability of each event depends only on the previous event states, often used to model real-world processes, such as queues of customers arriving at an airport and population growths.
-Markov processes are the basis for stochastic simulation methods known as Markov Chain Monte Carlo (MCMC), which allow one to perform random sampling over vast state spaces, i.e., computing expectations concerning complex, high dimensional probability distributions.
+Markov chains describe a sequence of possible events, in which the probability of each event depends only on the previous event, often used to model real-world processes, e.g., queues of customers arriving at an airport or population growths.
+Markov processes are the basis for stochastic simulation methods known as Markov Chain Monte Carlo (MCMC), which allow us to perform random sampling over vast state spaces, i.e., computing expectations concerning complex, high dimensional probability distributions.
 The idea behind MCMC is creating a Markov Chain, which will, over time, tend to an equilibrium that is close to desired density distribution.
-In robotics and control areas Markov Chains and MCMC are used as the basis foundation for a process known as probabilistic swarm guidance, which gifts autonomous robots with the capability of generating their trajectories independently of each other, in decentralized fashion and without ahead-of-time position allocation, so that sets of robots, known as swarm, converges to the desired distribution shape.
-Swarm guidance has the advantage over other methods for this problem because robots do not necessarily need to communicate with each other to perform their tasks, even if they could gain from it; their overall distribution can be adapted dynamically to the environment, and any damage to their desired formation is eventually self-repaired.
-Demonstration of these properties can be found in recent research by B.
-Açikmeçe \textit{et al.} \cite{psg-mca, psg-caa}, which inspired us to apply swarm guidance principles to P2P networks and file survivability.
+In robotics and control areas Markov Chains and MCMC are used as the foundation for a process known as probabilistic swarm guidance, which gifts autonomous agents with the capability of generating their trajectories independently of each other, in a decentralized fashion and without ahead-of-time position allocation, so that sets of agents converge to the desired shape.
+Swarm guidance has the advantage over other methods for this problem because agents do not necessarily need to communicate with each other to perform their tasks; Furthermore, the agents goal distribution can be adapted dynamically, and any damage to their formations is eventually self-repaired.
+Demonstration of these properties can be found in research by B.
+Açikmeçe \textit{et al.}\cite{psg-mca, psg-caa}, which inspired us to apply swarm guidance principles to P2P networks.
 Boyd et al.
-\cite{fast_mixing_mc} argue that the most popular MCMC methods, Metropolis-Hastings and Maximum-Degree Chain, that allow the creation of transition matrices, that can then be followed independently by autonomous robots in order to practice swarm guidance behavior are not optimal with respect to mixing rates, i.e.,  these algorithms may not result in matrices that converge in small amount of steps towards the desired density distribution.
-They propose that the problem of obtaining such matrices can be formulated as a convex optimization problem and prove that, for most cases, the resulting matrices have higher mixing rates.
-It is reasonable to assume that, concerning our problem in P2P networks, this might open the possibility of imposing additional constraints on the creation of the transition matrix, e.g., bounding the probability of a peer sending a file to some other peer in his hive to increase bandwidth balancing.
+\cite{fast_mixing_mc} argue that the most popular MCMC methods, Metropolis-Hastings and Maximum-Degree Chain, that allow the creation of transition matrices, that can then be followed independently by autonomous agents are not optimal with respect to mixing rates, i.e.,  these algorithms may not result in matrices that converge in relatively small amount of steps towards the desired density distribution.
+They propose that the problem of obtaining such matrices can be formulated as a convex optimization problem and prove that, for most cases, the resulting matrices have better mixing rates.
+It is reasonable to assume that, concerning our proposal, this opens the possibility of imposing additional constraints on the creation of the transition matrix, e.g., bounding the probability of a peer sending a file to some other peer, limiting the number of parts each peer will have, among others, by selecting proper steady-state vectors.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage\section{Solution Proposal}\label{sec:proposal}
+\section{Solution Proposal}\label{sec:proposal}
 In this section, we introduce a prototype architecture that targets the use case that inspired to explore the theme of this thesis. We explain it by decomposition.
@@
 %%%%%%%%%%%%%%%%%%%%%%%
-\newpage\section{Work Schedule}\label{sec:workschedule}
+\section{Work Schedule}\label{sec:workschedule}
 %%%%%%%%%%%%%%%%%%%%%

### 2020-01-03 docs/thesis-project/article.tex
Update article.tex

@@
     \item Implement a replication model that exploits the properties of the $M$
-    \item Evaluate solution performance in a Simulation or on in real life scenario
+    \item Evaluate solution performance in a simulation or on in real life scenario
 \end{itemize}
@@
 \centering
-\includegraphics[width=\textwidth]{images/gfs_paper.png}
+\includegraphics[width=0.8\textwidth]{images/gfs_paper.png}
 \caption{GFS Architecture}
@@
 \subsection{Ceph File System}
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=0.6\textwidth]{images/cephfs.png}
-\caption{CFS Architecture}
-\label{fig:cfs_arch}
-\end{figure}
 
@@
 
-\section{P2P Overlays}\label{sec:p2p-overlays}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.45\textwidth]{images/cephfs.PNG}
+    \caption{Ceph File System Architecture \cite{configure-ceph}}
+    \label{fig:ceph_arch}
+\end{figure}
 
-\begin{figure}[t]
+\newpage\section{P2P Overlays}\label{sec:p2p-overlays}
+
+\begin{figure}[ht]
     \centering
@@
 
-\section{Membership Management Frameworks}\label{sec:membership-management}
+\newpage\section{Membership Management Frameworks}\label{sec:membership-management}
 Gossip-based protocols are well known for their usability to solve problems like database replication, failure detection, and resource-monitoring.
@@
 
-\section{Cloud-Assisted P2P Networks}\label{sec:ca-networks}
+\newpage\section{Cloud-Assisted P2P Networks}\label{sec:ca-networks}
 
@@
 
-\section{Data Redundancy}\label{sec:dataredundancy}
+\newpage\section{Data Redundancy}\label{sec:dataredundancy}
 In DFS, not losing data is of the utmost importance.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Solution Proposal}\label{sec:proposal}
+\newpage\section{Solution Proposal}\label{sec:proposal}
 In this section, we introduce a prototype architecture that targets the use case that inspired to explore the theme of this thesis. We explain it by decomposition.
@@
 \subsection{Use Case}
-Fundamentally, the idea is that a company, Hives, with dedicated web servers called Hiveminds, provides a platform for registered users to sell storage belonging to their personal computers directly to other registered users.
+Fundamentally, the idea is that a company, Hives, with dedicated web servers called Hiveminds, provides a platform for users to sell storage belonging to their personal computers directly to other users.
 Clients pay a monthly fee directly to Hives based on how much storage space they need.
-Hiveminds then put together a set of sellers, which constitute a Hive, that serves adequate storage space to the clients.
-Sellers are selected according to the reputation they earn over time based on their reliability, with a bias towards more reputable sellers.
+Hiveminds then puts together a set of sellers, which constitute a Hive, that serves adequate storage space to the clients.
+Sellers are selected according to the reputation they earn over time, based on their reliability, with a bias towards more reputable sellers.
 The collected fee is distributed proportionately among Hive members according to their overall space contribution and up time, calculated weekly, for example.
 The company's revenue comes from keeping a small percentage of the monthly fees.
-If the client requires continuous availability, rather than durability, he may pay a premium, in which case, inspired on the related work mentioned in section \ref{sec:ca-networks}, cloud storage may be used to help the Hive in meeting the required service guarantees.
-Hive revenue shares are reduced if their reliability degrades too much.
+If the client requires continuous availability, rather than durability, he may pay a premium, in which case, inspired on the related work mentioned in section \ref{sec:ca-networks}, cloud storage may be used to help the Hive in meeting the required service.
+Seller shares are reduced if cloud storage is used at any point.
 
@@
 The Hivemind servers are the backbone of the system; they have to handle registration, login, subscription and payment procedures.
-When a client wishes to subscribe to a monthly plan, he contacts one Hivemind server at random by accessing his desktop application, and the Hivemind then sets a \textit{usedSpace} variable indicating how much data the Client uploaded into the system to zero.
+When a client wishes to subscribe to a monthly plan, he contacts one Hivemind by accessing his desktop application, and the Hivemind then sets a \textit{usedSpace} variable indicating how much data the Client uploaded into the system to zero.
 This variable is updated according to the Client's future requests.
@@
 In a POST operation, the Client first obtains permission to upload from the Hivemind; after that, he divides his file into parts, applies an erasure coding algorithm over each them, and uploads the encoded fragments to the workers at random.
-PUT updates existing files and operate similarly to POST, but whenever possible, try to send the new version to the Hive that contains the previous \textit{versionId}.
+PUT updates existing files and operate similarly to POST, but whenever possible, tries to send the new version to the Hive that contains the previous \textit{versionId}.
 
@@
 Workers will then redistribute the fragment using simple routing rules.
-This approach does not sacrifice fast lookup operations due to the structured nature of the system using a Hivemind that behaves somewhat like a BitTorrent tracker.
+This approach does not sacrifice fast lookup operations due to the structured nature of the system using a Hivemind that behaves somewhat like a BitTorrent\cite{bittorrent} tracker.
 The never-ending fragments exchange between peers also facilitates the timely detection of non-transient faults in the Hive, which helps to ensure file durability.
 Workers piggyback aggregate information to the fragments, including but not limited to, how many parts of a file they have, allowing us to implement a hybrid between a reactive and a proactive system concerning replica control, in hopes of minimizing bandwidth burst on top of already high bandwidth requirements.
-Fragment exchange may eventually be replaced with ping-like messages to minimize bandwidth cost further, hence the importance of having a fast mixing rate transition matrix.
+Fragment exchange may eventually be replaced with ping-like messages to minimize bandwidth cost further, hence the importance of having a fast mixing transition matrix.
 When a Worker suspects that some other Worker on a Hive is faulty for too long, he first contacts a dedicated server, called Keeper, who decides the new Hive membership.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Evaluation Methodology}\label{sec:methodology}
+\newpage\section{Evaluation Methodology}\label{sec:methodology}
 We will first evaluate the performance of different transition matrix generating algorithms by comparing their theoretical fast mixing property against the number of discrete time steps the matrices take to converge on a simple, cycle-based, locally ran, custom-built simulator.
@@
 \section{Work Schedule}\label{sec:workschedule}
+\begin{itemize}
+    \item January 6 to January 12: Implement PSG on basic custom made simulator;
+    \item January 13 to January 19: Perform preliminary tests;
+    \item January 20 to January 26: Improve fast mixing property;
+    \item January 27 to February 2: Write a paper describing initial results;
+    \item February 3 to February 9: Implement erasure coding algorithm
+    \item February 10 to February 16: Holiday Period
+    \item February 17 to February 23: Evaluate and improve erasure coding algorithm
+    \item February 24 to March 29: Attempt to port solution to PeerfactSim.KOM
+    \item March 30 1 to April 12: Preliminary tests and possible bug fixes
+    \item April 13 to April 26: Perform the extensive experimental evaluation
+    \item April 27 to June 15: Write draft for MSc dissertation
+    \item June 16 to August 30: Attempt to deploy proof of concept application
+    \item August 31 to September 27: Finish writing and deliver MSc dissertation
+\end{itemize}
 %%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%
+\newpage
 \bibliographystyle{splncs04}

### 2020-10-08 hive/app/scripts/python/plotting/convergence_analysis.py
Save images as PDF

@@ -57,8 +57,7 @@ def plotvalues(convergence_times_list, directory, state):
     bb.y1 -= 0.15  # yOffset
     leg.set_bbox_to_anchor(bb, transform=ax.transAxes)
     fig.tight_layout()
-    # plt.show()
-    plt.savefig("{}-{}-{}".format("convergence_sets", directory, state))
+    plt.savefig(f"convergence_sets-{directory}-{state}", bbox_inches="tight", format="pdf")
 
 
 def process_file(filepath, convergence_times_list):
@@ -196,7 +195,7 @@ def instantaneous_convergence_plot(
     plt.ylim(0, len(outfiles_view) + 10)
 
     figure_name = f"{plots_directory}/ic_plot_N{ns}O{opt}D{errs}L{ll}"
-    plt.savefig(figure_name, bbox_inches="tight")
+    plt.savefig(figure_name, bbox_inches="tight", format="pdf")
 
 
 def boxplot_first_convergence(outfiles_view):
@@ -231,7 +230,7 @@ def boxplot_first_convergence(outfiles_view):
                fontproperties=cfg.fp_axis_labels)
 
     figure_name = f"{plots_directory}/fc_boxplot_{ns}O{opt}D{errs}L{ll}"
-    plt.savefig(figure_name, bbox_inches="tight")
+    plt.savefig(figure_name, bbox_inches="tight", format="pdf")
 
 
 if __name__ == "__main__":

### 2020-10-08 hive/app/scripts/python/plotting/mean_vs_cumsum-mean.py

@@ -51,7 +51,7 @@ def plot_values(tkey, global_mean, global_cs_mean, terminations):
 
     # plt.show()
     figure_path = os.path.join(plots_directory, f'{figure_name}_{tkey}')
-    plt.savefig(figure_path)
+    plt.savefig(figure_path,  bbox_inches="tight", format="pdf")
 
 
 def process_file(key, outfile_json):

### 2020-10-08 hive/app/scripts/python/plotting/mixing_rate_sample_analysis.py

@@ -124,7 +124,7 @@ def __create_box_plot__(
 
     src = pl.Path(file_name).stem
     fig_name = f"{__MIXING_RATE_PLOTS_HOME__}/bp_f{src}-n{skey}_s{slen}"
-    plt.savefig(fig_name, bbox_inches='tight')
+    plt.savefig(fig_name, bbox_inches="tight", format="pdf")
 # endregion
 
 
@@ -221,7 +221,7 @@ def __create_pie_chart__(
 
     src = pl.Path(file_name).stem
     fig_name = f"{__MIXING_RATE_PLOTS_HOME__}/pc_f{src}-n{skey}_s{slen}"
-    plt.savefig(fig_name, bbox_inches='tight')
+    plt.savefig(fig_name, bbox_inches="tight", format="pdf")
 # endregion
 
 

### 2020-10-09 hive/app/scripts/python/plotting/convergence_analysis.py
Renamed fig_name to figname, and added .pdf extension to all of them.

@@ -55,9 +55,11 @@ def plotvalues(convergence_times_list, directory, state):
     bb = leg.get_bbox_to_anchor().inverse_transformed(ax.transAxes)
     bb.y0 -= 0.15  # yOffset
     bb.y1 -= 0.15  # yOffset
+
     leg.set_bbox_to_anchor(bb, transform=ax.transAxes)
-    fig.tight_layout()
-    plt.savefig(f"convergence_sets-{directory}-{state}", bbox_inches="tight", format="pdf")
+
+    figname = f"convergence_sets-{directory}-{state}.pdf"
+    plt.savefig(figname, bbox_inches="tight", format="pdf")
 
 
 def process_file(filepath, convergence_times_list):
@@ -194,8 +196,8 @@ def instantaneous_convergence_plot(
 
     plt.ylim(0, len(outfiles_view) + 10)
 
-    figure_name = f"{plots_directory}/ic_plot_N{ns}O{opt}D{errs}L{ll}"
-    plt.savefig(figure_name, bbox_inches="tight", format="pdf")
+    figname = f"{plots_directory}/ic_plot_N{ns}O{opt}D{errs}L{ll}.pdf"
+    plt.savefig(figname, bbox_inches="tight", format="pdf")
 
 
 def boxplot_first_convergence(outfiles_view):

### 2020-10-09 hive/app/scripts/python/plotting/mixing_rate_sample_analysis.py

@@ -109,9 +109,9 @@ def __create_box_plot__(
                labels=func_labels,
                rotation=45)
 
-    plt.title(f"Algorithm performance comparison for networks of size {skey}",
-              pad=cfg.title_pad,
-              fontproperties=cfg.fp_title)
+    # plt.title(f"Algorithm performance comparison for networks of size {skey}",
+    #           pad=cfg.title_pad,
+    #           fontproperties=cfg.fp_title)
     plt.xlabel("generating function",
                labelpad=cfg.labels_pad,
                fontproperties=cfg.fp_axis_labels)
@@ -123,8 +123,8 @@ def __create_box_plot__(
     plt.ylim(0.1, 1.1)
 
     src = pl.Path(file_name).stem
-    fig_name = f"{__MIXING_RATE_PLOTS_HOME__}/bp_f{src}-n{skey}_s{slen}"
-    plt.savefig(fig_name, bbox_inches="tight", format="pdf")
+    figname = f"{__MIXING_RATE_PLOTS_HOME__}/bp_f{src}-n{skey}_s{slen}.pdf"
+    plt.savefig(figname, bbox_inches="tight", format="pdf")
 # endregion
 
 
@@ -192,10 +192,10 @@ def __create_pie_chart__(
 
     fig1, ax = plt.subplots()
 
-    plt.title(f"Algorithm selection frequency for networks of size {skey}",
-              x=0.57, y=1,
-              pad=cfg.title_pad,
-              fontproperties=cfg.fp_title)
+    # plt.title(f"Algorithm selection frequency for networks of size {skey}",
+    #           x=0.57, y=1,
+    #           pad=cfg.title_pad,
+    #           fontproperties=cfg.fp_title)
 
     wedges, texts, autotexts = ax.pie(
         wins,
@@ -220,8 +220,8 @@ def __create_pie_chart__(
     ax.axis('equal')
 
     src = pl.Path(file_name).stem
-    fig_name = f"{__MIXING_RATE_PLOTS_HOME__}/pc_f{src}-n{skey}_s{slen}"
-    plt.savefig(fig_name, bbox_inches="tight", format="pdf")
+    figname = f"{__MIXING_RATE_PLOTS_HOME__}/pc_f{src}-n{skey}_s{slen}.pdf"
+    plt.savefig(figname, bbox_inches="tight", format="pdf")
 # endregion
 
 

### 2020-10-09 hive/app/scripts/python/plotting/mixing_rate_sample_analysis.py
Better name and final mixing rate plots in PDF

@@ -123,7 +123,7 @@ def __create_box_plot__(
     plt.ylim(0.1, 1.1)
 
     src = pl.Path(file_name).stem
-    figname = f"{__MIXING_RATE_PLOTS_HOME__}/bp_f{src}-n{skey}_s{slen}.pdf"
+    figname = f"{__MIXING_RATE_PLOTS_HOME__}/mr_s{skey}bp.pdf"
     plt.savefig(figname, bbox_inches="tight", format="pdf")
 # endregion
 
@@ -220,7 +220,7 @@ def __create_pie_chart__(
     ax.axis('equal')
 
     src = pl.Path(file_name).stem
-    figname = f"{__MIXING_RATE_PLOTS_HOME__}/pc_f{src}-n{skey}_s{slen}.pdf"
+    figname = f"{__MIXING_RATE_PLOTS_HOME__}/mr_s{skey}pc.pdf"
     plt.savefig(figname, bbox_inches="tight", format="pdf")
 # endregion
 

### 2020-10-09 hive/app/scripts/python/plotting/_matplotlib_configs.py
Update _matplotlib_configs.py

@@ -11,7 +11,7 @@ legends_pad = 12
 
 fp_title = FontProperties(
     family='sans-serif',
-    weight='bold',
+    weight='semibold',
     size='large'
 )
 
@@ -23,8 +23,8 @@ fp_subtitle = FontProperties(
 fp_axis_labels = FontProperties(
     family='sans-serif',
     style='italic',
-    weight='semibold',
-    size='medium'
+    weight='medium',
+    size='large'
 )
 
 outlyer_shape = {

### 2020-10-09 hive/app/scripts/python/plotting/_matplotlib_configs.py
Fiddling with font properties for better latex visibility at 100% magnification

@@ -11,20 +11,21 @@ legends_pad = 12
 
 fp_title = FontProperties(
     family='sans-serif',
-    weight='semibold',
-    size='large'
+    weight='bold',
+    size='x-large'
 )
 
 fp_subtitle = FontProperties(
     family='sans-serif',
-    size='small'
+    size='large',
+    weight='demibold'
 )
 
 fp_axis_labels = FontProperties(
     family='sans-serif',
     style='italic',
-    weight='medium',
-    size='large'
+    weight='semibold',
+    size='x-large'
 )
 
 outlyer_shape = {

### 2020-10-09 hive/app/scripts/python/plotting/mixing_rate_sample_analysis.py

@@ -105,9 +105,9 @@ def __create_box_plot__(
                 flierprops=outlyer_shape,
                 widths=1,
                 notch=True)
-    plt.xticks(ticks=range(0, func_count * 2, 2),
-               labels=func_labels,
-               rotation=45)
+    plt.xticks(ticks=range(0, func_count * 2, 2), labels=func_labels,
+               rotation=45, fontsize="large", fontweight="semibold")
+    plt.yticks(fontsize="large", fontweight="semibold")
 
     # plt.title(f"Algorithm performance comparison for networks of size {skey}",
     #           pad=cfg.title_pad,

### 2020-10-09 hive/app/scripts/python/plotting/_matplotlib_configs.py
Better titles

@@ -12,7 +12,7 @@ legends_pad = 12
 fp_title = FontProperties(
     family='sans-serif',
     weight='bold',
-    size='x-large'
+    size='large'
 )
 
 fp_subtitle = FontProperties(

### 2020-10-09 hive/app/scripts/python/plotting/mixing_rate_sample_analysis.py

@@ -109,9 +109,9 @@ def __create_box_plot__(
                rotation=45, fontsize="large", fontweight="semibold")
     plt.yticks(fontsize="large", fontweight="semibold")
 
-    # plt.title(f"Algorithm performance comparison for networks of size {skey}",
-    #           pad=cfg.title_pad,
-    #           fontproperties=cfg.fp_title)
+    plt.title(f"Generating functions' SLEM on clusters of size {skey}",
+              pad=cfg.title_pad,
+              fontproperties=cfg.fp_title)
     plt.xlabel("generating function",
                labelpad=cfg.labels_pad,
                fontproperties=cfg.fp_axis_labels)
@@ -192,10 +192,10 @@ def __create_pie_chart__(
 
     fig1, ax = plt.subplots()
 
-    # plt.title(f"Algorithm selection frequency for networks of size {skey}",
-    #           x=0.57, y=1,
-    #           pad=cfg.title_pad,
-    #           fontproperties=cfg.fp_title)
+    plt.title(f"Generating functions' selection frequency for networks of size {skey}",
+              x=0.57, y=1,
+              pad=cfg.title_pad,
+              fontproperties=cfg.fp_title)
 
     wedges, texts, autotexts = ax.pie(
         wins,

### 2020-10-09 hive/app/scripts/python/plotting/mixing_rate_sample_analysis.py
beter opts and ax name

@@ -5,13 +5,13 @@ import json
 import math
 import os
 import sys
+from typing import OrderedDict, List, Any, Dict
 
-import numpy as np
-import pathlib as pl
 import matplotlib.pyplot as plt
+import numpy as np
+
 import _matplotlib_configs as cfg
 
-from typing import OrderedDict, List, Any, Dict
 _SizeResultsDict: OrderedDict[str, List[float]]
 _ResultsDict: OrderedDict[str, _SizeResultsDict]
 
@@ -112,17 +112,16 @@ def __create_box_plot__(
     plt.title(f"Generating functions' SLEM on clusters of size {skey}",
               pad=cfg.title_pad,
               fontproperties=cfg.fp_title)
-    plt.xlabel("generating function",
+    plt.xlabel("function name abbreviation",
                labelpad=cfg.labels_pad,
                fontproperties=cfg.fp_axis_labels)
-    plt.ylabel("mixing rate",
+    plt.ylabel("slem",
                labelpad=cfg.labels_pad,
                fontproperties=cfg.fp_axis_labels)
 
     plt.xlim(-2, func_count * 2)
     plt.ylim(0.1, 1.1)
 
-    src = pl.Path(file_name).stem
     figname = f"{__MIXING_RATE_PLOTS_HOME__}/mr_s{skey}bp.pdf"
     plt.savefig(figname, bbox_inches="tight", format="pdf")
 # endregion
@@ -219,7 +218,6 @@ def __create_pie_chart__(
     leg._legend_box.sep = cfg.legends_pad
     ax.axis('equal')
 
-    src = pl.Path(file_name).stem
     figname = f"{__MIXING_RATE_PLOTS_HOME__}/mr_s{skey}pc.pdf"
     plt.savefig(figname, bbox_inches="tight", format="pdf")
 # endregion

### 2020-10-09 hive/app/scripts/python/plotting/hives_vs_peersim_performance.py
better plots

@@ -45,17 +45,18 @@ if __name__ == "__main__":
     ax.legend()
 
     # Use supertitle as title and title as subtitle
-    plt.suptitle("Simulators' performance comparison",
-                 fontproperties=cfg.fp_title,
-                 x=0.56, y=0.999)
-    plt.title("30 cycles of Newscast Shuffling and AverageFunction aggregation",
-              fontproperties=cfg.fp_subtitle)
-    plt.xlabel("Number of network nodes",
-               labelpad=cfg.labels_pad,
-               fontproperties=cfg.fp_axis_labels)
-    plt.ylabel("Execution time in seconds",
-               labelpad=cfg.labels_pad,
-               fontproperties=cfg.fp_axis_labels)
+    # plt.suptitle("Simulators' performance comparison, Hives vs. PeerSim",
+    #              fontproperties=cfg.fp_title,
+    #              x=0.56, y=0.999)
+    # plt.title("30 cycles of Newscast Shuffling and AverageFunction aggregation",
+    #           fontproperties=cfg.fp_subtitle)
+    plt.title("Simulators' performance comparison, Hives vs. PeerSim",
+              pad=cfg.title_pad, fontproperties=cfg.fp_title)
+
+    plt.xlabel("number of network nodes",
+               labelpad=cfg.labels_pad, fontproperties=cfg.fp_axis_labels)
+    plt.ylabel("time in seconds",
+               labelpad=cfg.labels_pad, fontproperties=cfg.fp_axis_labels)
 
     autolabel(rects1)
     autolabel(rects2)

### 2020-10-09 hive/app/scripts/python/plotting/mixing_rate_sample_analysis.py

@@ -110,14 +110,11 @@ def __create_box_plot__(
     plt.yticks(fontsize="large", fontweight="semibold")
 
     plt.title(f"Generating functions' SLEM on clusters of size {skey}",
-              pad=cfg.title_pad,
-              fontproperties=cfg.fp_title)
+              pad=cfg.title_pad, fontproperties=cfg.fp_title)
     plt.xlabel("function name abbreviation",
-               labelpad=cfg.labels_pad,
-               fontproperties=cfg.fp_axis_labels)
+               labelpad=cfg.labels_pad, fontproperties=cfg.fp_axis_labels)
     plt.ylabel("slem",
-               labelpad=cfg.labels_pad,
-               fontproperties=cfg.fp_axis_labels)
+               labelpad=cfg.labels_pad, fontproperties=cfg.fp_axis_labels)
 
     plt.xlim(-2, func_count * 2)
     plt.ylim(0.1, 1.1)

### 2020-10-09 hive/app/scripts/python/plotting/hives_vs_peersim_performance.py
better plots for simulator compare

@@ -5,6 +5,8 @@ Experiments differ in the number of network nodes that exist in the
 simulation."""
 from __future__ import annotations
 
+import os
+
 import numpy
 
 import matplotlib.pyplot as plt
@@ -15,11 +17,12 @@ def autolabel(rects):
     """Attach a text label above each bar in *rects*, displaying its height."""
     for rect in rects:
         height = rect.get_height()
-        ax.annotate("{}".format(height),
+        ax.annotate(f"{height}",
                     xy=(rect.get_x() + rect.get_width() / 2, height),
                     xytext=(0, 3),  # 3 points vertical offset
                     textcoords="offset points",
-                    ha="center", va="bottom")
+                    ha="center", va="bottom",
+                    fontsize="large", fontweight="semibold", color="dimgrey")
 
 
 if __name__ == "__main__":
@@ -55,8 +58,12 @@ if __name__ == "__main__":
 
     plt.xlabel("number of network nodes",
                labelpad=cfg.labels_pad, fontproperties=cfg.fp_axis_labels)
+
     plt.ylabel("time in seconds",
                labelpad=cfg.labels_pad, fontproperties=cfg.fp_axis_labels)
+
+    plt.xticks(fontsize="large", fontweight="semibold")
+    plt.yticks(fontsize="large", fontweight="semibold")
 
     autolabel(rects1)
     autolabel(rects2)
@@ -65,4 +72,8 @@ if __name__ == "__main__":
 
     fig.tight_layout()
 
-    plt.show()
+    figdir = os.path.abspath(os.path.join(
+        os.getcwd(), '..', '..', '..', 'static', 'outfiles', 'simulation_plots')
+    )
+    figname = f"{figdir}/simulators_execution_times.pdf"
+    plt.savefig(figname, bbox_inches="tight", format="pdf")

### 2020-10-11 hive/app/scripts/python/plotting/_matplotlib_configs.py
Even larger titles, ax labels and ticks

@@ -12,12 +12,12 @@ legends_pad = 12
 fp_title = FontProperties(
     family='sans-serif',
     weight='bold',
-    size='large'
+    size='xx-large'
 )
 
 fp_subtitle = FontProperties(
     family='sans-serif',
-    size='large',
+    size='xx-large',
     weight='demibold'
 )
 
@@ -25,7 +25,7 @@ fp_axis_labels = FontProperties(
     family='sans-serif',
     style='italic',
     weight='semibold',
-    size='x-large'
+    size='xx-large'
 )
 
 outlyer_shape = {

### 2020-10-11 hive/app/scripts/python/plotting/hives_vs_peersim_performance.py

@@ -22,7 +22,7 @@ def autolabel(rects):
                     xytext=(0, 3),  # 3 points vertical offset
                     textcoords="offset points",
                     ha="center", va="bottom",
-                    fontsize="large", fontweight="semibold", color="dimgrey")
+                    fontsize="x-large", fontweight="semibold", color="dimgrey")
 
 
 if __name__ == "__main__":
@@ -62,8 +62,8 @@ if __name__ == "__main__":
     plt.ylabel("time in seconds",
                labelpad=cfg.labels_pad, fontproperties=cfg.fp_axis_labels)
 
-    plt.xticks(fontsize="large", fontweight="semibold")
-    plt.yticks(fontsize="large", fontweight="semibold")
+    plt.xticks(fontsize="x-large", fontweight="semibold")
+    plt.yticks(fontsize="x-large", fontweight="semibold")
 
     autolabel(rects1)
     autolabel(rects2)

### 2020-10-11 hive/app/scripts/python/plotting/mixing_rate_sample_analysis.py

@@ -106,8 +106,8 @@ def __create_box_plot__(
                 widths=1,
                 notch=True)
     plt.xticks(ticks=range(0, func_count * 2, 2), labels=func_labels,
-               rotation=45, fontsize="large", fontweight="semibold")
-    plt.yticks(fontsize="large", fontweight="semibold")
+               rotation=45, fontsize="x-large", fontweight="semibold")
+    plt.yticks(fontsize="x-large", fontweight="semibold")
 
     plt.title(f"Generating functions' SLEM on clusters of size {skey}",
               pad=cfg.title_pad, fontproperties=cfg.fp_title)

### 2020-10-11 hive/app/scripts/python/plotting/hives_vs_peersim_performance.py
Update hives_vs_peersim_performance.py

@@ -22,7 +22,7 @@ def autolabel(rects):
                     xytext=(0, 3),  # 3 points vertical offset
                     textcoords="offset points",
                     ha="center", va="bottom",
-                    fontsize="x-large", fontweight="semibold", color="dimgrey")
+                    fontsize="large", fontweight="semibold", color="dimgrey")
 
 
 if __name__ == "__main__":

### 2020-10-21 hive/app/domain/cluster_groups.py
Fix a bug that broadcasted the same matrix twice

@@ -786,18 +786,15 @@ class SGCluster(Cluster):
         ``SGCluster`` will change, thus, more opportunities to perform a
         correct swarm guidance behavior will be possible.
         """
-        tries = 1
+        tries = 0
         result: pd.DataFrame = pd.DataFrame()
         while tries <= 5:
-            print(f"Creating new transition matrix... try #{tries}.")
+            print(f"Creating new transition matrix... try #{tries + 1}.")
+            tries += 1
             result = self.new_transition_matrix()
             if self._validate_transition_matrix(result, self.v_):
-                self.broadcast_transition_matrix(result)
                 break
-            tries += 1
-        # Only tries to generate a valid matrix up to three times,
-        # then resumes with the last generated matrix even if it never
-        # converges.
+        # Only tries to create a valid matrix up to five times before proceeding
         self.broadcast_transition_matrix(result)
 
     # noinspection PyIncorrectDocstring

### 2020-10-21 hive/app/scripts/python/plotting/convergence_analysis_v2_parts.py
now testing boxplot_node_degree

@@ -29,6 +29,10 @@ def setup_sources(source_patterns: List[str]) -> Tuple[Dict[str, Any], List[str]
     return sources_files, source_keys
 
 
+def __get_whole_frac__(num: float) -> Tuple[int, int]:
+    return int(num), int(str(num)[(len(str(int(num)))+1):])
+
+
 def __set_box_color__(bp: Any, color: str) -> None:
     """Changes the colors of a boxplot.
 
@@ -86,7 +90,7 @@ def __create_boxplot__(data_dict: Dict[str, Any],
     plt.suptitle(suptitle, fontproperties=cfg.fp_title)
     plt.xlabel(xlabel, labelpad=cfg.labels_pad, fontproperties=cfg.fp_axis_labels)
     plt.ylabel(ylabel, labelpad=cfg.labels_pad, fontproperties=cfg.fp_axis_labels)
-    plt.xticks(rotation=45, fontsize="x-large", fontweight="semibold")
+    plt.xticks(rotation=75, fontsize="x-large", fontweight="semibold")
     plt.yticks(fontsize="x-large", fontweight="semibold")
     if savefig:
         plt.savefig(f"{plots_directory}/{figname}.{figext}", format=figext, bbox_inches="tight")
@@ -95,6 +99,7 @@ def __create_boxplot__(data_dict: Dict[str, Any],
 
 def __create_double_boxplot__(left_data, right_data,
                               suptitle: str, xlabel: str, ylabel: str,
+                              labels: List[str],
                               figname: str, figext: str = "png",
                               left_color: Optional[str] = None,
                               right_color: Optional[str] = None,
@@ -107,21 +112,25 @@ def __create_double_boxplot__(left_data, right_data,
     bpr = plt.boxplot(right_data,  sym='', whis=0.75, widths=0.7, notch=True, positions=np.array(range(len(right_data))) * 2.0 + 0.4)
 
     if left_color:
-        __set_box_color__(bpl, "#55A868")
+        __set_box_color__(bpl, left_color)
         if left_label:
-            plt.plot([], c="#55A868", label=left_label)
+            plt.plot([], c=left_color, label=left_label)
             plt.legend(frameon=False, loc="best", prop=cfg.fp_axis_legend)
     if right_color:
-        __set_box_color__(bpr, "#C44E52")
+        __set_box_color__(bpr, right_color)
         if right_label:
-            plt.plot([], c="#C44E52", label=right_label)
[…]

### 2020-10-21 hive/app/domain/cluster_groups.py
now log in and out-degrees without float right most zero bug

@@ -762,12 +762,12 @@ class SGCluster(Cluster):
             queries and matrix calculations. For simplicity of implementation
             each cluster only manages one file.
         """
-        nodes_degrees: Dict[str, float] = {}
+        nodes_degrees: Dict[str, str] = {}
         out_degrees: pd.Series = m.apply(np.count_nonzero, axis=0)  # columns
         in_degrees: pd.Series = m.apply(np.count_nonzero, axis=1)  # rows
         for node in self.members.values():
             nid = node.id
-            nodes_degrees[nid] = float(f"{in_degrees[nid]}.{out_degrees[nid]}")
+            nodes_degrees[nid] = f"{in_degrees[nid]}i#o{out_degrees[nid]}"
             transition_vector: pd.DataFrame = m.loc[:, nid]
             node.set_file_routing(self.file.name, transition_vector)
         self.file.logger.log_matrices_degrees(nodes_degrees)

### 2020-10-21 hive/app/domain/helpers/smart_dataclasses.py

@@ -483,7 +483,7 @@ class LoggingData:
         self.delay_replication: List[float] = [0.0] * max_epochs_plus_one
         self.delay_suspects_detection: Dict[str, int] = {}
         self.initial_spread = ""
-        self.matrices_nodes_degrees: List[Dict[str, float]] = []
+        self.matrices_nodes_degrees: List[Dict[str, str]] = []
         self.off_node_count: List[int] = [0] * max_epochs
         self.topologies_goal_achieved: List[bool] = []
         self.topologies_goal_distance: List[float] = []
@@ -570,7 +570,7 @@ class LoggingData:
         self.topologies_goal_achieved.append(achieved_goal)
         self.topologies_goal_distance.append(distance_magnitude)
 
-    def log_matrices_degrees(self, nodes_degrees: Dict[str, float]):
+    def log_matrices_degrees(self, nodes_degrees: Dict[str, str]):
         """Logs the degree of all nodes in a Markov Matrix overlay, at the
         time of its creation, before any faults on the overlay occurs.
 
@@ -578,7 +578,7 @@ class LoggingData:
             nodes_degrees:
                 A dictionary mapping the :py:attr:`node identifiers
                 <app.domain.network_nodes.Node.id>` to their ``in-degree``
-                and ``out-degree`` concatenated as a float.
+                and ``out-degree`` seperated by the delimiter ``i#o``.
         """
         self.matrices_nodes_degrees.append(nodes_degrees)
 

### 2020-10-21 hive/app/domain/helpers/smart_dataclasses.py
Create NUMBER_PARTS along with READ_SIZE so user can choose to use one or the other.

@@ -125,6 +125,7 @@ class FileData:
             "hive_id": cluster.id,
             "file_name": self.name,
             "read_size": es.READ_SIZE,
+            "number_of_parts": es.NUMBER_PARTS,
             "critical_size_threshold": cluster.critical_size,
             "sufficient_size_threshold": cluster.sufficient_size,
             "original_size": cluster.original_size,

### 2020-10-21 hive/app/environment_settings.py

@@ -40,7 +40,6 @@ DEBUG: bool = False
 """Indicates if some debug related actions or prints to the terminal should 
 be performed."""
 
-
 READ_SIZE: int = 131072
 """Defines the raw size of each file block before it's wrapped in a 
 :py:class:`~app.domain.helpers.smart_dataclasses.FileBlockData` instance 
@@ -49,6 +48,25 @@ object.
 Some possible values include { 32KB = 32768B; 128KB = 131072B; 512KB = 524288B; 
 1MB = 1048576B; 20MB = 20971520B }.
 """
+
+
+def set_read_size(n: int) -> None:
+    """Changes READ_SIZE constant value at run time to the given n bytes."""
+    global READ_SIZE
+    READ_SIZE = n
+
+
+NUMBER_PARTS: int = 333
+"""Defines into how many 
+:py:class:`~app.domain.helpers.smart_dataclasses.FileBlockData` instances a file
+is divided into. Either use this or :py:const:`READ_SIZE` but not both."""
+
+
+def set_number_parts(n: int) -> None:
+    """Changes NUMBER_PARTS constant value at run time."""
+    global NUMBER_PARTS
+    NUMBER_PARTS = n
+
 
 NEWSCAST_CACHE_SIZE: int = 20
 """The maximum amount of neighbors a :py:attr:`NewscastNode view 

### 2020-10-21 hive/app/domain/helpers/smart_dataclasses.py
Fix docstring

@@ -228,7 +228,7 @@ class FileBlockData:
         self.sha256: str = crypto.sha256(self.data)
 
     # region Simulation Interface
-    def set_replication_epoch(self, epoch: int) -> int:
+    def set_replication_epoch(self, epoch: int) -> float:
         """Sets the epoch in which replication levels should be restored.
 
         This method tries to assign a new epoch, in the future, at which

### 2020-10-21 hive/app/simfile_generator.py
Simfile_generator only simulates one file at the time now.

@@ -232,28 +232,23 @@ def _init_persisting_dict() -> Dict[str, Any]:
         "~/cluster/app/static/shared/shared_backups"
     )
 
-    add_file: bool = True
-    while add_file:
-        file_name = _input_filename(
-            "Name the file (with extension) you wish to simulate persistence of: ")
+    file_name = _input_filename(
+        "Name the file (with extension) you wish to simulate persistence of: ")
 
-        options_message = ("\nSelect how files blocks are spread across "
-                           "clusters at the start of the simulation: {\n"
-                           "   u: uniform distribution among network nodes,\n"
-                           "   i: ideal distribution, e.g., near a steady-state vector, \n"
-                           "   a: all replicas given to N different nodes,\n"
-                           "   o: each network node receives one random replica\n"
-                           "}: ")
-        options_list = ["u", "U", "i", "I", "a", "A", "o", "O"]
-        option_choice = _input_character_option(options_message, options_list)
+    options_message = ("\nSelect how files blocks are spread across "
+                       "clusters at the start of the simulation: {\n"
+                       "   u: uniform distribution among network nodes,\n"
+                       "   i: ideal distribution, e.g., near a steady-state vector, \n"
+                       "   a: all replicas given to N different nodes,\n"
+                       "   o: each network node receives one random replica\n"
+                       "}: ")
+    options_list = ["u", "U", "i", "I", "a", "A", "o", "O"]
+    option_choice = _input_character_option(options_message, options_list)
 
-        persisting[file_name] = {}
-        persisting[file_name]["spread"] = option_choice.lower()
-        persisting[file_name]["cluster_size"] = _input_bounded_integer(
-            "\nNumber of nodes that should be sharing the next file: ")
-
-        add_file = _in_yes_no(
-            "\nSimulate persistence of another file in simulation?")
+    persisting[file_name] = {}
+    persisting[file_name]["spread"] = option_choice.lower()
+    persisting[file_name]["cluster_size"] = _input_bounded_integer(
+        "\nNumber of nodes that should be sharing the next file: ")
 
     return persisting
 # endregion

### 2020-10-21 hive/app/domain/helpers/smart_dataclasses.py
Change READ_SIZE to BLOCKS_SIZE and NUMBER_PARTS TO BLOCKS_COUNT

@@ -124,8 +124,8 @@ class FileData:
             "simfile_name": origin,
             "hive_id": cluster.id,
             "file_name": self.name,
-            "read_size": es.READ_SIZE,
-            "number_of_parts": es.NUMBER_PARTS,
+            "blocks_size": es.BLOCKS_SIZE,
+            "blocks_count": es.BLOCKS_COUNT,
             "critical_size_threshold": cluster.critical_size,
             "sufficient_size_threshold": cluster.sufficient_size,
             "original_size": cluster.original_size,

### 2020-10-21 hive/app/environment_settings.py

@@ -40,7 +40,7 @@ DEBUG: bool = False
 """Indicates if some debug related actions or prints to the terminal should 
 be performed."""
 
-READ_SIZE: int = 131072
+BLOCKS_SIZE: int = 131072
 """Defines the raw size of each file block before it's wrapped in a 
 :py:class:`~app.domain.helpers.smart_dataclasses.FileBlockData` instance 
 object. 
@@ -50,22 +50,22 @@ Some possible values include { 32KB = 32768B; 128KB = 131072B; 512KB = 524288B;
 """
 
 
-def set_read_size(n: int) -> None:
-    """Changes READ_SIZE constant value at run time to the given n bytes."""
-    global READ_SIZE
-    READ_SIZE = n
+def set_block_size(n: int) -> None:
+    """Changes BLOCKS_SIZE constant value at run time to the given n bytes."""
+    global BLOCKS_SIZE
+    BLOCKS_SIZE = n
 
 
-NUMBER_PARTS: int = 333
+BLOCKS_COUNT: int = 333
 """Defines into how many 
 :py:class:`~app.domain.helpers.smart_dataclasses.FileBlockData` instances a file
-is divided into. Either use this or :py:const:`READ_SIZE` but not both."""
+is divided into. Either use this or :py:const:`BLOCKS_SIZE` but not both."""
 
 
-def set_number_parts(n: int) -> None:
+def set_blocks_count(n: int) -> None:
     """Changes NUMBER_PARTS constant value at run time."""
-    global NUMBER_PARTS
-    NUMBER_PARTS = n
+    global BLOCKS_COUNT
+    BLOCKS_COUNT = n
 
 
 NEWSCAST_CACHE_SIZE: int = 20

### 2020-10-21 hive/app/environment_settings.py
set_block_size to set_blocks_size

@@ -50,7 +50,7 @@ Some possible values include { 32KB = 32768B; 128KB = 131072B; 512KB = 524288B;
 """
 
 
-def set_block_size(n: int) -> None:
+def set_blocks_size(n: int) -> None:
     """Changes BLOCKS_SIZE constant value at run time to the given n bytes."""
     global BLOCKS_SIZE
     BLOCKS_SIZE = n

### 2020-10-21 hive/app/domain/master_servers.py
Update master_servers.py

@@ -139,9 +139,9 @@ class Master:
                 size = d[fname]['cluster_size']
                 cluster = self._new_cluster_group(cluster_class, size, fname)
 
-                file_size = os.path.getsize(os.path.join(es.SHARED_ROOT, fname))
-                read_size = int(file_size / 333)
-                fblocks[fname] = self._split_files(fname, cluster, read_size)
+                filesize = os.path.getsize(os.path.join(es.SHARED_ROOT, fname))
+                es.set_blocks_size(int(filesize / es.BLOCKS_COUNT))
+                fblocks[fname] = self._split_files(fname, cluster, es.BLOCKS_SIZE)
 
                 # fblocks[fname] = self._split_files(fname, cluster, es.READ_SIZE)
                 
@@ -415,7 +415,11 @@ class HDFSMaster(Master):
                 fspreads[fname] = spread_strategy
                 size = d[fname]['cluster_size']
                 cluster = self._new_cluster_group(cluster_class, size, fname)
-                fblocks[fname] = self._split_files(fname, cluster, 1048576)
+
+                filesize = os.path.getsize(os.path.join(es.SHARED_ROOT, fname))
+                es.set_blocks_size(1 * 1024 * 1024)
+                es.set_blocks_count(int(filesize / es.BLOCKS_SIZE))
+                fblocks[fname] = self._split_files(fname, cluster, es.BLOCKS_SIZE)
 
             # Distribute files before starting simulation
             for cluster in self.cluster_groups.values():

### 2020-10-21 hive/app/domain/cluster_groups.py
Update cluster_groups.py

@@ -794,6 +794,7 @@ class SGCluster(Cluster):
             result = self.new_transition_matrix()
             if self._validate_transition_matrix(result, self.v_):
                 break
+            print(" [x] Invalid matrix.")
         # Only tries to create a valid matrix up to five times before proceeding
         self.broadcast_transition_matrix(result)
 

### 2020-10-21 hive/app/hive_simulation.py
Update hive_simulation.py

@@ -127,7 +127,8 @@ def _parallel_main(threads_count: int,
     """
     with ThreadPoolExecutor(max_workers=threads_count) as executor:
         if sdir:
-            snames = os.listdir(SIMULATION_ROOT)
+            snames = list(filter(
+                lambda x: "scenarios" not in x, os.listdir(SIMULATION_ROOT)))
             for sn in snames:
                 for i in range(1, iters + 1):
                     executor.submit(__start_simulation__, sn, i, epochs)
@@ -157,7 +158,8 @@ def _single_main(
             lasts.
     """
     if sdir:
-        snames = os.listdir(SIMULATION_ROOT)
+        snames = list(filter(
+            lambda x: "scenarios" not in x, os.listdir(SIMULATION_ROOT)))
         for sn in snames:
             for i in range(1, iters + 1):
                 __start_simulation__(sn, i, epochs)

### 2020-10-21 hive/app/hive_simulation.py
Exclude scenario files from hive_simulation.py

@@ -223,14 +223,12 @@ if __name__ == "__main__":
     cluster_class = "SGClusterExt"
     node_class = "SGNodeExt"
 
-    try:
-        short_opts = "df:i:t:e:m:c:n:"
-        long_opts = [
-            "directory", "file=", "iters=", "threading=", "epochs=",
-            "master_server=", "cluster_group=", "network_node="
-        ]
-        options, args = getopt.getopt(sys.argv[1:], short_opts, long_opts)
+    short_opts = "df:i:t:e:m:c:n:"
+    long_opts = ["directory", "file=", "iters=", "threading=", "epochs=",
+                 "master_server=", "cluster_group=", "network_node="]
 
+    try:
+        options, args = getopt.getopt(sys.argv[1:], short_opts, long_opts)
         for options, args in options:
             if options in ("-t", "--threading"):
                 threading = int(str(args).strip())

### 2020-10-21 hive/app/environment_settings.py
Created RESOURCES_ROOT and got rid of if checking on os.makedirs by setting , exist_ok=True

@@ -195,8 +195,10 @@ SIMULATION_ROOT: str = os.path.join(os.getcwd(), 'static', 'simfiles')
 :py:mod:`app.hive_simulation` are located."""
 
 OUTFILE_ROOT: str = os.path.join(os.getcwd(), 'static', 'outfiles')
-"""Path to the folder where simulation output files are located will be 
-stored."""
+"""Path to the folder where simulation output files are located."""
+
+RESOURCES_ROOT: str = os.path.join(os.getcwd(), 'static', 'resources')
+"""Path to the folder where miscellaneous files are located."""
 
 MIXING_RATE_SAMPLE_ROOT: str = os.path.join(OUTFILE_ROOT, 'mixing_rate_samples')
 

### 2020-10-21 hive/app/hive_simulation.py

@@ -57,15 +57,10 @@ from utils.convertions import class_name_to_obj
 def __makedirs__() -> None:
     """Helper method that reates required simulation working directories if
     they do not exist."""
-    if not os.path.exists(SHARED_ROOT):
-        os.makedirs(SHARED_ROOT)
-
-    if not os.path.exists(SIMULATION_ROOT):
-        os.makedirs(SIMULATION_ROOT)
-
-    if not os.path.exists(OUTFILE_ROOT):
-        os.makedirs(OUTFILE_ROOT)
-
+    os.makedirs(SHARED_ROOT, exist_ok=True)
+    os.makedirs(SIMULATION_ROOT, exist_ok=True)
+    os.makedirs(OUTFILE_ROOT, exist_ok=True)
+    os.makedirs(RESOURCES_ROOT, exist_ok=True)
 
 def __can_exec_simfile__(sname: str) -> None:
     """Asserts if simulation can proceed with user specified file.

### 2020-10-21 hive/app/mixing_rate_sampler.py

@@ -49,8 +49,7 @@ def main():
     The execution of the main method results in a JSON file outputed to
     :py:const:`~app.environment_settings.MIXING_RATE_SAMPLE_ROOT` folder.
     """
-    if not os.path.exists(MIXING_RATE_SAMPLE_ROOT):
-        os.makedirs(MIXING_RATE_SAMPLE_ROOT)
+    os.makedirs(MIXING_RATE_SAMPLE_ROOT, exist_ok=True)
 
     MatlabEngineContainer.get_instance()
 

### 2020-10-21 hive/app/sample_scenario_generator.py
Created scenarios file.

@@ -1,0 +1,48 @@
+import os
+import sys
+import ast
+import json
+import getopt
+
+import environment_settings as es
+import domain.helpers.matrices as mm
+
+from typing import Tuple, Any
+
+if __name__ == "__main__":
+    network_sizes: Tuple = (8, 16, 32)
+    samples: int = 100
+
+    short_opts = "n:s:"
+    long_opts = ["network_sizes=", "samples="]
+
+    try:
+        options, args = getopt.getopt(sys.argv[1:], short_opts, long_opts)
+        for options, args in options:
+            if options in ("-n", "--network_size"):
+                network_sizes = ast.literal_eval(str(args).strip())
+            if options in ("-s", "--samples"):
+                samples = int(str(args).strip())
+    except getopt.GetoptError:
+        sys.exit()
+    except ValueError:
+        sys.exit("Execution arguments should have the following data types:\n"
+                 "  --network_size -n (comma seperated list of int)\n"
+                 "  --samples -s (int)\n")
+
+    if not network_sizes or samples < 1:
+        sys.exit("Can't proceed with no samples parameter or empty networks.")
+
+    os.makedirs(es.RESOURCES_ROOT, exist_ok=True)
+
+    scenarios = {str(k): {"vectors": [], "matrices": []} for k in network_sizes}
+    for k in scenarios:
+        network_size = int(k)
+        for i in range(samples):
+            v_ = mm.new_vector(network_size).tolist()
+            m = mm.new_symmetric_connected_matrix(network_size).tolist()
+            scenarios[k]["vectors"].append(v_)
+            scenarios[k]["matrices"].append(m)
+
+    with open(os.path.join(es.RESOURCES_ROOT, "scenarios.json"), "w") as f:
+        json.dump(scenarios, f, indent=4, sort_keys=True, ensure_ascii=False)

### 2020-10-21 hive/app/mixing_rate_sampler.py
Fixed a type hint

@@ -86,7 +86,7 @@ def main():
 
 if __name__ == "__main__":
     samples: int = 30
-    network_sizes: Tuple[int] = (8, 16)
+    network_sizes: Tuple = (8, 16)
     module: Any = "domain.helpers.matrices"
     functions: List[str] = [
         "new_mh_transition_matrix",

### 2020-10-21 hive/app/hive_simulation.py
better imports in hive_simulation.py

@@ -40,10 +40,10 @@ from concurrent.futures.thread import ThreadPoolExecutor
 from typing import Optional
 
 import numpy
+import environment_settings as es
 
 from domain.helpers.matlab_utils import MatlabEngineContainer
-from environment_settings import SIMULATION_ROOT, OUTFILE_ROOT, SHARED_ROOT, \
-    MASTER_SERVERS
+
 
 __err_message__ = ("Invalid arguments. You must specify -f fname or -d, e.g.:\n"
                    "    $ python hive_simulation.py -f simfilename.json\n"
@@ -57,10 +57,11 @@ from utils.convertions import class_name_to_obj
 def __makedirs__() -> None:
     """Helper method that reates required simulation working directories if
     they do not exist."""
-    os.makedirs(SHARED_ROOT, exist_ok=True)
-    os.makedirs(SIMULATION_ROOT, exist_ok=True)
-    os.makedirs(OUTFILE_ROOT, exist_ok=True)
-    os.makedirs(RESOURCES_ROOT, exist_ok=True)
+    os.makedirs(es.SHARED_ROOT, exist_ok=True)
+    os.makedirs(es.SIMULATION_ROOT, exist_ok=True)
+    os.makedirs(es.OUTFILE_ROOT, exist_ok=True)
+    os.makedirs(es.RESOURCES_ROOT, exist_ok=True)
+
 
 def __can_exec_simfile__(sname: str) -> None:
     """Asserts if simulation can proceed with user specified file.
@@ -72,7 +73,7 @@ def __can_exec_simfile__(sname: str) -> None:
             :py:const:`~app.environment_settings.SIMULATION_ROOT` will be
             checked.
     """
-    spath = os.path.join(SIMULATION_ROOT, sname)
+    spath = os.path.join(es.SIMULATION_ROOT, sname)
     if not os.path.exists(spath):
         sys.exit("Specified simulation file does not exist in SIMULATION_ROOT.")
 
@@ -89,7 +90,7 @@ def __start_simulation__(sname: str, sid: int, epochs: int) -> None:
             The number of discrete time steps the simulation lasts.
     """
     master_server = class_name_to_obj(
-        MASTER_SERVERS,
+        es.MASTER_SERVERS,
         master_class,
         [sname, sid, epochs, cluster_class, node_class]
     )
@@ -123,7 +124,7 @@ def _parallel_main(threads_count: int,
     with ThreadPoolExecutor(max_workers=threads_count) as executor:
         if sdir:
[…]

### 2020-10-21 hive/app/domain/cluster_groups.py
SGClusterPerfect uses mm.new_vector

@@ -1003,9 +1003,7 @@ class SGClusterPerfect(SGCluster):
                 A list of floats with which represent how the files should be
                 distributed among network nodes in the long-run.
         """
-        u_ = np.random.random_sample(len(member_ids))
-        u_ /= np.sum(u_)
-
+        u_ = mm.new_vector(len(member_ids))
         self.v_ = pd.DataFrame(data=u_, index=member_ids)
         self.cv_ = pd.DataFrame(data=[0] * len(self.v_), index=member_ids)
         self.avg_ = pd.DataFrame(data=[0] * len(self.v_), index=member_ids)

### 2020-10-21 hive/app/domain/master_servers.py
SGMaster now loads scenarios if env settings DEBUG is set to True

@@ -335,7 +335,41 @@ class Master:
 
 
 class SGMaster(Master):
-    # region Master API
+    SCENARIOS: Dict[str, Dict[str, List]] = {}
+
+    def __init__(self,
+                 simfile_name: str,
+                 sid: int,
+                 epochs: int,
+                 cluster_class: str,
+                 node_class: str) -> None:
+        """Instantiates an Master object.
+
+        Args:
+            simfile_name:
+                A path to the simulation file to be run by the simulator.
+            sid:
+                Identifier that generates unique output file names,
+                thus guaranteeing that different simulation instances do not
+                overwrite previous out files.
+            epochs:
+                The number of discrete time steps the simulation lasts.
+            cluster_class:
+                The name of the class used to instantiate cluster group
+                instances through reflection. See :py:mod:`cluster groups module
+                <app.domain.cluster_groups>`.
+            node_class:
+                The name of the class used to instantiate network node
+                instances through reflection. See :py:mod:`network nodes module
+                <app.domain.network_nodes>`.
+        """
+        if es.DEBUG:
+            scenarios_path = os.path.join(es.RESOURCES_ROOT, "scenarios.json")
+            scenarios = open(scenarios_path, "r")
+            SGMaster.SCENARIOS = json.load(scenarios)
+            scenarios.close()
+        super().__init__(simfile_name, sid, epochs, cluster_class, node_class)
+
     def get_cloud_reference(self) -> str:
         """Use to obtain a reference to 3rd party cloud storage provider
 

### 2020-10-21 hive/app/environment_settings.py

@@ -36,7 +36,7 @@ from typing import List
 from utils.convertions import truncate_float_value
 import numpy
 
-DEBUG: bool = False
+DEBUG: bool = True
 """Indicates if some debug related actions or prints to the terminal should 
 be performed."""
 

### 2020-10-21 hive/app/domain/helpers/smart_dataclasses.py
Update smart_dataclasses.py

@@ -122,7 +122,7 @@ class FileData:
         extras: Dict[str, Any] = {
             "cluster_type": cluster.__class__.__name__,
             "simfile_name": origin,
-            "hive_id": cluster.id,
+            "cluster_id": cluster.id,
             "file_name": self.name,
             "blocks_size": es.BLOCKS_SIZE,
             "blocks_count": es.BLOCKS_COUNT,
@@ -179,7 +179,7 @@ class FileBlockData:
     parameters, e.g., replica control such or file block integrity.
 
     Attributes:
-        hive_id:
+        cluster_id:
             Unique identifier of the cluster that manages the file block.
         name:
             The name of the file the file block belongs to.
@@ -204,12 +204,12 @@ class FileBlockData:
     """
 
     def __init__(
-            self, hive_id: str, name: str, number: int, data: bytes
+            self, cluster_id: str, name: str, number: int, data: bytes
     ) -> None:
         """Creates an instance of `FileBlockData`.
 
         Args:
-            hive_id:
+            cluster_id:
                 Unique identifier of the cluster that manages the file block.
             name:
                 The name of the file the file block belongs to.
@@ -218,7 +218,7 @@ class FileBlockData:
             data:
                 Actual file block data as a sequence of bytes.
         """
-        self.hive_id = hive_id
+        self.cluster_id = cluster_id
         self.name: str = name
         self.number: int = number
         self.id: str = name + "_#_" + str(number)

### 2020-10-21 hive/app/domain/master_servers.py
Update master_servers.py

@@ -370,6 +370,13 @@ class SGMaster(Master):
             scenarios.close()
         super().__init__(simfile_name, sid, epochs, cluster_class, node_class)
 
+    def get_next_scenario(self, size_key: str) -> Tuple[np.ndarray, np.ndarray]:
+        if not DEBUG:
+            raise RuntimeError(
+                "get_next_scenario can not be called outside debug envs.")
+        d = SGMaster.SCENARIOS[size_key]
+        return np.ndarray(d["vectors"].pop()), np.ndarray(d["matrices"].pop())
+
     def get_cloud_reference(self) -> str:
         """Use to obtain a reference to 3rd party cloud storage provider
 

### 2020-10-21 hive/app/domain/master_servers.py
update docs

@@ -370,7 +370,8 @@ class SGMaster(Master):
             scenarios.close()
         super().__init__(simfile_name, sid, epochs, cluster_class, node_class)
 
-    def get_next_scenario(self, size_key: str) -> Tuple[np.ndarray, np.ndarray]:
+    @staticmethod
+    def get_next_scenario(size_key: str) -> Tuple[np.ndarray, np.ndarray]:
         if not DEBUG:
             raise RuntimeError(
                 "get_next_scenario can not be called outside debug envs.")

### 2020-10-21 hive/app/environment_settings.py
Create OPTIMIZE in environment settings

@@ -36,6 +36,8 @@ from typing import List
 from utils.convertions import truncate_float_value
 import numpy
 
+OPTIMIZE: bool = True
+
 DEBUG: bool = True
 """Indicates if some debug related actions or prints to the terminal should 
 be performed."""

### 2020-10-21 hive/app/domain/master_servers.py
return topology then vector in SGMaster.Scenarios

@@ -376,7 +376,7 @@ class SGMaster(Master):
             raise RuntimeError(
                 "get_next_scenario can not be called outside debug envs.")
         d = SGMaster.SCENARIOS[size_key]
-        return np.ndarray(d["vectors"].pop()), np.ndarray(d["matrices"].pop())
+        return np.ndarray(d["matrices"].pop()), np.ndarray(d["vectors"].pop())
 
     def get_cloud_reference(self) -> str:
         """Use to obtain a reference to 3rd party cloud storage provider

### 2020-10-21 hive/app/domain/cluster_groups.py
SGClusterPerfect deletes new_desired_distribution and override new_transition_matrix, also call super if OPT = True in env settings, else use MH only

@@ -982,34 +982,26 @@ class SGClusterPerfect(SGCluster):
         self.corruption_chances: List[float] = [0.0, 1.0]
 
     # region Swarm guidance structure management
-    def new_desired_distribution(
-            self, member_ids: List[str], member_uptimes: List[float]
-    ) -> np.ndarray:
-        """Creates a random desired distribution.
-
-        Overrides:
-            :py:meth:`app.domain.cluster_groups.SGCluster.new_desired_distribution`
-
-        Args:
-            member_ids:
-                A list of :py:attr:`node identifiers
-                <app.domain.network_nodes.Node.id>` who are
-                :py:attr:`~Cluster.members` of the ``SGCluster``.
-            member_uptimes:
-                This method's parameter is ignored and can be ``None``.
+    def new_transition_matrix(self) -> pd.DataFrame:
+        """Creates a new transition matrix that is likely to be a Markov Matrix.
 
         Returns:
-            :py:class:`~np:numpy.ndarray`:
-                A list of floats with which represent how the files should be
-                distributed among network nodes in the long-run.
+            :py:class:`~pd:pandas.DataFrame`:
+                The labeled matrix that has the fastests mixing rate from all
+                the pondered strategies.
         """
-        u_ = mm.new_vector(len(member_ids))
-        self.v_ = pd.DataFrame(data=u_, index=member_ids)
-        self.cv_ = pd.DataFrame(data=[0] * len(self.v_), index=member_ids)
-        self.avg_ = pd.DataFrame(data=[0] * len(self.v_), index=member_ids)
+        node_ids = [node.id for node in self.members.values()]
+
+        a, v_ = SGMaster.get_next_scenario(str(self.original_size))
+
+        self.v_ = pd.DataFrame(data=u_, index=node_ids)
+        self.cv_ = pd.DataFrame(data=[0] * len(self.v_), index=node_ids)
+        self.avg_ = pd.DataFrame(data=[0] * len(self.v_), index=node_ids)
         self._timer = 0
 
-        return u_
+        t = self.select_fastest_topology(a, v_)
+
+        return pd.DataFrame(t, index=node_ids, columns=node_ids)
     # endregion
 
     # region Simulation steps
@@ -1042,16 +1034,17 @@ class SGClusterPerfect(SGCluster):
     # endregion
 
     # region Helpers
-    """
     def select_fastest_topology(
[…]

### 2020-10-21 hive/app/domain/master_servers.py
Fixed import error

@@ -372,7 +372,7 @@ class SGMaster(Master):
 
     @staticmethod
     def get_next_scenario(size_key: str) -> Tuple[np.ndarray, np.ndarray]:
-        if not DEBUG:
+        if not es.DEBUG:
             raise RuntimeError(
                 "get_next_scenario can not be called outside debug envs.")
         d = SGMaster.SCENARIOS[size_key]

### 2020-10-21 hive/app/domain/cluster_groups.py
Fixed import error

@@ -992,7 +992,7 @@ class SGClusterPerfect(SGCluster):
         """
         node_ids = [node.id for node in self.members.values()]
 
-        a, v_ = SGMaster.get_next_scenario(str(self.original_size))
+        a, v_ = ms.SGMaster.get_next_scenario(str(self.original_size))
 
         self.v_ = pd.DataFrame(data=u_, index=node_ids)
         self.cv_ = pd.DataFrame(data=[0] * len(self.v_), index=node_ids)

### 2020-10-21 hive/app/domain/master_servers.py
Rever changes to SGMaster, moved Scenarios to Hive Simulation, to avoid restarts and have access to current iteration.

@@ -335,49 +335,6 @@ class Master:
 
 
 class SGMaster(Master):
-    SCENARIOS: Dict[str, Dict[str, List]] = {}
-
-    def __init__(self,
-                 simfile_name: str,
-                 sid: int,
-                 epochs: int,
-                 cluster_class: str,
-                 node_class: str) -> None:
-        """Instantiates an Master object.
-
-        Args:
-            simfile_name:
-                A path to the simulation file to be run by the simulator.
-            sid:
-                Identifier that generates unique output file names,
-                thus guaranteeing that different simulation instances do not
-                overwrite previous out files.
-            epochs:
-                The number of discrete time steps the simulation lasts.
-            cluster_class:
-                The name of the class used to instantiate cluster group
-                instances through reflection. See :py:mod:`cluster groups module
-                <app.domain.cluster_groups>`.
-            node_class:
-                The name of the class used to instantiate network node
-                instances through reflection. See :py:mod:`network nodes module
-                <app.domain.network_nodes>`.
-        """
-        if es.DEBUG:
-            scenarios_path = os.path.join(es.RESOURCES_ROOT, "scenarios.json")
-            scenarios = open(scenarios_path, "r")
-            SGMaster.SCENARIOS = json.load(scenarios)
-            scenarios.close()
-        super().__init__(simfile_name, sid, epochs, cluster_class, node_class)
-
-    @staticmethod
-    def get_next_scenario(size_key: str) -> Tuple[np.ndarray, np.ndarray]:
-        if not es.DEBUG:
-            raise RuntimeError(
-                "get_next_scenario can not be called outside debug envs.")
-        d = SGMaster.SCENARIOS[size_key]
-        return np.ndarray(d["matrices"].pop()), np.ndarray(d["vectors"].pop())
-
     def get_cloud_reference(self) -> str:
         """Use to obtain a reference to 3rd party cloud storage provider
 

### 2020-10-21 hive/app/hive_simulation.py
Update hive_simulation.py

@@ -39,9 +39,10 @@ import getopt
 from concurrent.futures.thread import ThreadPoolExecutor
 from typing import Optional
 
-import numpy
+import numpy as np
 import environment_settings as es
 
+from utils.convertions import class_name_to_obj
 from domain.helpers.matlab_utils import MatlabEngineContainer
 
 
@@ -49,11 +50,19 @@ __err_message__ = ("Invalid arguments. You must specify -f fname or -d, e.g.:\n"
                    "    $ python hive_simulation.py -f simfilename.json\n"
                    "    $ python hive_simulation.py -d")
 
+# region Sample Scenarios available for debug environments
+SCENARIOS: Dict[str, Dict[str, List]] = {}
+
+
+def get_next_scenario(k: str) -> Tuple[np.ndarray, np.ndarray]:
+    if not es.DEBUG:
+        warn("get_next_scenario should not be called outside debug envs.")
+    global SCENARIOS
+    return SCENARIOS[k]["matrices"].pop(), np.ndarray(SCENARIOS[k]["vectors"].pop())
+# endregion
+
 
 # region Module private functions (helpers)
-from utils.convertions import class_name_to_obj
-
-
 def __makedirs__() -> None:
     """Helper method that reates required simulation working directories if
     they do not exist."""
@@ -200,7 +209,7 @@ def main(threads_count: int,
 
     if threads_count != 0:
         _parallel_main(
-            numpy.abs(threads_count).item(), sdir, sname, iters, epochs)
+            np.abs(threads_count).item(), sdir, sname, iters, epochs)
     else:
 
         _single_main(sdir, sname, iters, epochs)
@@ -244,7 +253,6 @@ if __name__ == "__main__":
                 cluster_class = str(args).strip()
             if options in ("-n", "--network_node"):
                 node_class = str(args).strip()
-
     except getopt.GetoptError:
         sys.exit(__err_message__)
     except ValueError:
@@ -260,8 +268,20 @@ if __name__ == "__main__":
                  "Another cause of error might be a simulation file with "
                  "inconsistent values.")
 
+    try:
+        global SCENARIOS
+        scenarios_path = os.path.join(es.RESOURCES_ROOT, "scenarios.json")
+        scenarios = open(scenarios_path, "r")
+        # noinspection PyRedeclaration
+        SCENARIOS = json.load(scenarios)
+        scenarios.close()
+    except OSError:
+        warn(f"Could not load scenarios.json from {es.RESOURCES_ROOT}.\n"
[…]

### 2020-10-21 hive/app/domain/cluster_groups.py
Call get next scenario from hive_simulation.py

@@ -13,6 +13,7 @@ from tabulate import tabulate
 import numpy as np
 import pandas as pd
 import type_hints as th
+import hive_simulation as hs
 import environment_settings as es
 import domain.master_servers as ms
 import domain.helpers.enums as e
@@ -992,7 +993,7 @@ class SGClusterPerfect(SGCluster):
         """
         node_ids = [node.id for node in self.members.values()]
 
-        a, v_ = ms.SGMaster.get_next_scenario(str(self.original_size))
+        a, v_ = hs.get_next_scenario(str(self.original_size))
 
         self.v_ = pd.DataFrame(data=u_, index=node_ids)
         self.cv_ = pd.DataFrame(data=[0] * len(self.v_), index=node_ids)

### 2020-10-21 hive/app/hive_simulation.py
Update hive_simulation.py

@@ -57,7 +57,6 @@ SCENARIOS: Dict[str, Dict[str, List]] = {}
 def get_next_scenario(k: str) -> Tuple[np.ndarray, np.ndarray]:
     if not es.DEBUG:
         warn("get_next_scenario should not be called outside debug envs.")
-    global SCENARIOS
     return SCENARIOS[k]["matrices"].pop(), np.ndarray(SCENARIOS[k]["vectors"].pop())
 # endregion
 
@@ -269,7 +268,6 @@ if __name__ == "__main__":
                  "inconsistent values.")
 
     try:
-        global SCENARIOS
         scenarios_path = os.path.join(es.RESOURCES_ROOT, "scenarios.json")
         scenarios = open(scenarios_path, "r")
         # noinspection PyRedeclaration

### 2020-10-21 hive/app/hive_simulation.py
Fixed imports

@@ -35,9 +35,10 @@ Note:
 
 import os
 import sys
+import json
 import getopt
 from concurrent.futures.thread import ThreadPoolExecutor
-from typing import Optional
+from typing import Optional, Dict, Tuple, List
 
 import numpy as np
 import environment_settings as es

### 2020-10-21 hive/app/hive_simulation.py
__load_scenarios__ outside main to go around the global issue.

@@ -52,13 +52,30 @@ __err_message__ = ("Invalid arguments. You must specify -f fname or -d, e.g.:\n"
                    "    $ python hive_simulation.py -d")
 
 # region Sample Scenarios available for debug environments
-SCENARIOS: Dict[str, Dict[str, List]] = {}
+def __load_scenarios__():
+    scenarios = {}
+    try:
+        scenarios_path = os.path.join(es.RESOURCES_ROOT, "scenarios.json")
+        scenarios_file = open(scenarios_path, "r")
+        scenarios = json.load(scenarios_file)
+        scenarios_file.close()
+    except OSError:
+        warn(f"Could not load scenarios.json from {es.RESOURCES_ROOT}.\n"
+             " > if you need sample scenarios for swarm guidance in your code, "
+             "please refer to sample_scenario_generator.py, "
+             "otherwise, ignore this warning.")
+    return scenarios
+
+
+_scenarios: Dict[str, Dict[str, List]] = __load_scenarios__()
 
 
 def get_next_scenario(k: str) -> Tuple[np.ndarray, np.ndarray]:
     if not es.DEBUG:
         warn("get_next_scenario should not be called outside debug envs.")
-    return SCENARIOS[k]["matrices"].pop(), np.ndarray(SCENARIOS[k]["vectors"].pop())
+    topology = _scenarios[k]["matrices"].pop()
+    equilibrium = _scenarios[k]["vectors"].pop()
+    return topology, equilibrium
 # endregion
 
 
@@ -267,18 +284,6 @@ if __name__ == "__main__":
                  "  --network_node -n (str)\n"
                  "Another cause of error might be a simulation file with "
                  "inconsistent values.")
-
-    try:
-        scenarios_path = os.path.join(es.RESOURCES_ROOT, "scenarios.json")
-        scenarios = open(scenarios_path, "r")
-        # noinspection PyRedeclaration
-        SCENARIOS = json.load(scenarios)
-        scenarios.close()
-    except OSError:
-        warn(f"Could not load scenarios.json from {es.RESOURCES_ROOT}.\n"
-             " > if you need sample scenarios for swarm guidance in your code, "
-             "please refer to sample_scenario_generator.py, "
-             "otherwise, ignore this warning.")
 
     if simfile or simdirectory:
         main(threading, simdirectory, simfile, iterations, duration)

### 2020-10-21 hive/app/domain/cluster_groups.py
Fixed a bug where trying to get shape of list instead of nd array

@@ -994,8 +994,10 @@ class SGClusterPerfect(SGCluster):
         node_ids = [node.id for node in self.members.values()]
 
         a, v_ = hs.get_next_scenario(str(self.original_size))
+        print(a)
+        print(v_)
 
-        self.v_ = pd.DataFrame(data=u_, index=node_ids)
+        self.v_ = pd.DataFrame(data=v_, index=node_ids)
         self.cv_ = pd.DataFrame(data=[0] * len(self.v_), index=node_ids)
         self.avg_ = pd.DataFrame(data=[0] * len(self.v_), index=node_ids)
         self._timer = 0

### 2020-10-21 hive/app/hive_simulation.py

@@ -51,6 +51,7 @@ __err_message__ = ("Invalid arguments. You must specify -f fname or -d, e.g.:\n"
                    "    $ python hive_simulation.py -f simfilename.json\n"
                    "    $ python hive_simulation.py -d")
 
+
 # region Sample Scenarios available for debug environments
 def __load_scenarios__():
     scenarios = {}
@@ -73,8 +74,8 @@ _scenarios: Dict[str, Dict[str, List]] = __load_scenarios__()
 def get_next_scenario(k: str) -> Tuple[np.ndarray, np.ndarray]:
     if not es.DEBUG:
         warn("get_next_scenario should not be called outside debug envs.")
-    topology = _scenarios[k]["matrices"].pop()
-    equilibrium = _scenarios[k]["vectors"].pop()
+    topology = np.ndarray(_scenarios[k]["matrices"].pop())
+    equilibrium = np.ndarray(_scenarios[k]["vectors"].pop())
     return topology, equilibrium
 # endregion
 

### 2020-10-21 hive/app/hive_simulation.py
Fixed a bug where topology was being created as nparray instead of npasarray

@@ -74,8 +74,8 @@ _scenarios: Dict[str, Dict[str, List]] = __load_scenarios__()
 def get_next_scenario(k: str) -> Tuple[np.ndarray, np.ndarray]:
     if not es.DEBUG:
         warn("get_next_scenario should not be called outside debug envs.")
-    topology = np.ndarray(_scenarios[k]["matrices"].pop())
-    equilibrium = np.ndarray(_scenarios[k]["vectors"].pop())
+    topology = np.asarray(_scenarios[k]["matrices"].pop())
+    equilibrium = np.asarray(_scenarios[k]["vectors"].pop())
     return topology, equilibrium
 # endregion
 

### 2020-10-22 hive/app/domain/master_servers.py
Update master_servers.py

@@ -140,7 +140,7 @@ class Master:
                 cluster = self._new_cluster_group(cluster_class, size, fname)
 
                 filesize = os.path.getsize(os.path.join(es.SHARED_ROOT, fname))
-                es.set_blocks_size(int(filesize / es.BLOCKS_COUNT))
+                es.set_blocks_size(math.floor(filesize / es.BLOCKS_COUNT))
                 fblocks[fname] = self._split_files(fname, cluster, es.BLOCKS_SIZE)
 
                 # fblocks[fname] = self._split_files(fname, cluster, es.READ_SIZE)
@@ -417,7 +417,7 @@ class HDFSMaster(Master):
 
                 filesize = os.path.getsize(os.path.join(es.SHARED_ROOT, fname))
                 es.set_blocks_size(1 * 1024 * 1024)
-                es.set_blocks_count(int(filesize / es.BLOCKS_SIZE))
+                es.set_blocks_count(math.ceil(filesize / es.BLOCKS_SIZE))
                 fblocks[fname] = self._split_files(fname, cluster, es.BLOCKS_SIZE)
 
             # Distribute files before starting simulation

### 2020-10-22 hive/app/domain/cluster_groups.py
Restored ATOL to 5%.

@@ -918,8 +918,8 @@ class SGCluster(Cluster):
         """
         ptotal = self.file.existing_replicas
         target = self.v_.multiply(ptotal)
-        atol = np.clip(1 / self.original_size, 0.0, es.ATOL).item() * ptotal
-        return np.allclose(self.cv_, target, rtol=es.RTOL, atol=atol)
+        # atol = np.clip(1 / self.original_size, 0.0, es.ATOL).item() * ptotal
+        return np.allclose(self.cv_, target, rtol=es.RTOL, atol=es.ATOL * ptotal)
 
     def _log_evaluation(self, pcount: int, ptotal: int = -1) -> None:
         super()._log_evaluation(pcount, ptotal)
@@ -935,8 +935,8 @@ class SGCluster(Cluster):
         distance = np.abs(self.v_.subtract(self.avg_))
         magnitude = np.sqrt(distance).sum(axis=0).item()
 
-        atol = np.clip(1 / self.original_size, 0.0, es.ATOL).item()
-        goaled = np.allclose(self.avg_, self.v_, rtol=es.RTOL, atol=atol)
+        # atol = np.clip(1 / self.original_size, 0.0, es.ATOL).item()
+        goaled = np.allclose(self.avg_, self.v_, rtol=es.RTOL, atol=es.ATOL)
 
         self.file.logger.log_topology_goal_performance(goaled, magnitude)
 

### 2020-10-22 hive/app/domain/cluster_groups.py
Revert "Restored ATOL to 5%."
This reverts commit c620b247d175392077e979a16cde920fa6e9b269.

@@ -918,8 +918,8 @@ class SGCluster(Cluster):
         """
         ptotal = self.file.existing_replicas
         target = self.v_.multiply(ptotal)
-        # atol = np.clip(1 / self.original_size, 0.0, es.ATOL).item() * ptotal
-        return np.allclose(self.cv_, target, rtol=es.RTOL, atol=es.ATOL * ptotal)
+        atol = np.clip(1 / self.original_size, 0.0, es.ATOL).item() * ptotal
+        return np.allclose(self.cv_, target, rtol=es.RTOL, atol=atol)
 
     def _log_evaluation(self, pcount: int, ptotal: int = -1) -> None:
         super()._log_evaluation(pcount, ptotal)
@@ -935,8 +935,8 @@ class SGCluster(Cluster):
         distance = np.abs(self.v_.subtract(self.avg_))
         magnitude = np.sqrt(distance).sum(axis=0).item()
 
-        # atol = np.clip(1 / self.original_size, 0.0, es.ATOL).item()
-        goaled = np.allclose(self.avg_, self.v_, rtol=es.RTOL, atol=es.ATOL)
+        atol = np.clip(1 / self.original_size, 0.0, es.ATOL).item()
+        goaled = np.allclose(self.avg_, self.v_, rtol=es.RTOL, atol=atol)
 
         self.file.logger.log_topology_goal_performance(goaled, magnitude)
 

### 2020-10-22 hive/app/domain/cluster_groups.py
Keeping same behavior as before but always make everything depend on the number of parts explicitly

@@ -930,13 +930,13 @@ class SGCluster(Cluster):
 
     def _normalize_avg_(self):
         self.avg_ /= self._timer
-        self.avg_ /= np.sum(self.avg_)
-
-        distance = np.abs(self.v_.subtract(self.avg_))
+        distance = np.abs(self.v_.subtract(self.avg_ / np.sum(self.avg_)))
         magnitude = np.sqrt(distance).sum(axis=0).item()
 
-        atol = np.clip(1 / self.original_size, 0.0, es.ATOL).item()
-        goaled = np.allclose(self.avg_, self.v_, rtol=es.RTOL, atol=atol)
+        ptotal = es.BLOCKS_COUNT * es.REPLICATION_LEVEL
+        target = self.v_.multiply(ptotal)
+        atol = np.clip(1 / self.original_size, 0.0, es.ATOL).item() * ptotal
+        goaled = np.allclose(self.avg_, target, rtol=es.RTOL, atol=atol)
 
         self.file.logger.log_topology_goal_performance(goaled, magnitude)
 

### 2020-10-22 hive/app/environment_settings.py
OPT False and Blocks Count to 33

@@ -36,7 +36,7 @@ from typing import List
 from utils.convertions import truncate_float_value
 import numpy
 
-OPTIMIZE: bool = True
+OPTIMIZE: bool = False
 
 DEBUG: bool = True
 """Indicates if some debug related actions or prints to the terminal should 
@@ -58,7 +58,7 @@ def set_blocks_size(n: int) -> None:
     BLOCKS_SIZE = n
 
 
-BLOCKS_COUNT: int = 333
+BLOCKS_COUNT: int = 33
 """Defines into how many 
 :py:class:`~app.domain.helpers.smart_dataclasses.FileBlockData` instances a file
 is divided into. Either use this or :py:const:`BLOCKS_SIZE` but not both."""

### 2020-10-22 hive/app/domain/cluster_groups.py
Set loss chance to 0%

@@ -979,8 +979,8 @@ class SGClusterPerfect(SGCluster):
                  sim_id: int = 0,
                  origin: str = "") -> None:
         super().__init__(master, file_name, members, sim_id, origin)
-        # es.set_loss_chance(0.0)
         self.corruption_chances: List[float] = [0.0, 1.0]
+        es.set_loss_chance(0.0)
 
     # region Swarm guidance structure management
     def new_transition_matrix(self) -> pd.DataFrame:

### 2020-10-22 hive/app/environment_settings.py
Make it one thousand

@@ -58,7 +58,7 @@ def set_blocks_size(n: int) -> None:
     BLOCKS_SIZE = n
 
 
-BLOCKS_COUNT: int = 33
+BLOCKS_COUNT: int = 333
 """Defines into how many 
 :py:class:`~app.domain.helpers.smart_dataclasses.FileBlockData` instances a file
 is divided into. Either use this or :py:const:`BLOCKS_SIZE` but not both."""

### 2020-10-23 hive/app/environment_settings.py
Set optmize to True, to test 1000P Opt networks of size 8, 16 and 32

@@ -36,7 +36,7 @@ from typing import List
 from utils.convertions import truncate_float_value
 import numpy
 
-OPTIMIZE: bool = False
+OPTIMIZE: bool = True
 
 DEBUG: bool = True
 """Indicates if some debug related actions or prints to the terminal should 

### 2020-10-23 hive/app/domain/master_servers.py
Update master_servers.py

@@ -416,7 +416,7 @@ class HDFSMaster(Master):
                 cluster = self._new_cluster_group(cluster_class, size, fname)
 
                 filesize = os.path.getsize(os.path.join(es.SHARED_ROOT, fname))
-                es.set_blocks_size(1 * 1024 * 1024)
+                es.set_blocks_size(1 * 1024 * 1024)  # 1MB blocks.
                 es.set_blocks_count(math.ceil(filesize / es.BLOCKS_SIZE))
                 fblocks[fname] = self._split_files(fname, cluster, es.BLOCKS_SIZE)
 

### 2020-10-23 hive/app/environment_settings.py
Update environment_settings.py

@@ -42,7 +42,7 @@ DEBUG: bool = True
 """Indicates if some debug related actions or prints to the terminal should 
 be performed."""
 
-BLOCKS_SIZE: int = 131072
+BLOCKS_SIZE: int = 1 * 1024 * 1024
 """Defines the raw size of each file block before it's wrapped in a 
 :py:class:`~app.domain.helpers.smart_dataclasses.FileBlockData` instance 
 object. 

### 2020-10-23 hive/app/domain/cluster_groups.py
Update cluster_groups.py

@@ -994,8 +994,6 @@ class SGClusterPerfect(SGCluster):
         node_ids = [node.id for node in self.members.values()]
 
         a, v_ = hs.get_next_scenario(str(self.original_size))
-        print(a)
-        print(v_)
 
         self.v_ = pd.DataFrame(data=v_, index=node_ids)
         self.cv_ = pd.DataFrame(data=[0] * len(self.v_), index=node_ids)

### 2020-10-23 hive/app/hive_simulation.py
Fixed a bug where warn was not being imported from python's warning package

@@ -37,8 +37,10 @@ import os
 import sys
 import json
 import getopt
-from concurrent.futures.thread import ThreadPoolExecutor
+
+from warnings import warn
 from typing import Optional, Dict, Tuple, List
+from concurrent.futures.thread import ThreadPoolExecutor
 
 import numpy as np
 import environment_settings as es

### 2020-10-23 hive/app/sample_scenario_fixer.py
Added scenario generator and fixer

@@ -1,0 +1,63 @@
+import os
+import sys
+import json
+
+import numpy as np
+import pandas as pd
+import environment_settings as es
+import domain.helpers.matrices as mm
+
+
+def __validate_mc__(m: pd.DataFrame, v_: pd.DataFrame) -> bool:
+    """Asserts if the inputed Markov Matrix that converges to the desired equilibrium."""
+    t_pow = np.linalg.matrix_power(m.to_numpy(), 4096)
+    column_count = t_pow.shape[1]
+    for j in range(column_count):
+        test_target = t_pow[:, j]  # gets array column j
+        if not np.allclose(
+                test_target, v_[0].values, atol=1e-02):
+            return False
+    return True
+
+
+def __select_fastest_topology__(a: np.ndarray, v_: np.ndarray) -> np.ndarray:
+    """Emulates Swarm Guidance Clusters' fastest topology selection for MH algorithms."""
+    fastest_matrix, _ = mm.new_mh_transition_matrix(a, v_)
+    size = fastest_matrix.shape[0]
+    for j in range(size):
+        fastest_matrix[:, j] = np.absolute(fastest_matrix[:, j])
+        fastest_matrix[:, j] /= fastest_matrix[:, j].sum()
+    return fastest_matrix
+
+
+if __name__ == "__main__":
+    scenarios_path = os.path.join(es.RESOURCES_ROOT, "scenarios.json")
+    if not os.path.exists(scenarios_path):
+        sys.exit("Scenarios file does not exist.")
+
+    with open(scenarios_path, "r+") as f:
+        scenarios_dict = json.load(f)
+        for sk in scenarios_dict:
+            invalid_pair_indexes = []
+            samples_dict = scenarios_dict[sk]
+            matrices = samples_dict["matrices"]
+            vectors = samples_dict["vectors"]
+            for i in range(len(vectors)):
+                # print(f"Checking scenario #{i + 1}...")
+                a = np.asarray(matrices[i])
+                v = np.asarray(vectors[i])
+                m = __select_fastest_topology__(a, v)
+                is_valid_mc = __validate_mc__(pd.DataFrame(m), pd.DataFrame(v))
+                if not is_valid_mc:
+                    invalid_pair_indexes.append(i)
+            print(f"Found {len(invalid_pair_indexes)} invalid scenarios for size {sk}.")
+            invalid_pair_indexes.sort(reverse=True)
+            for i in invalid_pair_indexes:
+                print(f"Removing scenario entry #{i}")
[…]

### 2020-10-23 hive/app/sample_scenario_generator.py

@@ -1,3 +1,25 @@
+"""Creates an arbrirary number of symmetric connected topologies and equilibrium
+vectors that can be read during simulations for one to one comparison between
+algorithms. There is a small chance that generated pairs can not be solved by
+heuristic Markov chain generating algorithms such as our implementation of
+:py:meth:`Metropolis Hastings
+<app.domain.helpers.matrices._metropolis_hastings>`. To ensure that algorithm
+can be used over the generated pairs, run :py:mod:`sample_scenario_generator`,
+which removes all invalid entries from the generated json file.
+
+To execute this file run the following command (both arguments are optional)::
+
+    $ python sample_scenario_generator.py --samples=1000 --network_sizes=8,16,32
+
+Note:
+    The output of this script is a file named "scenarios.json" under the
+    :py:const:`~app.environment_settings.RESOURCES_ROOT` directory. If you wish
+    to modify this behavior you need to customize the script to accept one
+    additional argument which than saves the file under a different name. You
+    also need to ensure that all other uses of "scenarios.json" are changed
+    accordingly.
+"""
+
 import os
 import sys
 import ast

### 2020-10-23 hive/app/sample_scenario_fixer.py
Update docstrings

@@ -1,3 +1,20 @@
+"""Excludes all  <topologies, equilibrium> pairs in the ``scenarios.json`` file
+that are not synthetizable by our implementation of
+:py:meth:`Metropolis Hastings
+<app.domain.helpers.matrices._metropolis_hastings>`. Such JSON file is created
+using the script :py:mod:`sample_scenario_generator`.
+
+To execute this file run the following command::
+
+    $ python sample_scenario_generator.py
+
+Note:
+    Thhis script expects to fix a file named "scenarios.json" under the
+    :py:const:`~app.environment_settings.RESOURCES_ROOT` directory. If you wish
+    to modify this behavior you need to customize the script to accept one
+    additional argument which indicates the name of the file to be fixed.
+"""
+
 import os
 import sys
 import json

### 2020-10-23 hive/app/sample_scenario_generator.py

@@ -4,7 +4,7 @@ algorithms. There is a small chance that generated pairs can not be solved by
 heuristic Markov chain generating algorithms such as our implementation of
 :py:meth:`Metropolis Hastings
 <app.domain.helpers.matrices._metropolis_hastings>`. To ensure that algorithm
-can be used over the generated pairs, run :py:mod:`sample_scenario_generator`,
+can be used over the generated pairs, run :py:mod:`sample_scenario_fixer`,
 which removes all invalid entries from the generated json file.
 
 To execute this file run the following command (both arguments are optional)::

### 2020-10-23 hive/app/sample_scenario_fixer.py
Updated docs and webpage

@@ -9,7 +9,7 @@ To execute this file run the following command::
     $ python sample_scenario_generator.py
 
 Note:
-    Thhis script expects to fix a file named "scenarios.json" under the
+    This script expects to fix a file named "scenarios.json" under the
     :py:const:`~app.environment_settings.RESOURCES_ROOT` directory. If you wish
     to modify this behavior you need to customize the script to accept one
     additional argument which indicates the name of the file to be fixed.

### 2020-10-23 hive/app/environment_settings.py
Updated links in docs

@@ -53,7 +53,7 @@ Some possible values include { 32KB = 32768B; 128KB = 131072B; 512KB = 524288B;
 
 
 def set_blocks_size(n: int) -> None:
-    """Changes BLOCKS_SIZE constant value at run time to the given n bytes."""
+    """Changes :py:const:`BLOCKS_SIZE` constant value at run time to the given n bytes."""
     global BLOCKS_SIZE
     BLOCKS_SIZE = n
 
@@ -65,7 +65,7 @@ is divided into. Either use this or :py:const:`BLOCKS_SIZE` but not both."""
 
 
 def set_blocks_count(n: int) -> None:
-    """Changes NUMBER_PARTS constant value at run time."""
+    """Changes :py:const:`BLOCKS_COUNT` constant value at run time."""
     global BLOCKS_COUNT
     BLOCKS_COUNT = n
 
@@ -92,7 +92,7 @@ REPLICATION_LEVEL: int = 3
 
 
 def set_replication_level(n: int) -> None:
-    """Changes REPLICATION_LEVEL constant value at run time."""
+    """Changes :py:const:`REPLICATION_LEVEL` constant value at run time."""
     global REPLICATION_LEVEL
     REPLICATION_LEVEL = n
 
@@ -115,7 +115,7 @@ COMMUNICATION_CHANCES = [LOSS_CHANCE, DELIVER_CHANCE]
 
 
 def set_loss_chance(v: float) -> None:
-    """Changes LOSS_CHANCE constant value at run time."""
+    """Changes :py:const:`LOSS_CHANCE` constant value at run time."""
     global LOSS_CHANCE
     global DELIVER_CHANCE
     global COMMUNICATION_CHANCES

### 2020-10-23 hive/app/hive_simulation.py
Update hive_simulation.py

@@ -16,6 +16,12 @@ working threads. For example::
 
     $ python hive_simulation.py -d --iters=1 --threading=12
 
+Warning:
+     Be aware that python's
+    :py:class:`~py:concurrent.futures.ThreadPoolExecutor` conceals/suprresses
+    any uncaught exceptions, i.e., simulations may fail to execute or log items
+    properly and no debug information will be provided
+
 If you don't have a simulation file yet, run the following instead::
 
     $ python simfile_generator.py --file=filename.json

### 2020-10-23 hive/app/hive_simulation.py
Fix typo

@@ -17,8 +17,8 @@ working threads. For example::
     $ python hive_simulation.py -d --iters=1 --threading=12
 
 Warning:
-     Be aware that python's
-    :py:class:`~py:concurrent.futures.ThreadPoolExecutor` conceals/suprresses
+    Python's :py:class:`~py:concurrent.futures.ThreadPoolExecutor`
+    conceals/supresses
     any uncaught exceptions, i.e., simulations may fail to execute or log items
     properly and no debug information will be provided
 

### 2020-10-23 hive/app/make_docs.py
First version that conceals local paths in website

@@ -1,0 +1,43 @@
+"""Generates new sphinx html docs, removes local user's paths from the generated
+html, commits to the local repository and then updates docs at firebase website.
+If no repository exists or no remote firebase website exists the commands fail.
+"""
+import os
+import bs4
+
+app_html_path = os.path.abspath(
+    os.path.join(os.getcwd(), '..', 'docs', '_build', 'html', 'app.html'))
+
+db = r"\\"
+
+
+def remove_local_user_paths():
+    token = f"{db}hive{db}"
+    app_html_file = open(app_html_path, encoding="utf-8")
+    soup = bs4.BeautifulSoup(app_html_file.read(), "html.parser")
+    elements = soup.find_all("em", class_="property")
+    for tag in elements:
+        tag_text = tag.get_text()
+        if token in tag_text:
+            i = tag_text.find("'") + 1
+            j = tag_text.find(token) + 2
+            # exclude everything between the apostrophe and target token, discarding the first two backlashes
+            tag_text = f"{tag_text[0:i]}{tag_text[j:]}".replace(db, "/")
+            tag.string.replace_with(tag_text)
+            print(tag)
+    app_html_file.close()
+
+    html = soup.prettify("utf-8")
+    with open(app_html_path, "wb") as f:
+        f.write(html)
+
+
+if __name__ == "__main__":
+    # generate docs
+    # os.system("cd ..\\docs && make html")
+    # fix documents
+    remove_local_user_paths()
+    # save on remote
+    # os.system("git add -f * && git commit -m 'refresh docs' && git push")
+    # # update website
+    # os.system("firebase deploy")

### 2020-10-23 hive/app/make_docs.py
Improved version

@@ -13,23 +13,20 @@ db = r"\\"
 
 def remove_local_user_paths():
     token = f"{db}hive{db}"
-    app_html_file = open(app_html_path, encoding="utf-8")
-    soup = bs4.BeautifulSoup(app_html_file.read(), "html.parser")
-    elements = soup.find_all("em", class_="property")
-    for tag in elements:
-        tag_text = tag.get_text()
-        if token in tag_text:
-            i = tag_text.find("'") + 1
-            j = tag_text.find(token) + 2
-            # exclude everything between the apostrophe and target token, discarding the first two backlashes
-            tag_text = f"{tag_text[0:i]}{tag_text[j:]}".replace(db, "/")
-            tag.string.replace_with(tag_text)
-            print(tag)
-    app_html_file.close()
-
-    html = soup.prettify("utf-8")
-    with open(app_html_path, "wb") as f:
-        f.write(html)
+    with open(app_html_path, "rb+") as f:
+        soup = bs4.BeautifulSoup(f, "html.parser")
+        elements = soup.find_all("em", class_="property")
+        for tag in elements:
+            tag_text = tag.get_text()
+            if token in tag_text:
+                i = tag_text.find("'") + 1
+                j = tag_text.find(token) + 2
+                # exclude everything between the apostrophe and "hive"
+                tag_text = f"{tag_text[0:i]}{tag_text[j:]}".replace(db, "/")
+                tag.string.replace_with(tag_text)
+        f.seek(0)
+        f.write(soup.prettify("utf-8"))
+        f.truncate()
 
 
 if __name__ == "__main__":

### 2020-10-23 hive/app/make_docs.py
Decommented system calls

@@ -31,10 +31,10 @@ def remove_local_user_paths():
 
 if __name__ == "__main__":
     # generate docs
-    # os.system("cd ..\\docs && make html")
+    os.system("cd ..\\docs && make html")
     # fix documents
     remove_local_user_paths()
     # save on remote
-    # os.system("git add -f * && git commit -m 'refresh docs' && git push")
+    os.system("git add -f * && git commit -m 'refresh docs' && git push")
     # # update website
-    # os.system("firebase deploy")
+    os.system("firebase deploy")

### 2020-10-23 hive/app/make_docs.py
Update make_docs.py

@@ -31,7 +31,7 @@ def remove_local_user_paths():
 
 if __name__ == "__main__":
     # generate docs
-    os.system("cd ..\\docs && make html")
+    os.system("cd ../docs && make html && cd ..")
     # fix documents
     remove_local_user_paths()
     # save on remote

### 2020-10-23 hive/app/make_docs.py
Update make_docs.py

@@ -31,10 +31,10 @@ def remove_local_user_paths():
 
 if __name__ == "__main__":
     # generate docs
-    os.system("cd ../docs && make html && cd ..")
+    os.system("cd ../docs && make html")
     # fix documents
     remove_local_user_paths()
     # save on remote
-    os.system("git add -f * && git commit -m 'refresh docs' && git push")
+    os.system("git add -f ../docs/* && git commit -m 'refresh docs' && git push")
     # # update website
-    os.system("firebase deploy")
+    # os.system("firebase deploy")

### 2020-10-23 hive/app/make_docs.py
Delete make_docs.py

@@ -1,40 +1,0 @@
-"""Generates new sphinx html docs, removes local user's paths from the generated
-html, commits to the local repository and then updates docs at firebase website.
-If no repository exists or no remote firebase website exists the commands fail.
-"""
-import os
-import bs4
-
-app_html_path = os.path.abspath(
-    os.path.join(os.getcwd(), '..', 'docs', '_build', 'html', 'app.html'))
-
-db = r"\\"
-
-
-def remove_local_user_paths():
-    token = f"{db}hive{db}"
-    with open(app_html_path, "rb+") as f:
-        soup = bs4.BeautifulSoup(f, "html.parser")
-        elements = soup.find_all("em", class_="property")
-        for tag in elements:
-            tag_text = tag.get_text()
-            if token in tag_text:
-                i = tag_text.find("'") + 1
-                j = tag_text.find(token) + 2
-                # exclude everything between the apostrophe and "hive"
-                tag_text = f"{tag_text[0:i]}{tag_text[j:]}".replace(db, "/")
-                tag.string.replace_with(tag_text)
-        f.seek(0)
-        f.write(soup.prettify("utf-8"))
-        f.truncate()
-
-
-if __name__ == "__main__":
-    # generate docs
-    os.system("cd ../docs && make html")
-    # fix documents
-    remove_local_user_paths()
-    # save on remote
-    os.system("git add -f ../docs/* && git commit -m 'refresh docs' && git push")
-    # # update website
-    # os.system("firebase deploy")

### 2020-10-23 hive/app/make_docs.py
Create make_docs.py

@@ -1,0 +1,41 @@
+"""Generates new sphinx html docs, removes local user's paths from the generated
+html, commits to the local repository and then updates docs at firebase website.
+If no repository exists or no remote firebase website exists the commands fail.
+"""
+import os
+
+import bs4
+
+app_html_path = os.path.abspath(
+    os.path.join(os.getcwd(), '..', 'docs', '_build', 'html', 'app.html'))
+
+db = r'\\'
+
+
+def remove_local_user_paths():
+    token = f'{db}hive{db}'
+    with open(app_html_path, 'rb+') as f:
+        soup = bs4.BeautifulSoup(f, 'html.parser')
+        elements = soup.find_all('em', class_='property')
+        for tag in elements:
+            tag_text = tag.get_text()
+            if token in tag_text:
+                i = tag_text.find("'") + 1
+                j = tag_text.find(token) + 2
+                # exclude everything between the apostrophe and "hive"
+                tag_text = f'{tag_text[0:i]}{tag_text[j:]}'.replace(db, '/')
+                tag.string.replace_with(tag_text)
+        f.seek(0)
+        f.write(soup.prettify('utf-8'))
+        f.truncate()
+
+
+if __name__ == "__main__":
+    # generate docs
+    os.system('cd ../docs && make html')
+    # fix documents
+    remove_local_user_paths()
+    # save on remote
+    os.system('git add * && git commit -m "refresh docs" && git push')
+    # update website
+    os.system('firebase deploy')

### 2020-10-23 hive/app/hive_simulation.py
Code works.

@@ -18,9 +18,8 @@ working threads. For example::
 
 Warning:
     Python's :py:class:`~py:concurrent.futures.ThreadPoolExecutor`
-    conceals/supresses
-    any uncaught exceptions, i.e., simulations may fail to execute or log items
-    properly and no debug information will be provided
+    conceals/supresses any uncaught exceptions, i.e., simulations may fail to
+    execute or log items properly and no debug information will be provided
 
 If you don't have a simulation file yet, run the following instead::
 

### 2020-10-23 hive/app/make_docs.py

@@ -36,6 +36,6 @@ if __name__ == "__main__":
     # fix documents
     remove_local_user_paths()
     # save on remote
-    os.system('git add * && git commit -m "refresh docs" && git push')
+    os.system('git add ../docs/* && git commit -m "refresh docs" && git push')
     # update website
     os.system('firebase deploy')

### 2020-10-24 hive/app/environment_settings.py
Set debug false, optimize true, blocks count 46 for SGExt tests

@@ -36,9 +36,9 @@ from typing import List
 from utils.convertions import truncate_float_value
 import numpy
 
-OPTIMIZE: bool = False
+OPTIMIZE: bool = True
 
-DEBUG: bool = True
+DEBUG: bool = False
 """Indicates if some debug related actions or prints to the terminal should 
 be performed."""
 
@@ -58,7 +58,7 @@ def set_blocks_size(n: int) -> None:
     BLOCKS_SIZE = n
 
 
-BLOCKS_COUNT: int = 333
+BLOCKS_COUNT: int = 46
 """Defines into how many 
 :py:class:`~app.domain.helpers.smart_dataclasses.FileBlockData` instances a file
 is divided into. Either use this or :py:const:`BLOCKS_SIZE` but not both."""

### 2020-10-25 hive/app/hive_simulation.py
Create new param in hive_simulation

@@ -246,6 +246,7 @@ if __name__ == "__main__":
     threading = 0
     simdirectory = False
     simfile = None
+    start_iteration = 1
     iterations = 1
     duration = 480
 
@@ -253,23 +254,27 @@ if __name__ == "__main__":
     cluster_class = "SGClusterExt"
     node_class = "SGNodeExt"
 
-    short_opts = "df:i:t:e:m:c:n:"
-    long_opts = ["directory", "file=", "iters=", "threading=", "epochs=",
+    short_opts = "df:i:S:t:e:m:c:n:"
+    long_opts = ["directory", "file=",
+                 "iterations=", "start_iteration="
+                 "threading=", "epochs=",
                  "master_server=", "cluster_group=", "network_node="]
 
     try:
         options, args = getopt.getopt(sys.argv[1:], short_opts, long_opts)
         for options, args in options:
-            if options in ("-t", "--threading"):
-                threading = int(str(args).strip())
             if options in ("-d", "--directory"):
                 simdirectory = True
             elif options in ("-f", "--file"):
                 simfile = str(args).strip()
                 if simfile == "":
                     sys.exit("Simulation file name can not be blank.")
-            if options in ("-i", "--iters"):
+            if options in ("-i", "--iterations"):
                 iterations = int(str(args).strip())
+            if options in ("-S", "--start_iteration"):
+                start_iteration = int(str(args).strip())
+            if options in ("-t", "--threading"):
+                threading = int(str(args).strip())
             if options in ("-e", "--epochs"):
                 duration = int(str(args).strip())
             if options in ("-m", "--master_server"):
@@ -282,14 +287,15 @@ if __name__ == "__main__":
         sys.exit(__err_message__)
     except ValueError:
         sys.exit("Execution arguments should have the following data types:\n"
-                 "  --iterations -i (int)\n"
-                 "  --epochs -e (int)\n"
-                 "  --threading -t (int)\n"
                  "  --directory -d (void)\n"
-                 "  --file -f (str)\n"
-                 "  --master_server -m (str)\n"
-                 "  --cluster_group -c (str)\n"
-                 "  --network_node -n (str)\n"
+                 "  --iterations= -i (int)\n"
+                 "  --start_iteration= -S (int)\n"
+                 "  --epochs= -e (int)\n"
+                 "  --threading= -t (int)\n"
+                 "  --file= -f (str)\n"
+                 "  --master_server= -m (str)\n"
[…]

### 2020-10-25 hive/app/hive_simulation.py
Improve hive_simulation.py

@@ -2,7 +2,7 @@
 
 You can start a simulation by executing the following command::
 
-    $ python hive_simulation.py --file=a_simulation_name.json --iters=30
+    $ python hive_simulation.py --file=a_simulation_name.json --iterations=30
 
 You can also execute all simulation file that exist in
 :py:const:`~app.environment_settings.SIMULATION_ROOT` by instead executing::
@@ -14,7 +14,7 @@ can use the -t or --threading flag in either of the previously specified
 commands. The threading flag expects an integer that specifies the max
 working threads. For example::
 
-    $ python hive_simulation.py -d --iters=1 --threading=12
+    $ python hive_simulation.py -d --iterations=1 --threading=2
 
 Warning:
     Python's :py:class:`~py:concurrent.futures.ThreadPoolExecutor`
@@ -33,7 +33,7 @@ Note:
         2. Any file used by the simulation, e.g., a picture or a .pptx \
         document is accessible in \
         :py:const:`~app.environment_settings.SHARED_ROOT`.
-        3. An output file simdirectory exists with default path being: \
+        3. An output file directory exists with default path being: \
         :py:const:`~app.environment_settings.OUTFILE_ROOT`.
 
 """
@@ -87,7 +87,7 @@ def get_next_scenario(k: str) -> Tuple[np.ndarray, np.ndarray]:
 # endregion
 
 
-# region Module private functions (helpers)
+# region Helpers
 def __makedirs__() -> None:
     """Helper method that reates required simulation working directories if
     they do not exist."""
@@ -97,186 +97,105 @@ def __makedirs__() -> None:
     os.makedirs(es.RESOURCES_ROOT, exist_ok=True)
 
 
-def __can_exec_simfile__(sname: str) -> None:
+def __list_dir__() -> List[str]:
+    target_dir = os.listdir(es.SIMULATION_ROOT)
+    return list(filter(lambda x: "scenarios" not in x, target_dir))
+
+
+def _validate_simfile(simfile_name: str) -> None:
     """Asserts if simulation can proceed with user specified file.
 
     Args:
-        sname:
+        simfile_name:
             The name of the simulation file, including extension,
             whose existence inside
             :py:const:`~app.environment_settings.SIMULATION_ROOT` will be
             checked.
     """
-    spath = os.path.join(es.SIMULATION_ROOT, sname)
+    spath = os.path.join(es.SIMULATION_ROOT, simfile_name)
[…]

### 2020-10-25 hive/app/hive_simulation.py
Update hive_simulation.py with the use of Start Iteration

@@ -133,28 +133,28 @@ def _start_simulation(simfile_name: str, sid: int) -> None:
     master_server.execute_simulation()
 
 
-def _parallel_main() -> None:
+def _parallel_main(start: int, stop: int) -> None:
     """Helper method that initializes a multi-threaded simulation."""
     with ThreadPoolExecutor(max_workers=threading) as executor:
         if directory:
             for simfile_name in __list_dir__():
-                for i in range(1, iterations + 1):
+                for i in range(start, stop):
                     executor.submit(_start_simulation, simfile_name, i, epochs)
         else:
             _validate_simfile(simfile)
-            for i in range(1, iterations + 1):
+            for i in range(start, stop):
                 executor.submit(_start_simulation, simfile, i, epochs)
 
 
-def _single_main() -> None:
+def _single_main(start: int, stop: int) -> None:
     """Helper function that initializes a single-threaded simulation."""
     if directory:
         for simfile_name in __list_dir__():
-            for i in range(1, iterations + 1):
+            for i in range(start, stop):
                 _start_simulation(simfile_name, i)
     else:
         _validate_simfile(simfile)
-        for i in range(1, iterations + 1):
+        for i in range(start, stop):
             _start_simulation(simfile, i)
 # endregion
 
@@ -226,4 +226,7 @@ if __name__ == "__main__":
 
     MatlabEngineContainer.get_instance()
     threading = np.ceil(np.abs(threading)).item()
-    _single_main() if threading in (0, 1) else _parallel_main()
+
+    s = start_iteration
+    st = start_iteration + iterations + 1
+    _single_main(s, st) if threading in (0, 1) else _parallel_main(s, st)

### 2020-10-25 hive/app/hive_simulation.py
Update hive_simulation.py

@@ -44,7 +44,7 @@ import json
 import getopt
 
 from warnings import warn
-from typing import Optional, Dict, Tuple, List
+from typing import Dict, Tuple, List
 from concurrent.futures.thread import ThreadPoolExecutor
 
 import numpy as np
@@ -79,6 +79,22 @@ _scenarios: Dict[str, Dict[str, List]] = __load_scenarios__()
 
 
 def get_next_scenario(k: str) -> Tuple[np.ndarray, np.ndarray]:
+    """Function used for one-to-one testing of different swarm guidance
+    configurations.
+
+    Args:
+        k:
+            A string identifying the pool of matrix, vector pairs to get the
+            scenario. Usually, a string representation of an integer which
+            corresponds to the network size being tested.
+    Note:
+        This method should only be used when
+        :py:const:`app.environment_settings.DEBUG` is set to True.
+
+    Returns:
+        A topology matrix and a random equilibrium vector that can be used
+        to generate Markov chains used for Swarm Guidance.
+    """
     if not es.DEBUG:
         warn("get_next_scenario should not be called outside debug envs.")
     topology = np.asarray(_scenarios[k]["matrices"].pop())
@@ -134,7 +150,18 @@ def _start_simulation(simfile_name: str, sid: int) -> None:
 
 
 def _parallel_main(start: int, stop: int) -> None:
-    """Helper method that initializes a multi-threaded simulation."""
+    """Helper method that initializes a multi-threaded simulation.
+
+    Args:
+        start:
+            A number that marks the first desired identifier for the
+            simulations that will execute.
+        stop:
+            A number that marks the last desired identifier for the
+            simulations that will execute. Usually a sum of ``start`` and the
+            total number of iterations specified by the user in the scripts'
+            arguments.
+    """
     with ThreadPoolExecutor(max_workers=threading) as executor:
         if directory:
             for simfile_name in __list_dir__():
@@ -147,7 +174,18 @@ def _parallel_main(start: int, stop: int) -> None:
 
 
 def _single_main(start: int, stop: int) -> None:
-    """Helper function that initializes a single-threaded simulation."""
+    """Helper function that initializes a single-threaded simulation.
+
+    Args:
+        start:
+            A number that marks the first desired identifier for the
+            simulations that will execute.
+        stop:
+            A number that marks the last desired identifier for the
+            simulations that will execute. Usually a sum of ``start`` and the
+            total number of iterations specified by the user in the scripts'
+            arguments.
[…]

### 2020-10-25 hive/app/hive_simulation.py
update docs

@@ -82,14 +82,15 @@ def get_next_scenario(k: str) -> Tuple[np.ndarray, np.ndarray]:
     """Function used for one-to-one testing of different swarm guidance
     configurations.
 
+    Note:
+        This method should only be used when
+        :py:const:`app.environment_settings.DEBUG` is set to True.
+
     Args:
         k:
             A string identifying the pool of matrix, vector pairs to get the
             scenario. Usually, a string representation of an integer which
             corresponds to the network size being tested.
-    Note:
-        This method should only be used when
-        :py:const:`app.environment_settings.DEBUG` is set to True.
 
     Returns:
         A topology matrix and a random equilibrium vector that can be used

### 2020-10-25 hive/app/hive_simulation.py
Fixed a bug in our hive_simulation,py

@@ -241,9 +241,7 @@ if __name__ == "__main__":
                 cluster_class = str(args).strip()
             if options in ("-n", "--network_node"):
                 node_class = str(args).strip()
-    except getopt.GetoptError:
-        sys.exit(__err_message__)
-    except ValueError:
+    except (getopt.GetoptError, ValueError):
         sys.exit("Execution arguments should have the following data types:\n"
                  "  --directory -d (void)\n"
                  "  --iterations= -i (int)\n"
@@ -257,10 +255,9 @@ if __name__ == "__main__":
                  "Another cause of error might be a simulation file with "
                  "inconsistent values.")
 
-    if not simfile or not directory:
+    if simfile is None and not directory:
         sys.exit(__err_message__)
-
-    if not directory and simfile == "":
+    elif simfile == "" and not directory:
         sys.exit("File name can not be blank. Unless directory option is True.")
 
     MatlabEngineContainer.get_instance()

### 2020-10-25 hive/app/hive_simulation.py
Update hive_simulation.py

@@ -265,4 +265,4 @@ if __name__ == "__main__":
 
     s = start_iteration
     st = start_iteration + iterations + 1
-    _single_main(s, st) if threading in (0, 1) else _parallel_main(s, st)
+    _single_main(s, st) if threading in {0, 1} else _parallel_main(s, st)

### 2020-10-25 hive/app/hive_simulation.py
Update hive_simulation.py

@@ -42,6 +42,8 @@ import os
 import sys
 import json
 import getopt
+import traceback
+import concurrent.futures
 
 from warnings import warn
 from typing import Dict, Tuple, List
@@ -57,6 +59,10 @@ from domain.helpers.matlab_utils import MatlabEngineContainer
 __err_message__ = ("Invalid arguments. You must specify -f fname or -d, e.g.:\n"
                    "    $ python hive_simulation.py -f simfilename.json\n"
                    "    $ python hive_simulation.py -d")
+
+__log_thread_errors__: bool = True
+"""Wether or not the script should crash if a ThreadPoolExecutor 
+fails due to an exception and if the exception traceback should be provided."""
 
 
 # region Sample Scenarios available for debug environments
@@ -134,7 +140,7 @@ def _validate_simfile(simfile_name: str) -> None:
         sys.exit(f"The simulation file does not exist in {es.SIMULATION_ROOT}.")
 
 
-def _start_simulation(simfile_name: str, sid: int) -> None:
+def _simulate(simfile_name: str, sid: int) -> None:
     """Helper method that orders execution of one simulation instance.
 
     Args:
@@ -164,14 +170,22 @@ def _parallel_main(start: int, stop: int) -> None:
             arguments.
     """
     with ThreadPoolExecutor(max_workers=threading) as executor:
+        futures = []
         if directory:
             for simfile_name in __list_dir__():
                 for i in range(start, stop):
-                    executor.submit(_start_simulation, simfile_name, i, epochs)
+                    futures.append(executor.submit(_simulate, simfile_name, i))
         else:
             _validate_simfile(simfile)
             for i in range(start, stop):
-                executor.submit(_start_simulation, simfile, i, epochs)
+                futures.append(executor.submit(_simulate, simfile, i))
+
+        if __log_thread_errors__:
+            for future in concurrent.futures.as_completed(futures):
+                try:
+                    future.result()
+                except Exception:
+                    sys.exit(traceback.print_exc())
 
 
 def _single_main(start: int, stop: int) -> None:
@@ -190,11 +204,11 @@ def _single_main(start: int, stop: int) -> None:
     if directory:
         for simfile_name in __list_dir__():
             for i in range(start, stop):
-                _start_simulation(simfile_name, i)
+                _simulate(simfile_name, i)
     else:
         _validate_simfile(simfile)
         for i in range(start, stop):
-            _start_simulation(simfile, i)
+            _simulate(simfile, i)
 # endregion
 
 

### 2020-10-25 hive/app/scripts/python/plotting/swarm_guidance_analysis.py
Update swarm_guidance_analysis.py

@@ -22,9 +22,6 @@ def setup_sources(source_patterns: List[str]) -> Tuple[Dict[str, Any], List[str]
             if key in filename:
                 sources_files[key].append(filename)
                 break
-
-    # remove '-' and 'P' delimiters that avoid incorrect missmapping of sources
-    # sources_files = {k[1:-1]: v for k, v in sources_files.items()}
     source_keys = list(sources_files)
     return sources_files, source_keys
 
@@ -357,6 +354,10 @@ def piechart_avg_convergence_achieved(figname: str = "GA") -> None:
 # endregion
 
 
+def boxplot_time_to_detect_off_nodes(figname: str = "TSNR") -> None:
+    pass
+
+
 if __name__ == "__main__":
     # region args processing
     ns = 8
@@ -410,8 +411,7 @@ if __name__ == "__main__":
     subtitle = f"Cluster size: {ns}, Opt.: {opt}, P(de): {pde}%, P(ml): {pml}%"
     # endregion
 
-    source_patterns = ["SG8-100P", "SG8-1000P", "SG8-2000P"]
-    sources_files, source_keys = setup_sources(source_patterns)
+    sources_files, source_keys = setup_sources(["SG8-100P", "SG8-1000P", "SG8-2000P"])
     # Q1. Quantas mensagens passam na rede por epoch?
     boxplot_bandwidth(figname="bw_parts")
     # Q2. Existem mais conjuntos de convergencia perto do fim da simulação?
@@ -425,18 +425,14 @@ if __name__ == "__main__":
     boxplot_percent_time_instantaneous_convergence(figname="tic_parts")
     # Q6. Tecnicas de optimização influenciam as questões anteriores?
 
-    source_patterns = ["SG8-1000P", "SG8-Opt"]
-    # noinspection PyRedeclaration
-    sources_files, source_keys = setup_sources(source_patterns)
+    sources_files, source_keys = setup_sources(["SG8-1000P", "SG8-Opt"])
     barchart_instantaneous_convergence_vs_progress(bucket_size=5, figname="icp_opt")
     boxplot_first_convergence(figname="fc_opt")
     piechart_avg_convergence_achieved(figname="avgc_pie_opt")
     boxplot_avg_convergence_magnitude_distance(figname="avgc_dist_opt")
     boxplot_percent_time_instantaneous_convergence(figname="tic_opt")
 
[…]

### 2020-10-25 hive/app/scripts/python/plotting/_matplotlib_configs.py
Added color_palette to _matplotlib_configs

@@ -1,9 +1,13 @@
 import matplotlib.style
+import matplotlib.pyplot
+
 import matplotlib as mpl
 
 from matplotlib.font_manager import FontProperties
 
 mpl.style.use('seaborn-deep')
+
+color_palette = matplotlib.pyplot.style.library['seaborn-deep']
 
 title_pad = 20
 labels_pad = 15
